---
title: "Multi-Reference Indicator Comparison: `r params$candidate_name`"
author: "Delphi Group"
date: today
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
params:
  # Candidate Indicator
  candidate_source: doctor-visits
  candidate_indicator: smoothed_adj_cli
  candidate_name: "Doctor Visits: Smoothed Adj CLI"
  candidate_csv: !expr "NULL"
  # Reference Indicators
  reference_sources: !expr 'c("quidel", "hospital-admissions", "chng", "chng")'
  reference_indicators: !expr 'c("covid_ag_smoothed_pct_positive", "smoothed_covid19_from_claims", "7dav_inpatient_covid", "7dav_outpatient_covid")'
  reference_names: !expr 'c("Quidel: Smoothed % Positive COVID Antigen", "Hospital Admissions: Smoothed COVID-19 from Claims", "CHNG: 7-day Avg Inpatient COVID", "CHNG: 7-day Avg Outpatient COVID")'
  reference_csvs: !expr "NULL"
  # Shared settings
  geo_type: state
  time_type: day
  start_day: "2020-09-01"
  end_day: "2023-03-01"
  # Maximum locations to show in plots/tables
  max_locations_plot: 12
  api_timeout_seconds: 60
  api_chunk_days: 30
knitr:
  opts_chunk:
    comment: "#>"
    message: FALSE
    warning: FALSE
    out.width: "100%"
    dev: "png"
    fig.width: 9
    fig.height: 5
    fig.align: center
editor_options:
  chunk_output_type: console
execute:
  cache: true
---

```{r}
#| label: setup
#| include: false
#| cache: false
if (!exists("params")) {
    params <- list(
        candidate_source = "doctor-visits",
        candidate_indicator = "smoothed_adj_cli",
        candidate_name = "Doctor Visits: Smoothed Adj CLI",
        candidate_csv = NULL,
        reference_sources = c("quidel", "hospital-admissions", "chng", "chng"),
        reference_indicators = c(
            "covid_ag_smoothed_pct_positive",
            "smoothed_covid19_from_claims",
            "7dav_inpatient_covid",
            "7dav_outpatient_covid"
        ),
        reference_names = c(
            "Quidel: Smoothed % Positive COVID Antigen",
            "Hospital Admissions: Smoothed COVID-19 from Claims",
            "CHNG: 7-day Avg Inpatient COVID",
            "CHNG: 7-day Avg Outpatient COVID"
        ),
        reference_csvs = NULL,
        geo_type = "state",
        time_type = "day",
        start_day = "2020-09-01",
        end_day = "2023-03-01",
        max_locations_plot = 12,
        api_timeout_seconds = 60,
        api_chunk_days = 30
    )
}

options(repos = c(
    delphi = "https://cmu-delphi.r-universe.dev",
    CRAN = "https://cloud.r-project.org"
))
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    epidatr, epiprocess, covidcast,
    tidyverse, scales, patchwork,
    here, glue, DT, rlang
)

knitr::opts_chunk$set(cache.extra = list(params))

colours <- c(
    "#172431",
    "#ffffff",
    "#C41230",
    "#941120",
    "#007BC0",
    "#002676",
    "#FDB515",
    "#002145",
    "#00A7E1",
    "#8C338F",
    "#368f33"
)
dis_pal_list <- colours[c(3, 5, 7, 10, 11, 9)]

theme_set(
    theme_minimal(base_family = "sans", base_size = 14) +
        theme(
            plot.title = element_text(hjust = 0.5, size = 17, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5, size = 13, color = "gray45"),
            plot.caption = element_text(hjust = 0, size = 11, color = "gray45"),
            axis.title = element_text(size = 14),
            axis.text = element_text(size = 12),
            panel.grid = element_line(linetype = 2, linewidth = 0.3, color = "gray90"),
            panel.grid.minor = element_blank(),
            legend.position = "top",
            legend.title = element_text(size = 12),
            legend.text = element_text(size = 11),
            strip.background = element_rect(fill = "gray95", linetype = "blank"),
            strip.text = element_text(size = 12, face = "bold"),
            panel.border = element_rect(color = "gray95", fill = NA)
        )
)

options(
    ggplot2.discrete.colour = dis_pal_list,
    ggplot2.discrete.fill = dis_pal_list,
    ggplot2.continuous.colour = function() scale_colour_gradientn(colours = dis_pal_list),
    ggplot2.continuous.fill = function() scale_fill_gradientn(colours = dis_pal_list)
)

round_to_sunday <- function(d) d - as.integer(format(d, "%w"))

time_range <- function(start, end, time_type = params$time_type) {
    if (time_type == "week") {
        to_yw <- function(d) {
            d <- as.Date(d)
            m <- MMWRweek::MMWRweek(d)
            m$MMWRyear * 100L + m$MMWRweek
        }
        epirange(to_yw(start), to_yw(end))
    } else {
        epirange(start, end)
    }
}

geo_to_name <- function(geos, geo_type = params$geo_type) {
    if (geo_type == "state") {
        names <- covidcast::abbr_to_name(toupper(geos))
    } else if (geo_type == "county") {
        names <- covidcast::county_fips_to_name(as.character(geos))
    } else if (geo_type == "msa") {
        names <- covidcast::cbsa_to_name(as.character(geos))
    } else {
        return(geos)
    }
    names[is.na(names)] <- geos[is.na(names)]
    names
}

random_sample <- function(vec, n = 15) {
    if (length(vec) <= n) {
        return(vec)
    }
    withr::with_seed(42, sort(sample(vec, n)))
}

wrap_reference <- function(x, width = 30) {
    stringr::str_wrap(x, width = width)
}

reference_palette <- function(refs) {
    setNames(rep(dis_pal_list, length.out = length(refs)), refs)
}

lag_palette <- function(n) {
    grDevices::hcl.colors(n, palette = "Dark 3")
}

wrap_legend_label <- function(x, width = 26) {
    stringr::str_wrap(x, width = width)
}

safe_bottom_legend <- function(text_size = 10) {
    theme(
        legend.position = "bottom",
        legend.box = "vertical",
        legend.justification = "center",
        legend.text = element_text(size = text_size),
        legend.title = element_text(size = text_size + 1),
        legend.key.width = grid::unit(1.25, "lines"),
        legend.key.height = grid::unit(0.9, "lines"),
        legend.margin = margin(t = 2, r = 2, b = 2, l = 2),
        legend.box.margin = margin(t = 2, r = 2, b = 2, l = 2)
    )
}

safe_spearman <- function(x, y, min_n = 3) {
    ok <- complete.cases(x, y)
    if (sum(ok) < min_n || n_distinct(x[ok]) < 2 || n_distinct(y[ok]) < 2) {
        return(NA_real_)
    }
    suppressWarnings(cor(x[ok], y[ok], method = "spearman"))
}

zscore_safe <- function(x) {
    s <- sd(x, na.rm = TRUE)
    m <- mean(x, na.rm = TRUE)
    if (!is.finite(s) || s == 0 || !is.finite(m)) {
        return(rep(NA_real_, length(x)))
    }
    (x - m) / s
}

add_lag_units <- function(d, lag, time_type) {
    if (time_type == "week") {
        d + lubridate::weeks(lag)
    } else if (time_type == "month") {
        d + lubridate::months(lag)
    } else if (time_type == "year") {
        d + lubridate::years(lag)
    } else {
        d + lag
    }
}

new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")

date_chunks <- function(start, end, chunk_days = params$api_chunk_days) {
    starts <- seq.Date(as.Date(start), as.Date(end), by = sprintf("%d days", chunk_days))
    tibble(
        chunk_start = starts,
        chunk_end = pmin(starts + chunk_days - 1, as.Date(end))
    )
}

rename_first_available <- function(df, new_name, candidates) {
    if (new_name %in% names(df)) {
        return(df)
    }
    old_name <- intersect(candidates, names(df))[1]
    if (is.na(old_name)) {
        return(df)
    }
    rename(df, !!new_name := all_of(old_name))
}

canonicalize_indicator <- function(df, source, indicator) {
    if (nrow(df) == 0 || ncol(df) == 0) {
        return(tibble(
            geo_value = character(),
            time_value = as.Date(character()),
            value = double()
        ))
    }

    df <- df %>%
        rename_first_available("geo_value", c("geo", "geography", "location", "state")) %>%
        rename_first_available("time_value", c("reference_time", "time", "date")) %>%
        rename_first_available("value", c("signal_value", "estimate", "rate", "percent", "pct"))

    missing_cols <- setdiff(c("geo_value", "time_value", "value"), names(df))
    if (length(missing_cols) > 0) {
        stop(glue(
            "Could not normalize {source}/{indicator}; missing columns: {paste(missing_cols, collapse = ', ')}"
        ))
    }

    df %>%
        transmute(
            geo_value = as.character(geo_value),
            time_value = as.Date(time_value),
            value = as.numeric(value)
        )
}

load_indicator <- function(source, indicator, csv = NULL) {
    if (!is.null(csv) && !is.na(csv) && nzchar(csv)) {
        return(
            read.csv(csv) %>%
                canonicalize_indicator(source, indicator) %>%
                filter(
                    time_value >= as.Date(params$start_day),
                    time_value <= as.Date(params$end_day)
                )
        )
    }

    use_new_api <- FALSE
    if (source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(source)
                indicator %in% meta_res[[source]]$signals
            },
            error = function(e) FALSE
        )
        if (meta_available) use_new_api <- TRUE
    }

    if (use_new_api) {
        out <- epidata_snapshot(
            source = source,
            signals = indicator,
            geo_type = params$geo_type
        ) %>%
            rename(any_of(c(time_value = "reference_time", version = "report_time")))

        if ("fill_method" %in% colnames(out)) {
            out <- out %>%
                filter(fill_method == "source" | !any(fill_method == "source"))
        }
        if ("version" %in% colnames(out)) {
            out <- out %>%
                group_by(geo_value, time_value) %>%
                slice_max(version, n = 1, with_ties = FALSE) %>%
                ungroup()
        }

        out %>%
            canonicalize_indicator(source, indicator) %>%
            filter(
                time_value >= as.Date(params$start_day),
                time_value <= as.Date(params$end_day)
            )
    } else {
        date_chunks(params$start_day, params$end_day) %>%
            pmap_dfr(function(chunk_start, chunk_end) {
                pub_covidcast(
                    source,
                    indicator,
                    params$geo_type,
                    params$time_type,
                    time_values = time_range(chunk_start, chunk_end),
                    as_of = Sys.Date(),
                    fetch_args = fetch_args_list(
                        timeout_seconds = params$api_timeout_seconds,
                        return_empty = TRUE
                    )
                ) %>%
                    canonicalize_indicator(source, indicator)
            }) %>%
            distinct(geo_value, time_value, .keep_all = TRUE)
    }
}

cache_path <- file.path(getwd(), "epidatr_cache")
if (!dir.exists(cache_path)) dir.create(cache_path, recursive = TRUE)
set_cache(cache_dir = cache_path)
```

This notebook is the multi-reference version of the standard indicator comparison report. It evaluates one **candidate indicator** against multiple **reference indicators** in the same render. Every diagnostic is computed separately for each candidate/reference pair; reference indicators are not combined into a composite score.

```{r}
#| label: print-config
#| echo: false
reference_csvs <- params$reference_csvs
if (is.null(reference_csvs)) {
    reference_csvs <- rep(NA_character_, length(params$reference_sources))
}

reference_specs <- tibble(
    reference_source = params$reference_sources,
    reference_indicator = params$reference_indicators,
    reference_name = params$reference_names,
    reference_csv = reference_csvs
) %>%
    mutate(
        reference_id = make.unique(paste(reference_source, reference_indicator, sep = "::")),
        reference_label = wrap_reference(reference_name)
    )

if (nrow(reference_specs) == 0) {
    stop("At least one reference indicator is required.")
}
if (length(unique(c(
    length(params$reference_sources),
    length(params$reference_indicators),
    length(params$reference_names),
    length(reference_csvs)
))) != 1) {
    stop("reference_sources, reference_indicators, reference_names, and reference_csvs must have matching lengths.")
}

config_tbl <- bind_rows(
    tibble(
        Role = "Candidate",
        Source = params$candidate_source,
        Indicator = params$candidate_indicator,
        Label = params$candidate_name
    ),
    reference_specs %>%
        transmute(
            Role = "Reference",
            Source = reference_source,
            Indicator = reference_indicator,
            Label = reference_name
        )
)

DT::datatable(config_tbl, caption = sprintf(
    "Geo-level: %s | Period: %s to %s",
    params$geo_type, params$start_day, params$end_day
), rownames = FALSE, options = list(scrollX = TRUE, dom = "t"))
```

---

## Data Loading {#data}

The report loads the candidate once and each reference once, then creates candidate/reference panels by joining on location and time. All diagnostics below use the overlapping rows for each candidate/reference pair. This keeps the report equivalent to running several one-to-one comparisons, but avoids repeated downloads, repeated candidate storage, and repeated renders.

```{r}
#| label: load-data
#| echo: false
#| results: asis
df_candidate <- load_indicator(
    source = params$candidate_source,
    indicator = params$candidate_indicator,
    csv = params$candidate_csv
)

df_references <- pmap_dfr(
    reference_specs,
    function(reference_source, reference_indicator, reference_name, reference_csv, reference_id, reference_label) {
        load_indicator(reference_source, reference_indicator, reference_csv) %>%
            mutate(
                reference_id = reference_id,
                reference_source = reference_source,
                reference_indicator = reference_indicator,
                reference_name = reference_name,
                reference_label = reference_label
            )
    }
)

cat(sprintf(
    "- **Candidate**: %s rows across %s locations  \n",
    label_comma()(nrow(df_candidate)),
    label_comma()(n_distinct(df_candidate$geo_value))
))

reference_load_tbl <- df_references %>%
    group_by(reference_id) %>%
    summarize(
        rows = n(),
        locations = n_distinct(geo_value),
        first_date = min(time_value, na.rm = TRUE),
        last_date = max(time_value, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    right_join(reference_specs, by = "reference_id") %>%
    mutate(
        rows = replace_na(rows, 0L),
        locations = replace_na(locations, 0L),
        first_date = if_else(is.infinite(as.numeric(first_date)), as.Date(NA), first_date),
        last_date = if_else(is.infinite(as.numeric(last_date)), as.Date(NA), last_date)
    ) %>%
    transmute(
        reference_name,
        reference_source,
        reference_indicator,
        rows,
        locations,
        first_date,
        last_date
    )

DT::datatable(
    reference_load_tbl,
    caption = "Loaded reference indicators",
    rownames = FALSE,
    options = list(scrollX = TRUE, pageLength = 10)
)
```

```{r}
#| label: combine-data
#| echo: false
#| results: asis
env_time_type <- attributes(df_candidate)$metadata$time_type
if (is.null(env_time_type) || length(env_time_type) != 1 || is.na(env_time_type) || !is.character(env_time_type)) {
    env_time_type <- params$time_type
}

if (env_time_type == "week") {
    df_candidate <- df_candidate %>%
        mutate(time_value = round_to_sunday(time_value)) %>%
        group_by(geo_value, time_value) %>%
        summarize(value = mean(value, na.rm = TRUE), .groups = "drop") %>%
        as_epi_df()

    df_references <- df_references %>%
        mutate(time_value = round_to_sunday(time_value)) %>%
        group_by(reference_id, reference_source, reference_indicator, reference_name, geo_value, time_value) %>%
        summarize(value = mean(value, na.rm = TRUE), .groups = "drop")

    cat("- Aligned candidate and reference indicators to Sunday-start weeks  \n")
}

grid_start <- as.Date(params$start_day)
grid_end <- as.Date(params$end_day)
if (env_time_type == "week") {
    grid_start <- round_to_sunday(grid_start)
    grid_end <- round_to_sunday(grid_end)
}

df_pair <- df_references %>%
    select(reference_id, reference_source, reference_indicator, reference_name, reference_label, geo_value, time_value, reference = value) %>%
    left_join(
        df_candidate %>%
            select(geo_value, time_value, candidate = value),
        by = c("geo_value", "time_value")
    ) %>%
    group_by(reference_id, reference_source, reference_indicator, reference_name, geo_value) %>%
    complete(
        time_value = seq.Date(
            from = grid_start,
            to = grid_end,
            by = env_time_type
        )
    ) %>%
    ungroup()

df_inter <- df_pair %>%
    filter(!is.na(candidate), !is.na(reference))

intersection_tbl <- df_pair %>%
    group_by(reference_id, reference_name) %>%
    summarize(
        full_rows = n(),
        intersected_rows = sum(!is.na(candidate) & !is.na(reference)),
        intersected_locations = n_distinct(geo_value[!is.na(candidate) & !is.na(reference)]),
        intersected_dates = n_distinct(time_value[!is.na(candidate) & !is.na(reference)]),
        intersection_pct = 100 * intersected_rows / full_rows,
        .groups = "drop"
    ) %>%
    right_join(reference_specs %>% select(reference_id, reference_name), by = c("reference_id", "reference_name")) %>%
    mutate(
        full_rows = replace_na(full_rows, 0L),
        intersected_rows = replace_na(intersected_rows, 0L),
        intersected_locations = replace_na(intersected_locations, 0L),
        intersected_dates = replace_na(intersected_dates, 0L),
        intersection_pct = replace_na(intersection_pct, 0)
    )

DT::datatable(
    intersection_tbl %>%
        mutate(intersection_pct = sprintf("%.1f%%", intersection_pct)),
    caption = "Candidate/reference overlap",
    rownames = FALSE,
    options = list(scrollX = TRUE, pageLength = 10)
)

time_unit <- switch(env_time_type %||% "unknown",
    week = "weeks",
    month = "months",
    year = "years",
    "days"
)
lag_unit_label <- sprintf("Lag (%s)", time_unit)

lag_values <- switch(env_time_type %||% "unknown",
    week = -2:2,
    month = -2:2,
    year = -1:1,
    c(-14, -7, 0, 7, 14)
)

lags_sweep <- switch(env_time_type %||% "unknown",
    week = -5:5,
    month = -5:5,
    year = -2:2,
    -21:21
)

ref_cols <- reference_palette(sort(unique(df_inter$reference_label)))
```

---

```{r}
#| label: core-metrics
#| include: false
cor_by_time <- df_inter %>%
    group_by(reference_id, reference_name, reference_label, time_value) %>%
    summarize(
        n_locations = n_distinct(geo_value),
        cor = safe_spearman(candidate, reference),
        .groups = "drop"
    )

mean_geo_cor <- cor_by_time %>%
    group_by(reference_id, reference_name, reference_label) %>%
    summarize(
        mean_cor = mean(cor, na.rm = TRUE),
        median_cor = median(cor, na.rm = TRUE),
        p10_cor = quantile(cor, 0.10, na.rm = TRUE),
        p25_cor = quantile(cor, 0.25, na.rm = TRUE),
        p75_cor = quantile(cor, 0.75, na.rm = TRUE),
        p90_cor = quantile(cor, 0.90, na.rm = TRUE),
        dates = sum(!is.na(cor)),
        .groups = "drop"
    )

cor_by_geo <- df_inter %>%
    group_by(reference_id, reference_name, reference_label, geo_value) %>%
    summarize(
        n_dates = n_distinct(time_value),
        cor = safe_spearman(candidate, reference),
        .groups = "drop"
    )

temporal_summary <- cor_by_geo %>%
    group_by(reference_id, reference_name, reference_label) %>%
    summarize(
        locations = sum(!is.na(cor)),
        median_cor = median(cor, na.rm = TRUE),
        mean_cor = mean(cor, na.rm = TRUE),
        p10_cor = quantile(cor, 0.10, na.rm = TRUE),
        p25_cor = quantile(cor, 0.25, na.rm = TRUE),
        p75_cor = quantile(cor, 0.75, na.rm = TRUE),
        p90_cor = quantile(cor, 0.90, na.rm = TRUE),
        pct_abs_gt_05 = 100 * mean(abs(cor) > 0.5, na.rm = TRUE),
        .groups = "drop"
    )

candidate_lagged <- df_candidate %>%
    select(geo_value, candidate_time_value = time_value, candidate = value) %>%
    tidyr::crossing(lag = lags_sweep) %>%
    mutate(time_value = add_lag_units(candidate_time_value, lag, env_time_type)) %>%
    select(geo_value, time_value, lag, candidate)

df_lagged <- df_references %>%
    select(reference_id, reference_source, reference_indicator, reference_name, reference_label, geo_value, time_value, reference = value) %>%
    inner_join(candidate_lagged, by = c("geo_value", "time_value")) %>%
    filter(!is.na(candidate), !is.na(reference))

cor_sweep_geo <- df_lagged %>%
    group_by(reference_id, reference_name, reference_label, lag, geo_value) %>%
    summarize(
        n_dates = n_distinct(time_value),
        cor = safe_spearman(candidate, reference),
        .groups = "drop"
    )

lag_summary <- cor_sweep_geo %>%
    group_by(reference_id, reference_name, reference_label, lag) %>%
    summarize(
        mean_cor = mean(cor, na.rm = TRUE),
        median_cor = median(cor, na.rm = TRUE),
        q10 = quantile(cor, 0.10, na.rm = TRUE),
        q25 = quantile(cor, 0.25, na.rm = TRUE),
        q75 = quantile(cor, 0.75, na.rm = TRUE),
        q90 = quantile(cor, 0.90, na.rm = TRUE),
        locations = sum(!is.na(cor)),
        .groups = "drop"
    )

optimal_lag <- lag_summary %>%
    group_by(reference_id, reference_name, reference_label) %>%
    slice_max(mean_cor, n = 1, with_ties = FALSE) %>%
    ungroup()
```

---

## Exploratory Data Analysis {#eda}

Before comparing correlations, this section checks whether each candidate/reference pair has enough usable overlap and whether the indicators have broadly comparable behavior. These views mirror the original one-to-one comparison report, but the same calculations are grouped by reference indicator.

### Time Series Overview

This plot compares the candidate and references on a common scale. For each indicator and location, values are standardized as

$$
Z_{g,t}=\frac{X_{g,t}-\bar{X}_g}{s_g},
$$

where $X_{g,t}$ is the raw value at location $g$ and time $t$, $\bar{X}_g$ is that location's mean, and $s_g$ is that location's standard deviation. The plotted line is the median standardized value across locations at each date.

```{r}
#| label: normalized-series
#| echo: false
#| fig-height: 7.8
df_candidate_norm <- df_candidate %>%
    filter(geo_value %in% unique(df_inter$geo_value)) %>%
    mutate(indicator_label = wrap_reference(params$candidate_name)) %>%
    group_by(geo_value, indicator_label) %>%
    mutate(value_norm = (value - mean(value, na.rm = TRUE)) / sd(value, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(time_value, indicator_label) %>%
    summarize(median_norm = median(value_norm, na.rm = TRUE), .groups = "drop")

df_reference_norm <- df_references %>%
    filter(reference_id %in% unique(df_inter$reference_id)) %>%
    group_by(reference_id, reference_label, geo_value) %>%
    mutate(value_norm = (value - mean(value, na.rm = TRUE)) / sd(value, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(time_value, indicator_label = reference_label) %>%
    summarize(median_norm = median(value_norm, na.rm = TRUE), .groups = "drop")

trend_long <- bind_rows(df_candidate_norm, df_reference_norm) %>%
    filter(is.finite(median_norm))

trend_levels <- unique(trend_long$indicator_label)
trend_cols <- setNames(c("#172431", rep(dis_pal_list, length.out = length(trend_levels) - 1)), trend_levels)

ggplot(trend_long, aes(x = time_value, y = median_norm, color = indicator_label)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray70", linewidth = 0.35) +
    geom_line(linewidth = 0.85, alpha = 0.9) +
    scale_x_date(breaks = breaks_pretty(n = 7), labels = label_date_short()) +
    scale_y_continuous(breaks = breaks_pretty()) +
    scale_color_manual(values = trend_cols, labels = wrap_legend_label) +
    labs(
        title = "Median normalized trend across locations",
        subtitle = "Candidate and all references on a common standardized scale",
        x = "Date", y = "Median normalized value", color = "Indicator"
    ) +
    guides(color = guide_legend(ncol = 1, byrow = TRUE, override.aes = list(linewidth = 1.4))) +
    safe_bottom_legend(text_size = 10) +
    theme(axis.text.x = element_text(angle = 35, hjust = 1))
```

### Quantile Trends

This plot displays the cross-location distribution of the candidate and each reference over time. For each candidate/reference pair, the candidate quantiles are recomputed on that pair's overlap so that the comparison uses the same locations and dates as the later pairwise correlations.

```{r}
#| label: eda-quantiles
#| echo: false
#| fig-height: 9.5
df_quantiles_multi <- df_inter %>%
    pivot_longer(
        cols = c(candidate, reference),
        names_to = "indicator_role",
        values_to = "value"
    ) %>%
    mutate(
        indicator_role = recode(indicator_role, candidate = "Candidate", reference = "Reference"),
        indicator_label = if_else(indicator_role == "Candidate", params$candidate_name, reference_name)
    ) %>%
    group_by(reference_label, indicator_role, indicator_label, time_value) %>%
    summarize(
        q05 = quantile(value, 0.05, na.rm = TRUE),
        q25 = quantile(value, 0.25, na.rm = TRUE),
        q50 = quantile(value, 0.50, na.rm = TRUE),
        q75 = quantile(value, 0.75, na.rm = TRUE),
        q95 = quantile(value, 0.95, na.rm = TRUE),
        locations = sum(!is.na(value)),
        .groups = "drop"
    ) %>%
    filter(is.finite(q50))

ggplot(df_quantiles_multi, aes(x = time_value)) +
    geom_ribbon(aes(ymin = q05, ymax = q95), fill = dis_pal_list[2], alpha = 0.12) +
    geom_ribbon(aes(ymin = q25, ymax = q75), fill = dis_pal_list[2], alpha = 0.28) +
    geom_line(aes(y = q50), color = dis_pal_list[2], linewidth = 0.75) +
    facet_grid(reference_label ~ indicator_role, scales = "free_y") +
    scale_x_date(breaks = breaks_pretty(n = 5), labels = label_date_short()) +
    scale_y_continuous(labels = label_comma(), breaks = breaks_pretty(n = 4)) +
    labs(
        title = "Quantile trends by candidate/reference comparison",
        subtitle = "Line is median across locations; ribbons show 25th-75th and 5th-95th percentiles",
        x = "Date", y = "Value"
    ) +
    theme(
        axis.text.x = element_text(angle = 35, hjust = 1),
        strip.text = element_text(size = 11, face = "bold")
    )
```

### Coverage and Missingness

This plot shows how many locations are available for the candidate, for each reference, and for their pairwise overlap at each date. The "Both" line is the effective sample size used for date-specific spatial correlations.

```{r}
#| label: coverage-over-time
#| echo: false
#| fig-height: 6
coverage <- df_pair %>%
    group_by(reference_name, time_value) %>%
    summarize(
        candidate_available = sum(!is.na(candidate)),
        reference_available = sum(!is.na(reference)),
        both_available = sum(!is.na(candidate) & !is.na(reference)),
        .groups = "drop"
    ) %>%
    pivot_longer(
        cols = c(candidate_available, reference_available, both_available),
        names_to = "availability",
        values_to = "locations"
    ) %>%
    mutate(availability = recode(
        availability,
        candidate_available = "Candidate",
        reference_available = "Reference",
        both_available = "Both"
    ))

coverage_refs <- unique(coverage$reference_name)
ggplot(coverage, aes(x = time_value, y = locations, color = availability)) +
    geom_step(direction = "mid", linewidth = 0.5) +
    facet_wrap(~wrap_reference(reference_name), ncol = 1) +
    scale_x_date(breaks = breaks_pretty(), labels = label_date_short()) +
    scale_y_continuous(labels = label_comma()) +
    scale_color_manual(values = c(Candidate = dis_pal_list[2], Reference = dis_pal_list[1], Both = dis_pal_list[4])) +
    labs(
        title = "Data availability over time",
        x = "Date", y = "Locations", color = ""
    ) +
    theme(legend.position = "bottom")
```

### Summary Statistics and Missingness

This section summarizes availability and basic values by location for each candidate/reference pair.

```{r}
#| label: eda-summary-missingness
#| echo: false
#| fig-height: 5.5
location_summary <- df_pair %>%
    group_by(reference_id, reference_name, reference_label, geo_value) %>%
    summarize(
        periods = n(),
        candidate_mean = mean(candidate, na.rm = TRUE),
        candidate_missing_pct = 100 * mean(is.na(candidate)),
        reference_mean = mean(reference, na.rm = TRUE),
        reference_missing_pct = 100 * mean(is.na(reference)),
        pair_missing_pct = 100 * mean(is.na(candidate) | is.na(reference)),
        .groups = "drop"
    ) %>%
    mutate(geo_name = geo_to_name(geo_value))

missingness_summary <- location_summary %>%
    group_by(reference_name, reference_label) %>%
    summarize(
        locations = n_distinct(geo_value),
        median_candidate_missing_pct = median(candidate_missing_pct, na.rm = TRUE),
        median_reference_missing_pct = median(reference_missing_pct, na.rm = TRUE),
        median_pair_missing_pct = median(pair_missing_pct, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    right_join(reference_specs %>% select(reference_name, reference_label), by = c("reference_name", "reference_label")) %>%
    mutate(
        locations = replace_na(locations, 0L),
        across(ends_with("_pct"), ~ replace_na(.x, 100))
    )

missingness_long <- missingness_summary %>%
    pivot_longer(
        cols = ends_with("_pct"),
        names_to = "metric",
        values_to = "missing_pct"
    ) %>%
    mutate(metric = recode(
        metric,
        median_candidate_missing_pct = "Candidate",
        median_reference_missing_pct = "Reference",
        median_pair_missing_pct = "Either missing"
    ))

ggplot(missingness_long, aes(x = metric, y = reference_label, fill = missing_pct)) +
    geom_tile(color = "white", linewidth = 0.35) +
    geom_text(aes(label = sprintf("%.0f%%", missing_pct)), color = "gray15", size = 3.3) +
    scale_fill_gradient(low = "white", high = dis_pal_list[3], limits = c(0, 100), labels = label_percent(scale = 1)) +
    labs(
        title = "Median missingness by reference comparison",
        subtitle = "Pair missingness is the fraction of periods where either candidate or reference is missing",
        x = NULL, y = "Reference indicator", fill = "Median missing"
    ) +
    theme(axis.text.y = element_text(size = 11), legend.position = "right")
```

```{r}
#| label: eda-summary-table
#| echo: false
#| results: asis
summary_table <- location_summary %>%
    arrange(desc(pair_missing_pct), reference_name, geo_value) %>%
    transmute(
        Reference = reference_name,
        `Geo value` = geo_value,
        `Geo name` = geo_name,
        Periods = periods,
        `Candidate mean` = candidate_mean,
        `Candidate missing` = sprintf("%.1f%%", candidate_missing_pct),
        `Reference mean` = reference_mean,
        `Reference missing` = sprintf("%.1f%%", reference_missing_pct),
        `Either missing` = sprintf("%.1f%%", pair_missing_pct)
    )

DT::datatable(
    summary_table,
    caption = "Summary statistics by location and reference comparison.",
    rownames = FALSE,
    options = list(scrollX = TRUE, scrollY = "420px", scrollCollapse = TRUE, pageLength = 15)
) %>%
    DT::formatRound(columns = c("Candidate mean", "Reference mean"), digits = 2)
```

### Noise Assessment

Coefficient of variation (CV) is `sd(value) / abs(mean(value))` within each location. Lower CV means less variation relative to the location-specific mean.

```{r}
#| label: eda-noise
#| echo: false
noise_stats <- df_inter %>%
    group_by(reference_id, reference_name, reference_label, geo_value) %>%
    summarize(
        candidate_cv = sd(candidate, na.rm = TRUE) / abs(mean(candidate, na.rm = TRUE)),
        reference_cv = sd(reference, na.rm = TRUE) / abs(mean(reference, na.rm = TRUE)),
        .groups = "drop"
    ) %>%
    filter(if_all(c(candidate_cv, reference_cv), is.finite)) %>%
    mutate(
        cv_diff = candidate_cv - reference_cv,
        candidate_smoother = candidate_cv < reference_cv
    )

cv_upper <- quantile(c(noise_stats$candidate_cv, noise_stats$reference_cv), 0.99, na.rm = TRUE)
noise_stats_plot <- noise_stats %>%
    filter(candidate_cv <= cv_upper, reference_cv <= cv_upper)

noise_summary <- noise_stats %>%
    group_by(reference_name) %>%
    summarize(
        `Locations compared` = n(),
        `Median candidate CV` = median(candidate_cv, na.rm = TRUE),
        `Median reference CV` = median(reference_cv, na.rm = TRUE),
        `Median CV difference` = median(cv_diff, na.rm = TRUE),
        `% candidate smoother` = 100 * mean(candidate_smoother, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    right_join(reference_specs %>% select(reference_name), by = "reference_name") %>%
    mutate(
        `Locations compared` = replace_na(`Locations compared`, 0L),
        across(where(is.numeric), ~ if_else(is.nan(.x), NA_real_, .x))
    )

DT::datatable(
    noise_summary,
    caption = "Comparative CV summary. Negative median CV difference means the candidate is typically smoother than that reference.",
    rownames = FALSE,
    options = list(scrollX = TRUE, paging = FALSE, dom = "t")
) %>%
    DT::formatRound(columns = c("Median candidate CV", "Median reference CV", "Median CV difference", "% candidate smoother"), digits = 2)
```

```{r}
#| label: eda-noise-plot
#| echo: false
#| fig-height: 8.6
p_cv_diff <- ggplot(noise_stats_plot, aes(x = reference_label, y = cv_diff, fill = reference_label)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_violin(alpha = 0.35, color = NA, trim = TRUE) +
    geom_boxplot(width = 0.18, outlier.alpha = 0.25, color = "gray25") +
    coord_flip() +
    scale_fill_manual(values = ref_cols) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    guides(fill = "none") +
    labs(
        title = "Candidate/reference CV difference by location",
        subtitle = "Negative values mean the candidate is smoother than the reference",
        x = NULL, y = "Candidate CV - reference CV"
    )

p_cv_pair <- ggplot(noise_stats_plot, aes(x = reference_cv, y = candidate_cv, color = reference_label)) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey50") +
    geom_point(alpha = 0.55, size = 1.7) +
    scale_x_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    scale_y_continuous(labels = label_percent(), breaks = breaks_pretty()) +
    scale_color_manual(values = ref_cols, labels = wrap_legend_label) +
    labs(
        title = "Paired CV comparison",
        subtitle = "Points below the dashed line indicate locations where the candidate is smoother",
        x = "Reference CV", y = "Candidate CV", color = "Reference indicator"
    ) +
    safe_bottom_legend(text_size = 10) +
    guides(color = guide_legend(ncol = 1, byrow = TRUE, override.aes = list(size = 3, alpha = 1)))

p_cv_diff / p_cv_pair + plot_layout(heights = c(1, 1.2))
```

## Correlation Analysis {#correlation}

All correlation metrics are computed on candidate/reference intersections: rows where both the candidate and the given reference are non-missing. The report repeats the original one-to-one correlation diagnostics for each reference indicator in parallel.

### Correlation over Time

This section asks whether the candidate ranks locations similarly to each reference at each date. Here, $g$ indexes locations, $t$ indexes time, $k$ indexes reference indicators, $C_{g,t}$ is the candidate value, and $R^{(k)}_{g,t}$ is reference $k$.

$\operatorname{cor}_S(x,y)$ is Spearman rank correlation after dropping missing pairs. In this report, "agreement" means numeric association, not causality or ground truth.

For each individual reference $k$ and time $t$, the pairwise spatial correlation is computed across locations:

$$
\rho^{(k)}_{\text{space},t}
= \operatorname{cor}_S
\left(
  \{C_{g,t}\}_{g},
  \{R^{(k)}_{g,t}\}_{g}
\right).
$$

```{r}
#| label: cor-geographic-over-time
#| echo: false
#| fig-height: 7.8
ref_cols <- reference_palette(sort(unique(cor_by_time$reference_label)))

ggplot(cor_by_time, aes(x = time_value, y = cor, color = reference_label)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = c(0.3, 0.6, 0.8), linetype = "dotted", color = "gray85", linewidth = 0.3) +
    geom_line(alpha = 0.35, linewidth = 0.5) +
    geom_smooth(aes(fill = reference_label), method = "loess", span = 0.2, se = FALSE, linewidth = 1.1) +
    scale_x_date(breaks = breaks_pretty(), labels = label_date_short()) +
    scale_y_continuous(limits = c(-1, 1), breaks = breaks_pretty()) +
    scale_color_manual(values = ref_cols, labels = wrap_legend_label) +
    scale_fill_manual(values = ref_cols, labels = wrap_legend_label) +
    labs(
        title = "Pairwise geographic rank correlation over time",
        subtitle = "Faint lines show date-specific correlations; thicker smooth lines show LOESS trends for each reference",
        x = "Date", y = "Spearman ρ across locations",
        color = "Reference indicator", fill = "Reference indicator"
    ) +
    guides(
        color = guide_legend(ncol = 1, byrow = TRUE, override.aes = list(linewidth = 1.5, alpha = 1)),
        fill = "none"
    ) +
    safe_bottom_legend(text_size = 10) +
    theme(axis.text.x = element_text(angle = 35, hjust = 1))
```

---

### Distribution of Temporal Correlations

This section asks whether the candidate and references move together over time within each location at zero lag. Each correlation is computed across dates while holding one location fixed, then the distribution across locations is shown for each reference.

For each individual reference $k$ and location $g$, the pairwise temporal correlation is:

$$
\rho^{(k)}_{\text{time},g}
= \operatorname{cor}_S
\left(
  \{C_{g,t}\}_{t},
  \{R^{(k)}_{g,t}\}_{t}
\right).
$$

```{r}
#| label: cor-temporal-dist
#| echo: false
#| fig-height: 7.5
temporal_medians <- cor_by_geo %>%
    group_by(reference_id, reference_name, reference_label) %>%
    summarize(median_cor = median(cor, na.rm = TRUE), .groups = "drop")

ggplot(cor_by_geo %>% filter(is.finite(cor)), aes(x = cor, fill = reference_label)) +
    geom_histogram(bins = 24, color = "white", alpha = 0.75) +
    geom_vline(
        data = temporal_medians,
        aes(xintercept = median_cor),
        linetype = "dashed",
        color = "gray20",
        linewidth = 0.8
    ) +
    facet_wrap(~reference_label, ncol = 1) +
    scale_x_continuous(limits = c(-1, 1), breaks = breaks_pretty()) +
    scale_fill_manual(values = ref_cols) +
    guides(fill = "none") +
    labs(
        title = "Distribution of intra-location temporal correlations",
        subtitle = "Each histogram shows zero-lag Spearman correlations across locations; dashed line is the reference-specific median",
        x = expression(Spearman ~ rho),
        y = "Locations"
    ) +
    theme(strip.text = element_text(size = 12, face = "bold"))
```

### Lagged Correlations

This section scans candidate lead/lag values. The lag $\ell$ is measured in `r time_unit`. Positive lag means earlier candidate values are compared with current reference values, so the candidate leads the reference. Negative lag shifts the candidate later.

For lag $\ell$, the shifted candidate is compared with the reference:

$$
C_{g,t-\ell}
\quad \text{vs.} \quad
R^{(k)}_{g,t}.
$$

#### Per-location temporal correlation distributions, at different lags

```{r}
#| label: cor-temporal-by-location
#| echo: false
#| fig-height: 9.5
cor_lagged_geo_selected <- cor_sweep_geo %>%
    filter(lag %in% lag_values, is.finite(cor)) %>%
    mutate(lag = factor(lag, levels = lag_values))

if (nrow(cor_lagged_geo_selected) == 0) {
    cat("Pairwise temporal correlation distributions by lag are not available because the selected lags have insufficient overlap.\n\n")
} else {
    ggplot(cor_lagged_geo_selected, aes(x = cor, color = lag, fill = lag)) +
        geom_vline(xintercept = 0, linetype = "dashed", color = "gray65") +
        geom_density(alpha = 0.14, linewidth = 0.85, adjust = 1.1, na.rm = TRUE) +
        facet_wrap(~reference_label, ncol = 1) +
        scale_x_continuous(limits = c(-1, 1), breaks = breaks_pretty()) +
        scale_color_manual(values = lag_palette(length(lag_values))) +
        scale_fill_manual(values = lag_palette(length(lag_values))) +
        labs(
            title = "Pairwise per-location temporal correlation distributions at selected lags",
            subtitle = "Each curve summarizes location-level temporal Spearman ρ for one candidate lag; positive lag means candidate leads reference",
            x = "Per-location temporal Spearman ρ",
            y = "Density",
            color = lag_unit_label,
            fill = lag_unit_label
        ) +
        guides(
            color = guide_legend(nrow = 1, byrow = TRUE),
            fill = guide_legend(nrow = 1, byrow = TRUE)
        ) +
        safe_bottom_legend(text_size = 10) +
        theme(strip.text = element_text(size = 12, face = "bold"))
}
```

#### Empirical Uncertainty

These intervals summarize variation in the pairwise correlation distributions. Spatial intervals vary across dates; temporal intervals vary across locations. They are empirical summaries, not model-based confidence intervals.

```{r}
#| label: empirical-uncertainty
#| echo: false
#| fig-height: 7
uncertainty_tbl <- bind_rows(
    mean_geo_cor %>%
        transmute(
            reference_id, reference_name, reference_label,
            metric = "Spatial ρ across dates",
            median = median_cor,
            p10 = p10_cor,
            p25 = p25_cor,
            p75 = p75_cor,
            p90 = p90_cor,
            units = dates
        ),
    temporal_summary %>%
        transmute(
            reference_id, reference_name, reference_label,
            metric = "Temporal ρ across locations",
            median = median_cor,
            p10 = p10_cor,
            p25 = p25_cor,
            p75 = p75_cor,
            p90 = p90_cor,
            units = locations
        )
) %>%
    filter(is.finite(median)) %>%
    mutate(metric = factor(metric, levels = c("Spatial ρ across dates", "Temporal ρ across locations")))

ggplot(uncertainty_tbl, aes(y = reference_label, x = median, color = reference_label)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray65") +
    geom_linerange(aes(xmin = p10, xmax = p90), alpha = 0.25, linewidth = 2) +
    geom_linerange(aes(xmin = p25, xmax = p75), alpha = 0.75, linewidth = 3.5) +
    geom_point(size = 2.8, color = "gray15") +
    facet_wrap(~metric, ncol = 1) +
    scale_x_continuous(limits = c(-1, 1), breaks = breaks_pretty()) +
    scale_color_manual(values = ref_cols) +
    guides(color = "none") +
    labs(
        title = "Empirical uncertainty in pairwise agreement metrics",
        subtitle = "Thick intervals show 25th-75th percentiles; thin intervals show 10th-90th percentiles",
        x = "Spearman ρ", y = "Reference indicator"
    ) +
    theme(axis.text.y = element_text(size = 11))
```

---

#### Systematic Lag Analysis

The pairwise lag response curve summarizes within-location temporal correlations by reference and lag. The marked lag is the value with the largest mean temporal correlation across locations for that reference.

```{r}
#| label: lag-response-curve
#| echo: false
#| fig-height: 7.8
ggplot(lag_summary, aes(x = lag, y = median_cor, color = reference_label, fill = reference_label)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
    geom_vline(xintercept = 0, linetype = "dotted", color = "gray70") +
    geom_ribbon(aes(ymin = q25, ymax = q75), alpha = 0.12, color = NA) +
    geom_line(linewidth = 0.9) +
    geom_point(
        data = optimal_lag,
        aes(x = lag, y = median_cor),
        inherit.aes = FALSE,
        shape = 21, fill = "white", color = "black", size = 3, stroke = 0.8
    ) +
    scale_color_manual(values = ref_cols, labels = wrap_legend_label) +
    scale_fill_manual(values = ref_cols, labels = wrap_legend_label) +
    scale_x_continuous(breaks = breaks_pretty(n = 12)) +
    scale_y_continuous(limits = c(-1, 1), breaks = breaks_pretty()) +
    labs(
        title = "Pairwise lag response curves",
        subtitle = "Line is median within-location ρ; ribbon is 25th-75th percentile across locations",
        x = lag_unit_label, y = "Spearman ρ over time",
        color = "Reference indicator", fill = "Reference indicator"
    ) +
    guides(
        color = guide_legend(ncol = 1, byrow = TRUE, override.aes = list(linewidth = 1.5)),
        fill = "none"
    ) +
    safe_bottom_legend(text_size = 10)
```

```{r}
#| label: lag-sweep
#| echo: false
#| fig-height: 5.5
ggplot(lag_summary, aes(x = lag, y = reference_label, fill = mean_cor)) +
    geom_tile(color = "white", linewidth = 0.25) +
    geom_point(
        data = optimal_lag,
        aes(x = lag, y = reference_label),
        inherit.aes = FALSE,
        shape = 21, size = 3.2, stroke = 1,
        fill = "white", color = "black"
    ) +
    scale_x_continuous(breaks = breaks_pretty(n = 12)) +
    scale_fill_gradient2(
        low = dis_pal_list[3],
        mid = "white",
        high = dis_pal_list[2],
        midpoint = 0,
        limits = c(-1, 1),
        breaks = breaks_pretty(n = 5),
        na.value = "gray90"
    ) +
    labs(
        title = "Pairwise lagged temporal agreement matrix",
        subtitle = "Cell color is mean within-location Spearman ρ; white dot marks each reference's best lag",
        x = lag_unit_label, y = "Reference indicator", fill = "Mean ρ"
    ) +
    theme(
        legend.position = "right",
        axis.text.y = element_text(size = 11)
    )
```

#### Correlation over time, at different `r time_unit`

This view computes correlation across locations at each date after shifting the candidate. It shows whether selected lags remain stable over calendar time, rather than only improving the location-level temporal distribution.

```{r}
#| label: lagged-geographic-time
#| echo: false
#| fig-height: 9.4
cor_lagged_time <- df_lagged %>%
    filter(lag %in% lag_values) %>%
    group_by(reference_id, reference_name, reference_label, lag, time_value) %>%
    summarize(
        n_locations = n_distinct(geo_value),
        cor = safe_spearman(candidate, reference),
        .groups = "drop"
    ) %>%
    mutate(lag = factor(lag, levels = lag_values))

ggplot(cor_lagged_time, aes(x = time_value, y = cor, color = reference_label, linetype = lag)) +
    geom_line(alpha = 0.3, linewidth = 0.45) +
    geom_smooth(aes(group = interaction(reference_label, lag)), method = "loess", span = 0.2, se = FALSE, linewidth = 0.8) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    scale_x_date(breaks = breaks_pretty(), labels = label_date_short()) +
    scale_y_continuous(limits = c(-1, 1), breaks = breaks_pretty()) +
    scale_color_manual(values = ref_cols, labels = wrap_legend_label) +
    labs(
        title = sprintf("Pairwise geographic correlation over time at selected %s", time_unit),
        subtitle = "Color compares reference indicators; line type compares candidate lead/lag",
        x = "Date", y = "Spearman ρ across locations",
        color = "Reference indicator", linetype = lag_unit_label
    ) +
    guides(
        color = guide_legend(ncol = 1, byrow = TRUE, override.aes = list(linewidth = 1.5, alpha = 1)),
        linetype = guide_legend(nrow = 1, byrow = TRUE)
    ) +
    safe_bottom_legend(text_size = 10) +
    theme(axis.text.x = element_text(angle = 35, hjust = 1))
```

---

### Choropleth Map at Optimal Lag

These maps and tables show where temporal agreement is stronger or weaker. For each reference, the map uses that reference's optimal lag from the lead-lag scan above, defined as the lag with the largest mean within-location temporal Spearman correlation across locations.

```{r}
#| label: cor-choropleth
#| echo: false
#| fig-height: 8
cor_optimal_multi <- cor_sweep_geo %>%
    inner_join(
        optimal_lag %>% select(reference_id, optimal_lag = lag),
        by = "reference_id"
    ) %>%
    filter(lag == optimal_lag) %>%
    mutate(geo_name = geo_to_name(geo_value))

looks_like_state <- params$geo_type == "state" &&
    nrow(cor_optimal_multi) > 0 &&
    all(nchar(cor_optimal_multi$geo_value) == 2)
looks_like_county <- params$geo_type == "county" &&
    nrow(cor_optimal_multi) > 0 &&
    all(nchar(cor_optimal_multi$geo_value) == 5)

if (looks_like_state || looks_like_county) {
    map_plots <- cor_optimal_multi %>%
        split(.$reference_label) %>%
        imap(function(dat, label) {
            cor_choro <- dat %>%
                transmute(
                    geo_value,
                    value = cor,
                    time_value = as.Date(params$start_day),
                    issue = as.Date(params$start_day),
                    data_source = "multi_reference_evaluation",
                    signal = "correlation"
                )

            attributes(cor_choro)$metadata$geo_type <- params$geo_type
            class(cor_choro) <- c("covidcast_signal", "data.frame")

            plot(cor_choro,
                range = c(-1, 1),
                choro_col = c(dis_pal_list[3], "white", dis_pal_list[2]),
                title = stringr::str_wrap(sprintf("%s\noptimal lag: %s %s", label, unique(dat$optimal_lag), time_unit), 44)
            )
        })

    wrap_plots(map_plots, ncol = min(2, length(map_plots)))
} else {
    cat("Choropleth maps are available for state or county geographies with valid location codes.\n\n")
}
```

```{r}
#| label: cor-choropleth-table
#| echo: false
#| results: asis
cor_top <- cor_optimal_multi %>%
    arrange(reference_name, desc(cor)) %>%
    group_by(reference_name) %>%
    mutate(rank_within_reference = row_number()) %>%
    ungroup() %>%
    transmute(
        Reference = reference_name,
        `Geo value` = geo_value,
        `Geo name` = geo_name,
        `Optimal lag` = optimal_lag,
        `Spearman ρ` = cor,
        Rank = rank_within_reference
    )

DT::datatable(
    cor_top,
    caption = "Location-level temporal correlation at each pairwise reference's optimal lag.",
    rownames = FALSE,
    options = list(scrollX = TRUE, scrollY = "420px", scrollCollapse = TRUE, pageLength = 15)
) %>%
    DT::formatRound(columns = "Spearman ρ", digits = 3)
```

## Summary {#summary}

This final table collects the main pairwise metrics for quick review. It intentionally does not combine references or produce an automated decision; it reports the same quantities for each candidate/reference pair so reviewers can compare patterns across references.

```{r}
#| label: screening-summary
#| echo: false
screening_tbl <- mean_geo_cor %>%
    full_join(
        temporal_summary,
        by = c("reference_id", "reference_name", "reference_label"),
        suffix = c("_geographic", "_temporal")
    ) %>%
    full_join(
        optimal_lag %>%
            select(reference_id, reference_name, reference_label, optimal_lag = lag, mean_cor_at_optimal_lag = mean_cor, median_cor_at_optimal_lag = median_cor),
        by = c("reference_id", "reference_name", "reference_label")
    ) %>%
    right_join(reference_specs %>% select(reference_id, reference_name), by = c("reference_id", "reference_name")) %>%
    transmute(
        Reference = reference_name,
        `Mean geographic ρ` = if_else(is.na(mean_cor_geographic), "No overlap", sprintf("%.3f", mean_cor_geographic)),
        `Median geographic ρ` = if_else(is.na(median_cor_geographic), "No overlap", sprintf("%.3f", median_cor_geographic)),
        `Mean temporal ρ` = if_else(is.na(mean_cor_temporal), "No overlap", sprintf("%.3f", mean_cor_temporal)),
        `Median temporal ρ` = if_else(is.na(median_cor_temporal), "No overlap", sprintf("%.3f", median_cor_temporal)),
        `% locations |ρ| > 0.5` = if_else(is.na(pct_abs_gt_05), "No overlap", sprintf("%.0f%%", pct_abs_gt_05)),
        `Optimal lag` = if_else(is.na(optimal_lag), "No overlap", as.character(optimal_lag)),
        `Mean ρ at optimal lag` = if_else(is.na(mean_cor_at_optimal_lag), "No overlap", sprintf("%.3f", mean_cor_at_optimal_lag)),
        `Median ρ at optimal lag` = if_else(is.na(median_cor_at_optimal_lag), "No overlap", sprintf("%.3f", median_cor_at_optimal_lag))
    ) %>%
    arrange(desc(suppressWarnings(as.numeric(`Mean ρ at optimal lag`))))

DT::datatable(
    screening_tbl,
    caption = "Multi-reference screening metrics",
    rownames = FALSE,
    options = list(scrollX = TRUE, pageLength = 10)
)
```
