---
title: "Indicator Exploratory Analysis: `r paste(if (!is.null(params$name)) params$name else params$signal, collapse = ', ')`"
author: "Delphi Group"
date: today
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
params:
  # Candidate Indicator
  source:
  signal:
  name:
  input_csv: !expr "NULL"
  # Shared settings
  geo_type:
  time_type:
  start_day:
  end_day:
  # Extra key columns beyond geo_value/time_value (character vector), e.g. "age_group"
  extra_keys: !expr "character(0)"
  # First month of the epidemiological year used by the seasonality panels.
  # 7 = July-June, the usual influenza convention.
  season_start_month: 7
  # Smooth the quantile trends with a moving average. Meant for raw data with a
  # day-of-week effect; set to false for indicators already smoothed at source.
  smooth_quantiles: true
  # Show the cross-sectional geographic outlier table (2.2.1)
  show_geo_outliers: true
  # Show the versioning summary. Off also skips the version-history query, which
  # is the slowest part of the report; use revision_analysis.qmd for the detail.
  show_versioning: true
  # Maximum locations to show in plots/tables
  max_locations_plot: 60
  max_archive_locs: 60
  min_archive_days: 60  # Minimum days of version history to enable revision analysis
knitr:
  opts_chunk:
    comment: "#>"
    message: FALSE
    warning: FALSE
    out.width: "100%"
    dev: "png"
    fig.width: 8
    fig.height: 4.5
    fig.align: center
editor_options: 
  chunk_output_type: console
execute:
  cache: false
---

```{r}
#| label: setup
#| include: false
#| cache: false
# When running interactively, populate params from defaults
if (!exists("params")) {
    params <- list(
        source = "beta_pophive",
        signal = "epic_pct_covid_total",
        name = "PopHive: Epic % COVID Total",
        input_csv = NULL,
        geo_type = "state",
        time_type = NULL,
        end_day = format(Sys.Date()),
        start_day = format(Sys.Date() - 365 * 2),
        extra_keys = character(0),
        season_start_month = 7,
        smooth_quantiles = TRUE,
        show_geo_outliers = TRUE,
        show_versioning = TRUE,
        max_locations_plot = 60,
        max_archive_locs = 60,
        min_archive_days = 60
    )
}

# Runtime defaults for dynamic params. Unlock to allow modification.
if (exists("params", envir = .GlobalEnv, inherits = FALSE) &&
    bindingIsLocked("params", .GlobalEnv)) {
    unlockBinding("params", .GlobalEnv)
}
# Helper to flatten parameter lists (e.g. when Quarto CLI parses colons as nested lists)
flatten_param <- function(x) {
    if (is.null(x)) return(NULL)
    if (is.list(x)) {
        if (!is.null(names(x)) && any(nzchar(names(x)))) {
            paste(names(x), unlist(x), sep = ": ")
        } else {
            as.character(unlist(x))
        }
    } else {
        as.character(x)
    }
}

params$source <- flatten_param(params$source)
params$signal <- flatten_param(params$signal)
params$name <- flatten_param(params$name)
params$geo_type <- flatten_param(params$geo_type)
params$time_type <- flatten_param(params$time_type)
params$input_csv <- flatten_param(params$input_csv)

# Helper to resolve relative file paths across render working directories
resolve_path <- function(path) {
    if (is.null(path) || !nzchar(path)) return(path)
    if (file.exists(path)) return(path)
    cur_dir <- basename(getwd())
    prefix <- paste0("^", cur_dir, "/")
    if (grepl(prefix, path)) {
        stripped_path <- sub(prefix, "", path)
        if (file.exists(stripped_path)) {
            return(stripped_path)
        }
    }
    parent_path <- file.path("..", path)
    if (file.exists(parent_path)) {
        return(parent_path)
    }
    return(path)
}
params$input_csv <- resolve_path(params$input_csv)

if (!is.null(params$input_csv) && nzchar(params$input_csv)) {
    if (!file.exists(params$input_csv)) {
        stop(sprintf("Local CSV file not found at path: %s", params$input_csv))
    }
}

# When using CSV route, source/indicator labels default to the CSV filename
if (!is.null(params$input_csv) &&
    (is.null(params$source) || !nzchar(params$source))) {
    params$source <- tools::file_path_sans_ext(basename(params$input_csv))
}

if (is.null(params$end_day)   || !nzchar(params$end_day))   params$end_day   <- format(Sys.Date())
if (is.null(params$start_day) || !nzchar(params$start_day)) params$start_day <- format(as.Date(params$end_day) - 365 * 2)

# Load libraries
if (!require("pak")) install.packages("pak")
if (!requireNamespace("epiprocess", quietly = TRUE)) pak::pkg_install("cmu-delphi/epiprocess@dev")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    epidatr, epiprocess,
    tidyverse, scales, patchwork,
    here, glue, ggtext
)
# Custom HTML-friendly label wrapper for ggtext::element_markdown
label_wrap_html <- function(width = 25) {
  fun <- label_wrap_gen(width = width)
  function(labels) {
    res <- fun(labels)
    res[] <- lapply(res, function(x) {
      if (is.character(x) || is.factor(x)) {
        gsub("\n", "<br>", as.character(x), fixed = TRUE)
      } else {
        x
      }
    })
    res
  }
}

# Invalidate knitr cache when params change
knitr::opts_chunk$set(cache.extra = list(params))

# Normalize signal/name params to unique character vectors.
signals_vec <- if (!is.null(params$signal) && length(params$signal) > 0 &&
    any(nzchar(as.character(params$signal)))) {
    unique(as.character(params$signal)[nzchar(as.character(params$signal))])
} else if (!is.null(params$input_csv) && file.exists(params$input_csv)) {
    # Signal omitted: peek at CSV to resolve signal names early (needed for config table)
    csv_cols <- readr::read_csv(params$input_csv, n_max = 0, show_col_types = FALSE)
    if ("signal" %in% colnames(csv_cols)) {
        sort(unique(readr::read_csv(params$input_csv, col_select = "signal", show_col_types = FALSE)$signal))
    } else {
        "value"
    }
} else {
    character(0)
}
multi_signal <- length(signals_vec) > 1

# Keys beyond geo_value/time_value. With more than one indicator, `signal` is
# part of the key as well: without it, (geo_value, time_value, stratum) is not
# unique and the epi_df/epi_archive constructors reject the data.
candidate_keys <- c(if (multi_signal) "signal", params$extra_keys)

# Prose helpers, so the narrative text is plural only when more than one
# indicator was loaded. Used inline in the markdown between the chunks.
ind <- if (multi_signal) "indicators" else "indicator"
ind_poss <- if (multi_signal) "indicators'" else "indicator's"
sg_pl <- function(singular, plural) if (multi_signal) plural else singular

# Build signal labels (name falls back to signal if not provided)
name_vec <- {
    n <- params$name
    if (!is.null(n) && length(n) == length(signals_vec)) as.character(n)
    else signals_vec
}
name_vec <- str_wrap(name_vec, width = 15)
signal_labels <- setNames(name_vec, signals_vec)

# Resolve geo_type parameter (infer if not provided)
if (is.null(params$geo_type) || !nzchar(params$geo_type)) {
    if (!is.null(params$input_csv) && file.exists(params$input_csv)) {
        sample_df <- readr::read_csv(params$input_csv, n_max = 5, show_col_types = FALSE) %>%
                epiprocess::as_epi_df()
        params$geo_type <- attributes(sample_df)$metadata$geo_type
    } else {
        stop("Parameter 'geo_type' is required when querying datasets from the API.")
    }
}

# Set theme and colors
colours <- c(
    "#172431",
    "#ffffff",
    "#C41230",
    "#941120",
    "#007BC0",
    "#002676",
    "#FDB515",
    "#002145",
    "#00A7E1",
    "#8C338F",
    "#368f33"
)

# Palette selection
dis_pal_list <- colours[c(3, 5, 7, 10, 11, 9)]

theme_set(
    theme_minimal(base_size = 13) +
        theme(
            plot.title = ggtext::element_textbox_simple(hjust = 0.5, halign = 0.5, size = 15, face = "bold", margin = margin(b = 6)),
            plot.subtitle = element_text(hjust = 0.5, color = "gray50", margin = margin(b = 10)),
            plot.caption = ggtext::element_textbox_simple(hjust = 0, halign = 0, color = "gray50", margin = margin(t = 10)),
            panel.grid = element_line(linetype = 2, linewidth = 0.3, color = "gray90"),
            panel.grid.minor = element_blank(),
            legend.position = "top",
            legend.text = ggtext::element_markdown(size = 10),
            legend.title = ggtext::element_markdown(size = 10, face = "bold"),
            strip.text = ggtext::element_markdown(hjust = 0.5, size = 11, face = "bold", margin = margin(t = 4, b = 4)),
            strip.background = element_rect(fill = "gray95", linetype = "blank"),
            panel.border = element_rect(color = "gray95", fill = NA)
        )
)

# Palette as global defaults for discrete and continuous scales
options(
    ggplot2.discrete.colour = dis_pal_list,
    ggplot2.discrete.fill = dis_pal_list,
    ggplot2.continuous.colour = function() scale_colour_gradientn(colours = dis_pal_list),
    ggplot2.continuous.fill = function() scale_fill_gradientn(colours = dis_pal_list)
)

# Update defaults for geoms/stats
tibble::tribble(
    ~name,      ~aes,     ~alpha, ~is_stat,
    "line",     "colour", 1.0,    FALSE,
    "point",    "colour", 0.7,    FALSE,
    "function", "colour", 1.0,    FALSE,
    "bar",      "fill",   0.9,    FALSE,
    "col",      "fill",   0.9,    FALSE,
    "boxplot",  "fill",   0.8,    FALSE,
    "bin",      "fill",   0.8,    TRUE,
    "identity", "fill",   0.7,    TRUE
) %>%
    purrr::pwalk(function(name, aes, alpha, is_stat) {
        updater <- if (is_stat) update_stat_defaults else update_geom_defaults
        updater(name, setNames(list(dis_pal_list[1], alpha), c(aes, "alpha")))
    })

# Helper: round any date to the immediately preceding Sunday (or same day if already Sunday)
round_to_sunday <- function(d) d - as.integer(format(d, "%w"))

# Helper: build epirange with correct format for the indicator's time_type
time_range <- function(start, end, time_type = params$time_type) {
    if (time_type == "week") {
        to_yw <- function(d) {
            d <- as.Date(d)
            m <- MMWRweek::MMWRweek(d)
            m$MMWRyear * 100L + m$MMWRweek
        }
        epirange(to_yw(start), to_yw(end))
    } else {
        epirange(start, end)
    }
}

# Helper: map geo codes to readable names
geo_to_name <- function(geos, geo_type = params$geo_type) {
    if (geo_type == "state") {
        names <- covidcast::abbr_to_name(toupper(geos))
    } else if (geo_type == "county") {
        names <- covidcast::county_fips_to_name(as.character(geos))
    } else if (geo_type == "msa") {
        names <- covidcast::cbsa_to_name(as.character(geos))
    } else {
        return(geos)
    }
    # Fallback to original codes for any missing mappings
    names[is.na(names)] <- geos[is.na(names)]
    return(names)
}

# Helper: consistent random sample from a vector
random_sample <- function(vec, n = 15) {
    if (length(vec) <= n) {
        return(vec)
    }
    withr::with_seed(42, sort(sample(vec, n)))
}

# Helper function to print list (up to 15 obs)
format_excluded <- function(locs, names) {
    if (length(locs) == 0) {
        return(NULL)
    }
    sampled_idx <- match(random_sample(locs), locs)
    sampled_locs <- locs[sampled_idx]
    sampled_names <- names[sampled_idx]
    entries <- ifelse(
        sampled_locs == sampled_names,
        sampled_locs,
        sprintf("%s (%s)", sampled_locs, sampled_names)
    )
    if (length(locs) > 15) {
        paste0(paste(entries, collapse = ", "), ", etc.")
    } else {
        paste(entries, collapse = ", ")
    }
}

# Helper: order stratum levels the way a reader expects rather than
# alphabetically. Any aggregate level ("all", "total", ...) comes first, then the
# remaining levels by their leading number so "5-17" precedes "18-49", with
# open-ended low bins such as "<1" sorted below their own number.
order_strata <- function(x) {
    lv <- sort(unique(as.character(x[!is.na(x)])))
    is_agg <- tolower(lv) %in% c("all", "total", "overall", "all ages", "all_ages", "any")
    rest <- lv[!is_agg]
    num <- suppressWarnings(as.numeric(str_extract(rest, "[0-9]+\\.?[0-9]*")))
    num <- num - 0.5 * str_detect(tolower(rest), "^\\s*(<|under|less|up to)")
    factor(x, levels = c(lv[is_agg], rest[order(num, rest, na.last = TRUE)]))
}

# Helper: order stratum columns and append the number of locations reporting
# each level to its label. Panels drawn from different location sets are not
# directly comparable, so the label has to say how many locations each one has.
label_strata_coverage <- function(df, keys, src = df_candidate) {
    src <- as_tibble(src) %>% filter(!is.na(value))
    for (k in keys) {
        f <- order_strata(df[[k]])
        lv <- levels(f)
        cnt <- src %>%
            group_by(.data[[k]]) %>%
            summarize(n = n_distinct(geo_value), .groups = "drop")
        n <- cnt$n[match(lv, as.character(cnt[[k]]))]
        df[[k]] <- factor(f,
            levels = lv,
            labels = if_else(is.na(n), lv, sprintf("%s (%s locs)", lv, n))
        )
    }
    df
}

# Helper: interactive time series with one dropdown entry per location and one
# trace per stratum. Missing periods are left as gaps instead of being bridged
# by the line, and markers show which periods actually carry a value. A
# confidence band (when the data has one) shares its stratum's legend entry, so
# hiding a stratum hides its band with it.
plot_series_dropdown <- function(df, trace_col = NULL, ci_cols = NULL) {
    groups <- as.character(sort(unique(df$geo_value)))
    # An estimate on a near-empty denominator carries an interval about as wide
    # as the indicator's entire range. Drawn, it is a full-height column that
    # buries every series behind it while saying nothing, so leave those out and
    # report how many below the plot. The cut-off is the span of the values
    # themselves, so it scales with the indicator instead of being a constant.
    band_limit <- if (!is.null(ci_cols)) diff(range(df$value, na.rm = TRUE)) else NULL
    n_band_wide <- 0L
    n_band_obs <- 0L
    strata <- if (is.null(trace_col)) NA_character_ else levels(order_strata(df[[trace_col]]))
    # One distinct colour per stratum. Recycling the palette would give two
    # strata the same colour, which is unreadable once bands are drawn too.
    pal <- setNames(
        if (length(strata) <= length(dis_pal_list)) {
            dis_pal_list[seq_along(strata)]
        } else {
            grDevices::colorRampPalette(dis_pal_list)(length(strata))
        },
        strata
    )

    specs <- list()
    for (g in groups) {
        for (s in strata) {
            d <- df %>% filter(geo_value == g)
            if (!is.na(s)) d <- d %>% filter(.data[[trace_col]] == s)
            d <- d %>% arrange(time_value)
            if (nrow(d) == 0 || all(is.na(d$value))) next
            lab <- if (is.na(s)) "Value" else s
            col <- unname(if (is.na(s)) dis_pal_list[1] else pal[[s]])
            if (!is.null(ci_cols)) {
                # One band per contiguous run of reported bounds. A band is
                # drawn as a c(x, rev(x)) polygon, and an NA anywhere inside it
                # shifts the return leg, pairing each lower bound with the wrong
                # period and detaching the band from its line. Splitting on the
                # gaps also stops a band from spanning a period with no data,
                # which is what the line already does.
                ok <- !is.na(d[[ci_cols[1]]]) & !is.na(d[[ci_cols[2]]])
                too_wide <- ok & (d[[ci_cols[2]]] - d[[ci_cols[1]]]) > band_limit
                n_band_obs <- n_band_obs + sum(ok)
                n_band_wide <- n_band_wide + sum(too_wide)
                ok <- ok & !too_wide
                for (seg in split(d[ok, , drop = FALSE], cumsum(!ok)[ok])) {
                    specs <- append(specs, list(list(
                        d = seg, col = col, lab = lab, grp = g,
                        kind = "ci", legend_ok = FALSE
                    )))
                }
            }
            specs <- append(specs, list(list(
                d = d, col = col, lab = lab, grp = g,
                kind = "line", legend_ok = !is.na(s)
            )))
        }
    }
    if (length(specs) == 0) return(NULL)

    trace_grp <- map_chr(specs, "grp")
    legend_ok <- map_lgl(specs, "legend_ok")
    active <- trace_grp == groups[1]

    p <- reduce(seq_along(specs), function(p, i) {
        sp <- specs[[i]]
        if (sp$kind == "ci") {
            plotly::add_ribbons(p,
                x = sp$d$time_value,
                ymin = sp$d[[ci_cols[1]]], ymax = sp$d[[ci_cols[2]]],
                line = list(width = 0), fillcolor = plotly::toRGB(sp$col, 0.2),
                name = sp$lab, legendgroup = sp$lab,
                showlegend = FALSE, visible = active[i],
                hoverinfo = "skip"
            )
        } else {
            plotly::add_trace(p,
                x = sp$d$time_value, y = sp$d$value,
                type = "scatter", mode = "lines+markers", connectgaps = FALSE,
                line = list(color = sp$col, width = 1.5),
                marker = list(color = sp$col, size = 3),
                name = sp$lab, legendgroup = sp$lab,
                showlegend = active[i] && legend_ok[i], visible = active[i],
                hovertemplate = paste0(
                    "%{x|%Y-%m-%d}<br>Value: %{y:.3f}",
                    "<extra>", sp$lab, "</extra>"
                )
            )
        }
    }, .init = plotly::plot_ly())

    # Scale the y axis to the values, not to the bands. A near-empty denominator
    # produces a confidence interval spanning almost the whole range of the
    # indicator (here 10% of Alabama's rows reach above 90%), and letting those
    # set the axis compresses every series into a sliver along the bottom, so the
    # line reads as if its band belonged to something else. Bands reaching past
    # the top are clipped there; the y axis stays zoomable so they can be seen.
    yrange <- NULL
    if (!is.null(ci_cols)) {
        vals <- df$value[is.finite(df$value)]
        if (length(vals) > 0 && diff(range(vals)) > 0) {
            yrange <- c(min(0, min(vals)), max(vals) + 0.15 * diff(range(vals)))
        }
    }

    buttons <- map(groups, function(g) {
        on <- trace_grp == g
        list(
            method = "update",
            args = list(
                list(visible = as.list(on), showlegend = as.list(on & legend_ok)),
                list(title = list(text = paste0("Location: ", g)))
            ),
            label = g
        )
    })

    p <- p %>%
        plotly::layout(
            title = list(text = paste0("Location: ", groups[1])),
            xaxis = list(title = "Date"),
            yaxis = list(title = "Value", range = yrange),
            legend = list(title = list(text = trace_col %||% "")),
            updatemenus = list(list(
                type = "dropdown", active = 0, buttons = buttons, x = 0.05, y = 1.15
            ))
        ) %>%
        plotly::config(modeBarButtonsToRemove = c("zoomIn2d", "zoomOut2d"))

    # Counts ride along on the returned plot so the chunk can report them.
    attr(p, "bands_wide") <- n_band_wide
    attr(p, "bands_total") <- n_band_obs
    p
}

# Optional-section switches; all default to on when the param is absent.
smooth_quantiles <- !isFALSE(params$smooth_quantiles)
show_geo_outliers <- !isFALSE(params$show_geo_outliers)
show_versioning <- !isFALSE(params$show_versioning)

# Helper: assign each observation to an epidemiological year starting in
# `season_start_month` (7 = July-June, the usual influenza convention).
season_start_month <- as.integer(params$season_start_month %||% 7L)
season_label <- sprintf(
    "%s-%s", month.name[season_start_month],
    month.name[((season_start_month + 10L) %% 12L) + 1L]
)
add_season <- function(df) {
    df %>%
        mutate(
            yr = as.integer(format(time_value, "%Y")),
            month = as.integer(format(time_value, "%m")),
            season_start = if_else(month >= season_start_month, yr, yr - 1L),
            season = paste0(season_start, "-", substr(as.character(season_start + 1L), 3L, 4L)),
            season_month = ((month - season_start_month) %% 12L) + 1L
        )
}
season_month_labels <- function(x) {
    month.abb[((season_start_month - 1L + x - 1L) %% 12L) + 1L]
}

# Set cache
cache_path <- file.path(getwd(), "epidatr_cache")
if (!dir.exists(cache_path)) dir.create(cache_path, recursive = TRUE)

set_cache(cache_dir = cache_path)
```

This notebook evaluates `r sg_pl("a **candidate indicator**", "the **candidate indicators**")` to understand `r sg_pl("its", "their")` characteristics, coverage, missingness, and versioning/revision behavior.

```{r}
#| label: print-config
#| echo: false
config_tbl <- tibble(
    Field = c("Source", "Indicator(s)", "Label(s)", "Geography", "Stratifier(s)", "Requested period"),
    Value = c(
        params$source,
        paste(signals_vec, collapse = ", "),
        paste(str_remove(name_vec, "\\n"), collapse = ", "),
        if (is.null(params$geo_type) || !nzchar(params$geo_type)) NA_character_ else params$geo_type,
        if (length(params$extra_keys) == 0) NA_character_ else paste(params$extra_keys, collapse = ", "),
        paste(params$start_day, "to", params$end_day)
    )
) %>%
    filter(!is.na(Value))

DT::datatable(
    config_tbl,
    caption = "Analysis configuration",
    rownames = FALSE,
    options = list(scrollX = TRUE, dom = "t", paging = FALSE)
)
```

---

## Data Loading {#data}

```{r}
#| label: load-candidate
#| results: asis
# Load candidate indicator
candidate_archive <- NULL
if (!is.null(params$input_csv)) {
    df_candidate_raw <- read.csv(params$input_csv) %>%
        mutate(time_value = as.Date(time_value)) %>%
        filter(time_value >= as.Date(params$start_day) & time_value <= as.Date(params$end_day))

    # Add signal column if missing (required for long signal format)
    if (!("signal" %in% colnames(df_candidate_raw))) {
        df_candidate_raw$signal <- if (length(signals_vec) > 0) signals_vec[1] else "value"
    }

    # CSV may include version/issue columns for revision analysis
    has_versions <- any(c("version", "issue") %in% colnames(df_candidate_raw))
    if (has_versions) {
        df_candidate_raw <- df_candidate_raw %>%
            {
                if ("issue" %in% names(.)) rename(., version = issue) else .
            } %>%
            mutate(version = as.Date(version))

        # A row with no version cannot be placed on a version timeline. Drop
        # those rows and say how many, because a file holding both versioned and
        # unversioned copies of the same observations is usually a build
        # accident, and keeping them would double every series.
        n_no_version <- sum(is.na(df_candidate_raw$version))
        if (n_no_version > 0) {
            cat(sprintf(
                "- **Warning**: %s of %s CSV rows carry no version and were dropped.\n",
                label_comma()(n_no_version), label_comma()(nrow(df_candidate_raw))
            ))
            df_candidate_raw <- df_candidate_raw %>% filter(!is.na(version))
        }

        # Confidence bounds ride along as extra value columns, so the overview
        # plot can draw bands for versioned CSVs too.
        csv_archive_df  <- df_candidate_raw %>%
            select(geo_value, time_value, version, any_of("signal"), any_of(params$extra_keys),
                   value, any_of(c("ci_lower", "ci_upper"))) %>%
            group_by(geo_value, time_value, version, pick(any_of(c("signal", params$extra_keys)))) %>%
            slice_max(value, n = 1, with_ties = FALSE) %>%
            ungroup()

        candidate_archive <- csv_archive_df %>%
            as_epi_archive(signal_format = "as_is", other_keys = candidate_keys)
        df_candidate <- epix_as_of(candidate_archive, candidate_archive$versions_end)
        cat(sprintf(
            "- CSV contains version history: %s versions\n",
            label_comma()(n_distinct(df_candidate_raw$version))
        ))
    } else {
        df_candidate <- df_candidate_raw %>%
            as_epi_df(signal_format = "as_is", other_keys = candidate_keys)
    }
} else {
    # Determine if new API can be used for candidate source
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE
    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                all(signals_vec %in% meta_res[[params$source]]$signals)
            },
            error = function(e) FALSE
        )
        if (meta_available) use_new_api <- TRUE
    }
    if (use_new_api) {
        # epidata_snapshot does not accept multiple signals
        df_candidate <- purrr::map(signals_vec, ~ {
            epidata_snapshot(
                source = params$source,
                signals = .x,
                geo_type = params$geo_type
            ) %>%
                rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
                {
                    if ("fill_method" %in% colnames(.)) {
                        filter(., fill_method == "source" | !any(fill_method == "source"))
                    } else .
                } %>%
                # {
                #     if ("version" %in% colnames(.)) {
                #         group_by(., geo_value, time_value) %>%
                #             slice_max(version, n = 1, with_ties = FALSE) %>%
                #             ungroup()
                #     } else .
                # } %>%
                filter(
                    time_value >= as.Date(params$start_day),
                    time_value <= as.Date(params$end_day)
                ) %>%
                select(geo_value, any_of("signal"), time_value, any_of(params$extra_keys),
                       value, any_of(c("ci_lower", "ci_upper")))
        }) %>%
            bind_rows() %>%
            as_epi_df(signal_format = "as_is", other_keys = candidate_keys)
    } else {
        # pub_covidcast accepts a signal vector and returns a signal column
        df_candidate <- pub_covidcast(
            params$source, signals_vec,
            params$geo_type, params$time_type,
            time_values = time_range(params$start_day, params$end_day),
            as_of = Sys.Date()
        ) %>% as_epi_df(signal_format = "as_is", other_keys = candidate_keys)
    }
}

# Determine time type and number of locations. Resolved here (rather than after
# the archive fetch) so the summary below can report periods and strata.
env_time_type <- attributes(df_candidate)$metadata$time_type
if (is.null(env_time_type) || length(env_time_type) != 1 || is.na(env_time_type) || !is.character(env_time_type)) {
    env_time_type <- params$time_type
}
if (is.null(env_time_type) || !nzchar(env_time_type)) {
    env_time_type <- "day"
}
params$time_type <- env_time_type

env_geo_type <- attributes(df_candidate)$metadata$geo_type
if (is.null(params$geo_type) || length(params$geo_type) != 1 || is.na(params$geo_type) || !is.character(params$geo_type)) {
    params$geo_type <- env_geo_type
}

n_locations <- n_distinct(df_candidate$geo_value)
n_periods <- n_distinct(df_candidate$time_value)

# Stratum levels per extra key, and the number of (period x stratum) cells a
# location could in principle fill. Used here and by the missingness section.
strata_levels <- map(set_names(params$extra_keys), ~ sort(unique(as.character(df_candidate[[.x]]))))
n_key_combos <- if (length(params$extra_keys) > 0) {
    df_candidate %>% as_tibble() %>% distinct(across(all_of(params$extra_keys))) %>% nrow()
} else {
    1L
}
n_strata_cells <- max(1L, prod(lengths(strata_levels)))

cat(sprintf(
    "- **Summary**: %s rows across %s locations, %s %s period(s)%s, and %d indicator(s)\n",
    label_comma()(nrow(df_candidate)),
    label_comma()(n_locations),
    label_comma()(n_periods), env_time_type,
    if (length(params$extra_keys) > 0) {
        sprintf(
            ", stratified by %s",
            paste(sprintf(
                "%s (%d level(s): %s)", names(strata_levels), lengths(strata_levels),
                map_chr(strata_levels, ~ paste(levels(order_strata(.x)), collapse = ", "))
            ), collapse = "; ")
        )
    } else "",
    length(signals_vec)
))

# Length of the requested window, in the indicator's own periods
req_days <- as.integer(as.Date(params$end_day) - as.Date(params$start_day)) + 1L
req_periods <- round(req_days / switch(env_time_type,
    day = 1, week = 7, month = 30.44, year = 365.25, 1
))
cat(sprintf(
    "- **Observed period**: %s to %s (requested %s to %s, about %s %s(s))\n",
    format(min(df_candidate$time_value)), format(max(df_candidate$time_value)),
    params$start_day, params$end_day, label_comma()(req_periods), env_time_type
))
cat(sprintf(
    "- **Grid**: %s %s period(s) x %s stratum cell(s) = %s possible values per location\n",
    label_comma()(n_periods), env_time_type,
    label_comma()(n_strata_cells), label_comma()(n_periods * n_strata_cells)
))
```

```{r}
#| label: load-candidate-archive
#| results: asis
# Fetch revision history from API if not already built from CSV. Skipped entirely
# when the versioning section is off — the issues = "*" query is the slowest one.
if (show_versioning && is.null(candidate_archive) && is.null(params$input_csv)) {
    # Only fetch version history if the source supports revisions
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE
    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                all(signals_vec %in% meta_res[[params$source]]$signals)
            },
            error = function(e) FALSE
        )
        if (meta_available) use_new_api <- TRUE
    }

    # Fetch archive for revision behavior
    archive_geos <- unique(df_candidate$geo_value)
    # Limit number of locations for version history queries to avoid API limits
    if (length(archive_geos) > params$max_locations_plot) {
        archive_geos <- random_sample(archive_geos, params$max_locations_plot)
    }

    # Determine window of issues available
    archive_start <- as.Date(params$start_day) - 60

    archive_info <- list(
        n_sampled = length(archive_geos),
        n_total = length(unique(df_candidate$geo_value)),
        time_value_start = archive_start,
        time_value_end = params$end_day
    )

    if (use_new_api) {
        # epidata_archive does not accept multiple signals
        df_archive_raw <- purrr::map(signals_vec, ~ {
            epidata_archive(
                source = params$source,
                signals = .x,
                geo_type = params$geo_type
            ) %>%
                rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
                filter(
                    geo_value %in% archive_geos,
                    time_value >= as.Date(archive_start),
                    time_value <= as.Date(params$end_day)
                )
        }) %>% bind_rows()
    } else {
        # pub_covidcast accepts a signal vector and returns a signal column
        df_archive_raw <- pub_covidcast(
            source = params$source,
            signal = signals_vec,
            geo_type = params$geo_type,
            time_type = params$time_type,
            geo_values = archive_geos,
            time_values = time_range(archive_start, params$end_day),
            issues = "*",
        )
    }

    if (nrow(df_archive_raw) > 0) {
        df_archive_raw <- df_archive_raw %>%
            rename(any_of(c(time_value = "reference_time", version = "issue"))) %>%
            rename(any_of(c(version = "report_time"))) %>%
            select(geo_value, time_value, version, any_of("signal"), any_of(params$extra_keys), value) #%>%
            # Deduplicate: sources like NWSS can have multiple rows per
            # (geo, time, version) from different measurement sources.
            # group_by(geo_value, time_value, version, across(any_of("signal"))) %>%
            # slice_max(value, n = 1, with_ties = FALSE) %>%
            # ungroup()

        if ("version" %in% colnames(df_archive_raw)) {
            archive_info$version_start <- min(df_archive_raw$version)
            archive_info$version_end <- max(df_archive_raw$version)
            candidate_archive <- df_archive_raw %>%
                as_epi_archive(signal_format = "as_is", other_keys = candidate_keys)
        }
    }
}
```

```{r}
#| label: quarto-metadata
#| results: asis
#| echo: false
# Dynamically set Quarto metadata so we can use it in ::: {.content-visible}
# blocks later.
cat(sprintf(
    "\n---\nmulti_location: %s\ngeo_outliers: %s\nversioning: %s\n---\n",
    tolower(n_locations > 1),
    tolower(show_geo_outliers && n_locations > 1),
    tolower(show_versioning)
))
```

## Exploratory Data Analysis {#eda}

The data is aggregated at the `r params$geo_type` level. The dataset contains `r n_locations` valid locations measured at a `r env_time_type`-level frequency.

Throughout this notebook, an **observation** (equivalently, a reported value) is one row of the dataset: a single value for one location, one `r env_time_type``r if (length(params$extra_keys) > 0) sprintf(" and one level of %s", paste(params$extra_keys, collapse = "/")) else ""`. A **location** is one `r params$geo_type`, and counts of locations are counts of distinct `r params$geo_type`s, not of observations.

::: {.content-visible when-meta="multi_location"}
### Coverage and Missingness

This section comes first because everything after it should be read as a statement about the values that exist. It explores the availability of data across time and space, so that the plots and summaries below can be understood as covering only the non-missing observations.

A value is treated as **missing** when the dataset contains no row for a given (location, period`r if (length(params$extra_keys) > 0) ", stratum" else ""`) cell. The grid of cells a location could fill is the `r label_comma()(n_periods)` `r env_time_type` period(s) observed anywhere in the data`r if (n_strata_cells > 1) sprintf(" times the %s stratum cell(s)", label_comma()(n_strata_cells)) else ""`, so each location has `r label_comma()(n_periods * n_strata_cells)` possible values. Periods before the first date any location reports are outside this grid entirely; the Data Loading summary above contrasts the observed period with the requested one.

The area chart below tracks the number of geographic locations with valid data on any given date, against the `r n_locations` locations present in the dataset (dotted line).

```{r}
#| label: eda-coverage
#| echo: false
#| results: asis
coverage <- df_candidate %>%
    group_by(across(all_of(c("signal", "time_value", params$extra_keys)))) %>%
    summarize(n_available = sum(!is.na(value)), .groups = "drop") %>%
    mutate(geo_value = "all") %>%
    as_epi_df(signal_format = "as_is", other_keys = c("signal", params$extra_keys)) %>%
    as_tibble() %>%
    mutate(indicator = signal_labels[signal]) %>%
    label_strata_coverage(params$extra_keys)

# The full (period x stratum) x location grid, with an NA value wherever a
# location has no row. Every missingness figure below is derived from this, so
# the denominator is the same in the per-location and per-stratum views.
summary_base <- df_candidate %>%
    as_tibble() %>%
    select(all_of(c("signal", "geo_value", "time_value", params$extra_keys)), value) %>%
    complete(nesting(signal, time_value), !!!rlang::syms(params$extra_keys), geo_value)
```

```{r}
#| label: eda-coverage-plot
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 2), 12)
#| echo: true
ggplot(coverage, aes(x = time_value, y = n_available,
                     color = indicator, linetype = indicator)) +
    geom_hline(yintercept = n_locations, linetype = 3, linewidth = 0.4, color = "gray50") +
    geom_line(linewidth = 0.8) +
    scale_x_date(breaks = breaks_pretty(), label = label_date_short()) +
    scale_y_continuous(labels = label_comma(), limits = c(0, NA)) +
    labs(
        title = "Data coverage over time",
        subtitle = sprintf("Number of locations with non-missing values per %s (dotted line: %d locations in the dataset)",
                           env_time_type, n_locations),
        x = "Date", y = "Number of locations",
        color = "Indicator", linetype = "Indicator"
    ) +
    {if (length(signals_vec) > 6 || length(params$extra_keys) > 0) facet_wrap(reformulate(c(if (length(signals_vec) > 6) "indicator", params$extra_keys)), ncol = 2, labeller = label_wrap_html(width = 25)) else NULL}
```

#### Missingness by Stratifier

```{r}
#| label: eda-missingness-by-stratum
#| echo: false
#| eval: !expr length(params$extra_keys) > 0
strat_miss <- summary_base %>%
    group_by(across(all_of(c("signal", params$extra_keys)))) %>%
    summarize(
        n_possible = n(),
        n_obs = sum(!is.na(value)),
        n_locs = n_distinct(geo_value[!is.na(value)]),
        n_periods_obs = n_distinct(time_value[!is.na(value)]),
        mean_val = mean(value, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(
        pct_missing = 100 * (1 - n_obs / n_possible),
        signal_label = signal_labels[signal],
        across(all_of(params$extra_keys), order_strata)
    ) %>%
    arrange(across(all_of(params$extra_keys)), signal) %>%
    {
        if (multi_signal) select(., signal_label, all_of(params$extra_keys), n_locs, n_periods_obs, n_obs, n_possible, pct_missing, mean_val)
        else select(., all_of(params$extra_keys), n_locs, n_periods_obs, n_obs, n_possible, pct_missing, mean_val)
    }

DT::datatable(
    strat_miss,
    colnames = c(
        if (multi_signal) "Indicator" else NULL,
        str_to_title(str_replace_all(params$extra_keys, "_", " ")),
        "Locations Reporting",
        sprintf("%ss Reporting", str_to_title(env_time_type)),
        "Values Reported", "Possible Values", "% Missing", "Avg. Value"
    ),
    caption = sprintf(
        "Missingness by stratifier. Possible values = %s %s period(s) x %s location(s) for each stratum.",
        label_comma()(n_periods), env_time_type, label_comma()(n_locations)
    ),
    rownames = FALSE, options = list(scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
) %>% DT::formatRound(columns = if (multi_signal) 7:8 else 6:7, digits = 2)
```

:::

#### Summary Statistics by Location

This table summarizes the available data across all included locations. Locations with no candidate information are not considered.

```{r}
#| label: eda-summary
#| results: asis
#| echo: false
# Compute base stats per (signal, geo_value) against the full grid, so that a
# location missing an entire stratum is counted as missing rather than ignored.
summary_tbl_all <- summary_base %>%
    group_by(signal, geo_value) %>%
    summarize(
        n_possible = n(),
        n_obs = sum(!is.na(value)),
        cand_mean = mean(value, na.rm = TRUE),
        cand_var = var(value, na.rm = TRUE),
        cand_miss_frac = 1 - n_obs / n_possible,
        cand_miss = sprintf("%.1f%%", 100 * (1 - n_obs / n_possible)),
        .groups = "drop"
    ) %>%
    mutate(
        geo_name     = geo_to_name(geo_value),
        signal_label = signal_labels[signal]
    )

excluded_locs_df <- summary_tbl_all %>%
    filter(is.na(cand_mean) | cand_miss_frac == 1)
if (nrow(excluded_locs_df) > 0) {
    for (sig in signals_vec) {
        ex_sig <- excluded_locs_df %>% filter(signal == sig)
        if (nrow(ex_sig) > 0) {
            cat(sprintf(
                "**Excluded locations (no data) — %s**: %s\n\n",
                signal_labels[sig],
                format_excluded(ex_sig$geo_value, ex_sig$geo_name)
            ))
        }
    }
}

zero_var_cand_locs_df <- summary_tbl_all %>%
    filter(!is.na(cand_var) & cand_var == 0 & cand_miss_frac < 1)
if (nrow(zero_var_cand_locs_df) > 0) {
    for (sig in signals_vec) {
        zv_sig <- zero_var_cand_locs_df %>% filter(signal == sig)
        if (nrow(zv_sig) > 0) {
            cat(sprintf(
                "**Zero variance locations (flatlined) — %s**: %s\n\n",
                signal_labels[sig],
                format_excluded(zv_sig$geo_value, zv_sig$geo_name)
            ))
        }
    }
}

# Summary of missingness levels per signal
miss_levels <- summary_tbl_all %>%
    mutate(
        cand_miss_group = case_when(
            cand_miss_frac == 1 ~ "100% missing",
            cand_miss_frac < 0.05 ~ "<5% missing",
            cand_miss_frac < 0.25 ~ "5-25% missing",
            cand_miss_frac < 0.50 ~ "25-50% missing",
            TRUE ~ "50-99.9% missing"
        ),
        cand_miss_group = factor(cand_miss_group, levels = c("<5% missing", "5-25% missing", "25-50% missing", "50-99.9% missing", "100% missing"))
    )

for (sig in signals_vec) {
    cat(sprintf("**Missingness Overview%s:**\n\n",
                if (multi_signal) sprintf(" — %s", signal_labels[sig]) else " (All Locations)"))
    cand_counts <- miss_levels %>%
        filter(signal == sig) %>%
        count(cand_miss_group) %>%
        complete(cand_miss_group, fill = list(n = 0))
    for (i in seq_len(nrow(cand_counts))) {
        cat(sprintf("- **%s**: %d locations\n",
                    cand_counts$cand_miss_group[i], cand_counts$n[i]))
    }
    cat("\n")
}

summary_tbl <- summary_tbl_all %>%
    filter(!is.na(cand_mean) & cand_miss_frac < 1) %>%
    arrange(desc(cand_miss_frac)) %>%
    {
        if (multi_signal) select(., signal_label, geo_value, geo_name, n_obs, n_possible, cand_mean, cand_miss)
        else select(., geo_value, geo_name, n_obs, n_possible, cand_mean, cand_miss)
    }

DT::datatable(
    summary_tbl,
    colnames = c(
        if (multi_signal) "Indicator" else NULL,
        sprintf("Geo-level (%s)", params$geo_type),
        "Name",
        "Values Reported",
        "Possible Values",
        "Avg. Value",
        "% Missing"
    ),
    caption = sprintf(
        "Summary statistics by location. Possible values per location = %s %s period(s)%s = %s; %% missing is the share of those cells with no reported value.",
        label_comma()(n_periods), env_time_type,
        if (n_strata_cells > 1) sprintf(" x %s stratum cell(s)", label_comma()(n_strata_cells)) else "",
        label_comma()(n_periods * n_strata_cells)
    ),
    rownames = FALSE, options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
) %>% DT::formatRound(columns = if (multi_signal) 6 else 5, digits = 2)
```

### Time Series Overview

This section provides a visual overview of the candidate `r ind` over time, one location at a time (use the dropdown to switch). Values are shown as published: the line breaks wherever a period has no reported value, and each marker is one reported observation, so gaps in coverage read as gaps rather than as interpolated straight lines.

```{r}
#| label: eda-overview
#| fig-height: 4
#| results: asis
# Pick a diverse sample of locations for visualization
all_geos <- sort(unique(df_candidate$geo_value))
n_plot <- min(length(all_geos), params$max_locations_plot)
sample_geos <- random_sample(all_geos, n_plot)

# Confidence bounds are plotted when the source supplies them.
ci_cols <- intersect(c("ci_lower", "ci_upper"), colnames(df_candidate))
if (length(ci_cols) < 2) ci_cols <- NULL

df_sample <- df_candidate %>%
    as_tibble() %>%
    filter(geo_value %in% sample_geos) %>%
    # Fill the grid so that missing periods become explicit NAs; plotly then
    # leaves a gap instead of joining across them.
    complete(nesting(signal, time_value), !!!rlang::syms(params$extra_keys), geo_value) %>%
    mutate(indicator = signal_labels[signal])

if (!is.null(ci_cols)) {
    cat(sprintf(
        "Shaded bands are the confidence interval reported with each value%s.\n\n",
        if (length(params$extra_keys) > 0) {
            sprintf(", drawn in the colour of their %s and sharing its legend entry, so clicking a %s in the legend hides its line and its band together",
                    params$extra_keys[1], params$extra_keys[1])
        } else ""
    ))
}

# One plot per indicator so the dropdown only ever switches location.
band_notes <- c()
overview_plots <- map(signals_vec, function(sig) {
    d <- df_sample %>% filter(signal == sig)
    p <- plot_series_dropdown(
        d,
        # ponytail: one trace per level of the first stratifier only; nest the
        # rest into a combined key here if a second stratifier ever needs panels.
        trace_col = if (length(params$extra_keys) > 0) params$extra_keys[1] else NULL,
        ci_cols = ci_cols
    )
    if (is.null(p)) return(NULL)
    if (!is.null(attr(p, "bands_wide")) && attr(p, "bands_wide") > 0) {
        band_notes[[sig]] <<- sprintf(
            "%s%s of %s reported values (%s) carry an interval wider than the full range of the values themselves, so no band is drawn for them",
            if (multi_signal) sprintf("%s: ", signal_labels[sig]) else "",
            label_comma()(attr(p, "bands_wide")), label_comma()(attr(p, "bands_total")),
            label_percent(accuracy = 0.1)(attr(p, "bands_wide") / attr(p, "bands_total"))
        )
    }
    # Kept as SVG (no toWebGL): WebGL cannot fill a ribbon polygon, so it would
    # break the bands. Only one location is visible at a time, so the rendered
    # trace count is small and SVG is fast enough.
    p <- p %>% plotly::partial_bundle()
    if (!multi_signal) return(p)
    htmltools::tagList(htmltools::tags$h5(signal_labels[sig]), p)
}) %>% compact()

if (length(band_notes) > 0) {
    cat(sprintf(
        "An estimate resting on almost no denominator has a confidence interval about as wide as the indicator itself; drawn, it is a full-height column that hides every other series. %s. Those observations still show as a line and a marker, and the y axis is scaled to the values rather than to the bands, so a band running off the top of the plot is clipped there — zoom out to see it.\n\n",
        paste(band_notes, collapse = "; ")
    ))
}

if (multi_signal) htmltools::tagList(overview_plots) else overview_plots[[1]]
```

::: {.content-visible when-meta="multi_location"}

### Quantile Trends

`r sg_pl("This plot displays", "These plots display")` the distribution of the `r ind` over time across all included locations. The solid line represents the median, the inner ribbon shows the 25th to 75th percentile range (the middle 50% of locations), and the outer ribbon covers the 5th to 95th percentile range. We use a different scale for each panel so each distribution is clearly visible.

`r if (smooth_quantiles) "The quantiles are smoothed with a moving average. That is aimed at raw data with a day-of-week effect, which would otherwise dominate the ribbons; an indicator already smoothed at the source does not need it, and smoothing it again only widens the effective window. Set the <code>smooth_quantiles</code> parameter to <code>false</code> to plot the quantiles as published." else "Quantiles are plotted as published, without additional smoothing (<code>smooth_quantiles: false</code>). Set the parameter to <code>true</code> for raw data where a day-of-week effect would otherwise dominate the ribbons."`

By observing the width of these ribbons, you can get a sense of geographic variability. Narrow ribbons mean the values are tightly clustered across locations, while wide ribbons show more spread. This helps us assess the overall stability and systemic trends of the `r ind`.

Each quantile is taken cross whichever locations reported a value in that window, so panels whose location counts differ (shown in each panel label) are not directly comparable. A spike confined to one panel is often a composition effect: a location that reports only the aggregate stratum, on a small denominator, can move that panel's upper quantile without appearing in any other panel.

```{r}
#| label: eda-quantiles
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 1.5), 15)
#| eval: !expr n_locations > 1
cat("\n")

# Aggregate location values per (signal, time_value); use signal name as geo_value
# so epi_slide partitions by signal automatically.
df_quantiles_wide <- df_candidate %>%
    group_by(across(all_of(c("signal", "time_value", params$extra_keys)))) %>%
    summarize(
        vals = list(value),
        geo_value = "all",
        .groups = "drop"
    ) %>%
    as_epi_df(signal_format = "as_is", other_keys = c("signal", params$extra_keys))

# Handle time frequency differences for .window_size.
window_size <- switch(env_time_type %||% "unknown",
    day   = 7L, # 7-day smoothing
    week  = as.difftime(3, units = "weeks"), # 3-week smoothing
    month = as.difftime(3, units = "months"), # 3-month smoothing
    year  = as.difftime(3, units = "years"), # 3-year smoothing
    Inf # Default to cumulative for unknown types
    # (too sparse for sliding averages)
)

q_tbl <- function(v) {
    tibble(
        `0.05` = quantile(v, 0.05, na.rm = TRUE),
        `0.25` = quantile(v, 0.25, na.rm = TRUE),
        `0.5`  = quantile(v, 0.5, na.rm = TRUE),
        `0.75` = quantile(v, 0.75, na.rm = TRUE),
        `0.95` = quantile(v, 0.95, na.rm = TRUE)
    )
}

df_quantiles_wide <- {
    if (smooth_quantiles) {
        df_quantiles_wide %>%
            group_by(across(all_of(c("geo_value", "signal", params$extra_keys)))) %>%
            epi_slide(~ q_tbl(unlist(.x$vals)), .window_size = window_size) %>%
            ungroup()
    } else {
        # Quantiles of the values published in that period alone.
        bind_cols(df_quantiles_wide, map_dfr(df_quantiles_wide$vals, q_tbl))
    }
} %>%
    as_tibble() %>%
    mutate(indicator = signal_labels[signal]) %>%
    label_strata_coverage(params$extra_keys)

ggplot(df_quantiles_wide, aes(x = time_value)) +
    geom_ribbon(aes(ymin = `0.05`, ymax = `0.95`, fill = "5%-95%"), alpha = 0.2) +
    geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`, fill = "25%-75%"), alpha = 0.4) +
    geom_line(aes(y = `0.5`, color = "Median"), linewidth = 1) +
    facet_wrap(reformulate(c("indicator", params$extra_keys)), ncol = 1, scales = "free_y") +
    scale_x_date(breaks = breaks_pretty(), label = label_date_short()) +
    scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_si(""))) +
    scale_fill_manual(values = c("5%-95%" = dis_pal_list[2], "25%-75%" = dis_pal_list[2])) +
    scale_color_manual(values = c("Median" = dis_pal_list[2])) +
    labs(
        title = if (smooth_quantiles) "Smoothed Quantile Trends" else "Quantile Trends",
        subtitle = sprintf("Across the locations reporting each panel (of %d in total)%s",
                           n_locations,
                           if (smooth_quantiles) sprintf(", smoothed window: %s", format(window_size)) else ", as published"),
        x = "Date", y = if (smooth_quantiles) "Value (smoothed)" else "Value",
        fill = "Range", color = "Trend"
    ) +
    theme(legend.position = "bottom")
```

::: {.content-visible when-meta="geo_outliers"}

#### Cross-Sectional Geographic Outliers

While the quantile plots summarize the main body of the distribution, they may hide the behavior of individual highly erratic locations. This table complements them by placing each location within the distribution: it tracks how often a location sits in either extreme tail — above the cross-sectional 95th percentile, or below the 5th percentile, recorded on that specific date and stratum. Both percentages are shares of the periods the location actually reported, not of the analysis window, and the number of reported values is shown so that a location with only a handful of observations is not read as a stable 100%.

How to read it: a location in the tail on roughly 5% of its periods is where chance alone would put it, so the table is about the ones far from 5%. A location above the 95th percentile most of the time is not having a long outbreak — it is sitting at a different level from everyone else, which usually points at its denominator, coverage or coding rather than at epidemiology, and is worth checking against the raw counts. The same holds at the bottom, where a high zero share is the more common symptom. A location that appears in *both* tails is unstable rather than merely high or low.

::: {.callout-note}
This approach compares locations against one another on each date, so it only tells you about relative position — never about whether a value is correct. If the `r ind` `r sg_pl("relies", "rely")` on raw counts rather than population-standardized rates or percentages, this table will frequently just identify locations with the largest populations; in those cases, this metric reflects size rather than outliers. Set the `show_geo_outliers` parameter to `false` to drop the section.
:::

```{r}
#| label: eda-aberrant-locations
#| echo: false
#| results: asis
#| eval: !expr (show_geo_outliers && n_locations > 1)
strat_cols <- c("signal", params$extra_keys)

# Only reported values take part: a location cannot be in a tail on a period it
# never reported, and it should not be charged for one either.
df_reported <- df_candidate %>%
    as_tibble() %>%
    filter(!is.na(value))

# A non-negative indicator makes the share of exact zeros worth reporting
# alongside the lower tail, since the 5th percentile is often zero itself.
show_zero_share <- all(df_reported$value >= 0)

# Cross-sectional tails per (indicator, stratum, period)
daily_tails <- df_reported %>%
    group_by(across(all_of(c(strat_cols, "time_value")))) %>%
    summarize(
        cand_p05 = quantile(value, 0.05),
        cand_p95 = quantile(value, 0.95),
        .groups = "drop"
    )

aberrant_locs <- df_reported %>%
    left_join(daily_tails, by = c(strat_cols, "time_value")) %>%
    group_by(across(all_of(c(strat_cols, "geo_value")))) %>%
    summarize(
        n_obs = n(),
        upper_pct = 100 * mean(value > cand_p95),
        lower_pct = 100 * mean(value < cand_p05),
        zero_pct  = 100 * mean(value == 0),
        .groups = "drop"
    ) %>%
    mutate(
        geo_name     = geo_to_name(geo_value),
        signal_label = signal_labels[signal]
    ) %>%
    arrange(desc(upper_pct))

if (nrow(aberrant_locs) > 0) {
    aberrant_display <- aberrant_locs %>%
        select(
            any_of(if (multi_signal) "signal_label" else character(0)),
            all_of(params$extra_keys), geo_value, geo_name, n_obs,
            upper_pct, lower_pct, any_of(if (show_zero_share) "zero_pct" else character(0))
        )
    DT::datatable(
        aberrant_display,
        colnames = c(
            if (multi_signal) "Indicator" else NULL,
            str_to_title(str_replace_all(params$extra_keys, "_", " ")),
            "Geo-level",
            "Name",
            "Values Reported",
            "% Above 95th",
            "% Below 5th",
            if (show_zero_share) "% Exactly Zero" else NULL
        ),
        caption = "Summary of aberrant locations (time spent in each tail, as a share of the periods the location reported).",
        rownames = FALSE, options = list(scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
    ) %>% DT::formatRound(
        columns = which(names(aberrant_display) %in% c("upper_pct", "lower_pct", "zero_pct")),
        digits = 1
    )
} else {
    cat("Insufficient data to calculate aberrant time for these locations.\n\n")
}
```

:::


:::


::: {.content-visible when-meta="multi_location"}

### Temporal Variability

We calculate the Coefficient of Variation (CV) for each location to compare how much the `r ind` `r sg_pl("moves", "move")` over time across geographic areas. A high CV is not by itself evidence of noise: real epidemic dynamics, seasonality and genuine level shifts all raise it, so read it as the size of the temporal variation and use the seasonality and revision sections to judge how much of that variation is signal. What it is good for is comparison — a location whose CV is far above its peers is usually the one to inspect first.

The CV is computed over the values a location actually reported, so missing periods neither contribute nor count against it. One CV is computed per (location, stratum) rather than pooling strata together, because a location's strata usually sit at different levels and pooling them measures between-stratum spread rather than variation over time.

A CV is undefined where a location has a mean of zero or fewer than two reported values; those series cannot appear in the histogram, and their number is stated in the caption so the bars can be reconciled against the location count.

$$\text{CV}_i = \frac{\text{SD}(X_i)}{|\bar{X}_i|}$$

```{r}
#| label: eda-noise
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 0.8), 12)
#| eval: !expr n_locations > 1
# One CV per (indicator, stratum, location), over reported values only
noise_stats <- df_candidate %>%
    as_tibble() %>%
    filter(!is.na(value)) %>%
    group_by(across(all_of(c("signal", params$extra_keys, "geo_value")))) %>%
    summarize(
        n_obs = n(),
        cand_cv = sd(value) / abs(mean(value)),
        .groups = "drop"
    )

n_series <- nrow(noise_stats)
noise_undefined <- noise_stats %>% filter(!is.finite(cand_cv))
noise_stats <- noise_stats %>% filter(is.finite(cand_cv))

# Clip extreme outliers for robust visualization
cv_upper <- quantile(noise_stats$cand_cv, 0.99, na.rm = TRUE)
noise_stats_plot <- noise_stats %>%
    filter(cand_cv <= cv_upper) %>%
    mutate(signal_label = signal_labels[signal]) %>%
    label_strata_coverage(params$extra_keys)

n_excluded <- nrow(noise_stats) - nrow(noise_stats_plot)

# Summary Table — one row per (indicator, stratum)
noise_summary <- noise_stats %>%
    group_by(across(all_of(c("signal", params$extra_keys)))) %>%
    summarize(median_cv = median(cand_cv), n_locs = n_distinct(geo_value), .groups = "drop") %>%
    mutate(
        indicator = signal_labels[signal],
        across(all_of(params$extra_keys), order_strata)
    ) %>%
    arrange(across(all_of(params$extra_keys))) %>%
    select(indicator, all_of(params$extra_keys), n_locs, median_cv)

DT::datatable(noise_summary,
    colnames = c(
        "Indicator",
        str_to_title(str_replace_all(params$extra_keys, "_", " ")),
        "Locations", "Median CV"
    ),
    caption = sprintf(
        "CV summary across locations. %d of %d (location%s) series have a defined CV%s%s.",
        nrow(noise_stats), n_series,
        if (length(params$extra_keys) > 0) " x stratum" else "",
        if (nrow(noise_undefined) > 0) {
            sprintf("; %d dropped as undefined (zero mean or a single reported value)", nrow(noise_undefined))
        } else "",
        if (n_excluded > 0) sprintf("; a further %d above the 99th percentile are excluded from the plot", n_excluded) else ""
    ),
    rownames = FALSE, options = list(scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
) %>% DT::formatRound(columns = ncol(noise_summary), digits = 3)

# Histogram (faceted by indicator x stratum)
ggplot(noise_stats_plot, aes(x = cand_cv)) +
    geom_histogram(bins = 30, color = "white", fill = dis_pal_list[2], alpha = 0.7) +
    scale_x_continuous(labels = label_percent()) +
    {if (multi_signal || length(params$extra_keys) > 0) facet_wrap(reformulate(c(if (multi_signal) "signal_label", params$extra_keys)), scales = "free", labeller = label_wrap_html(width = 25)) else NULL} +
    labs(
        title = "CV Distribution by Location",
        subtitle = "Higher values mean more variation over time, from real dynamics or from noise",
        x = "Coefficient of Variation (CV)", y = "Count"
    ) +
    theme(legend.position = "none")
```





:::

```{r}
#| label: revision-analysis-setup
#| echo: false
#| dependson: "load-candidate-archive"
# Get versioned data
archive <- NULL
archive_has_history <- FALSE
if (!is.null(candidate_archive) &&
    n_distinct(candidate_archive$DT$version) > 1) {
    archive <- candidate_archive
    archive_span_days <- as.numeric(
        max(archive$DT$version) - min(archive$DT$version),
        units = "days"
    )
    archive_has_history <- archive_span_days >= params$min_archive_days
}
```





::: {.content-visible when-meta="versioning"}

### Candidate Versioning

Versioning is not the focus of this notebook. This section answers only whether the `r ind` `r sg_pl("revises", "revise")` at all and roughly how much, which is enough to decide whether the values above can be read as final. The statistics below summarize how often and how much observations are revised after their initial release. Relative revisions measure the magnitude of a revision relative to the `r ind_poss` own scale: a relative revision of 0.1 means the observation shifted by 10% of its own magnitude across revisions, while values near 1 indicate revisions as large as the `r ind` `r sg_pl("itself", "themselves")`.

::: {.callout-tip}
For the full treatment — revision lags by location, convergence over time, and how much of the latest value is knowable at each lag — use the dedicated notebook, `revision_analysis/revision_analysis.qmd`, which exists for that purpose.
:::

```{r}
#| label: revision-analysis-text
#| echo: false
#| results: asis
#| eval: !expr show_versioning
if (is.null(archive)) {
    cat(sprintf(
        "No version history is available for `%s / %s` — versioning analysis is skipped.\n\n",
        params$source, paste(signals_vec, collapse = ", ")
    ))
} else if (!archive_has_history) {
    version_start <- min(archive$DT$version)
    cat(sprintf(
        "Versioning began on %s. At least 60 days of version history are needed for meaningful revision analysis; the earliest this will be possible is %s.\n\n",
        format(version_start), format(version_start + 60)
    ))
    cat(sprintf(
        "Current versions: %s.\n\n",
        paste(sort(unique(archive$DT$version)), collapse = ", ")
    ))
} else {
    cat("> The `revision_summary` is under revision and may produce misleading results in certain edge cases. Use the summary statistics below with caution while development is ongoing.\n\n")

    if (exists("archive_info") && !is.null(archive_info)) {
        cat(sprintf(
            "Revision analysis is performed on a sample of %d out of %d locations, with versions from %s to %s, and time from %s to %s.\n\n",
            archive_info$n_sampled, archive_info$n_total, archive_info$version_start, archive_info$version_end, archive_info$time_value_start, archive_info$time_value_end
        ))
    }

    # Helper: print revision stats block for one signal's rev_details
    print_rev_stats <- function(rd, archive_obj) {
        rb <- rd$revision_behavior
        cat(sprintf(
            "- **Time range**: %s to %s\n",
            rd$range_time_values[1], rd$range_time_values[2]
        ))
        lag_stats <- summary(as.numeric(rb$min_lag, units = "days"))
        cat("- **Min lag (days to first version)**: ")
        cat(paste(names(lag_stats), round(as.numeric(lag_stats), 1), sep = ": ", collapse = ", "), "\n")
        quick_threshold <- 3
        few_threshold <- 3
        rel_spread_threshold <- 0.1
        abs_spread_threshold <- 0.05 * rd$max_val
        no_rev_pct   <- mean(rb$n_revisions == 0)
        quick_rev_pct <- mean(as.numeric(rb$max_lag, units = "days") <= quick_threshold)
        few_rev_pct  <- mean(rb$n_revisions <= few_threshold)
        cat(sprintf("- **Total number of versions**: %d\n", n_distinct(archive_obj$DT$version)))
        cat(sprintf("- **No revisions**: %s\n", label_percent()(no_rev_pct)))
        cat(sprintf("- **Quick revisions (<= %d days)**: %s\n", quick_threshold, label_percent()(quick_rev_pct)))
        cat(sprintf("- **Few revisions (<= %d)**: %s\n", few_threshold, label_percent()(few_rev_pct)))
        revised_beh <- rb %>% filter(n_revisions > 0)
        if (nrow(revised_beh) > 0) {
            low_rel_spread_pct  <- mean(revised_beh$rel_spread < rel_spread_threshold, na.rm = TRUE)
            high_abs_spread_pct <- mean(revised_beh$spread > abs_spread_threshold, na.rm = TRUE)
            cat(sprintf("- **Minor relative revisions (< %.1f)**: %s\n",
                        rel_spread_threshold, label_percent()(low_rel_spread_pct)))
            cat(sprintf("- **Major absolute revisions (> %.3f)**: %s\n",
                        abs_spread_threshold, label_percent()(high_abs_spread_pct)))
        }
        cat(sprintf("- **Days until within %d%% of latest**: ", rd$within_latest * 100))
        conv_stats <- summary(as.numeric(rb$lag_near_latest, units = "days"))
        cat(paste(names(conv_stats), round(as.numeric(conv_stats), 1), sep = ": ", collapse = ", "), "\n\n")
    }

    # Run revision_summary per signal so results are always signal-specific
    for (sig in signals_vec) {
        if (multi_signal) cat(sprintf("\n#### %s\n\n", signal_labels[sig]))

        # Subset archive to this signal (drop signal col for revision_summary)
        sub_archive <- archive$DT %>%
            as_tibble() %>%
            filter(signal == sig) %>%
            select(geo_value, time_value, version, value) %>%
            as_epi_archive()

        rd_sig <- revision_summary(sub_archive)
        print_rev_stats(rd_sig, sub_archive)
    }

}
```

:::



### Indicator sanity checks

This section is by its nature more difficult than others to do systematically. Things to look for:

- Do the values you see match the values you expected to see? This is the check that catches the most, and only you can make it: state the range, level and shape you expected from the `r ind_poss` definition before reading the plots, then compare.
- If it's a percentage or percentile, is it between 0 and 100?
- If it's a count, is it roughly proportional to the population of each location? A count that does not scale with population points at a coverage or denominator problem. A rate or percentage is already normalized, so the level should instead be roughly comparable across locations of very different sizes, and a location whose level is far off the rest is worth investigating. Note that the relevant denominator is usually not the whole resident population but the part of it the source observes (for example, the patients of one healthcare system). *(requires external population data — inspect manually)*
- Are the actual values reasonable? E.g. if it's a percentage, does it ever approach 100 outside of literal pandemics?
- If it's for a seasonal disease, are rates highest in the winter and lowest in the summer?
- Are geo-aggregated coarser levels plausible aggregations? *(checked below for state-level indicators vs HHS/nation)*

The `r sg_pl("histogram", "histograms")` and table below show the distribution of all observed values and the value range per location for the candidate `r ind`, to assess whether the range is plausible given the `r ind_poss` expected scale.

```{r}
#| label: sanity-value-range
#| fig-height: !expr min(max(5, length(signals_vec) * 0.5), 12)
df_candidate_plot <- df_candidate %>%
    mutate(indicator = signal_labels[signal])

ggplot(df_candidate_plot, aes(x = value)) +
    geom_histogram(bins = 40, alpha = 0.8, color = "white") +
    scale_x_continuous(labels = label_number(scale_cut = cut_si(""))) +
    scale_y_continuous(labels = label_number(scale_cut = cut_si(""))) +
    {if (multi_signal) facet_wrap(~indicator, scales = "free", labeller = label_wrap_html(width = 25)) else NULL} +
    labs(
        title = if (multi_signal) "Value distribution by indicator"
                else sprintf("Value distribution: %s", signal_labels[signals_vec]),
        subtitle = sprintf(
            "%s reported values across %d locations",
            label_comma()(nrow(df_candidate)), n_locations
        ),
        x = "Value", y = "Count"
    )
```

```{r}
#| label: sanity-value-range-table
#| echo: false
range_tbl <- df_candidate %>%
    as_tibble() %>%
    filter(!is.na(value)) %>%
    group_by(across(all_of(c("signal", params$extra_keys, "geo_value")))) %>%
    summarize(
        n = n(),
        min = min(value),
        p05 = quantile(value, 0.05),
        median = median(value),
        p95 = quantile(value, 0.95),
        max = max(value),
        .groups = "drop"
    ) %>%
    mutate(signal_label = signal_labels[signal]) %>%
    arrange(desc(max)) %>%
    select(
        any_of(if (multi_signal) "signal_label" else character(0)),
        all_of(params$extra_keys), geo_value, n:max
    )

DT::datatable(
    range_tbl,
    colnames = c(
        if (multi_signal) "Indicator" else NULL,
        str_to_title(str_replace_all(params$extra_keys, "_", " ")),
        "Location", "N", "Min", "5th pct", "Median", "95th pct", "Max"
    ),
    caption = sprintf(
        "Value range per location%s. Sort by Min or Max to surface potential out-of-bounds values.",
        if (multi_signal) "" else sprintf(" for %s", signal_labels[signals_vec])
    ),
    rownames = FALSE,
    options = list(
        scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
        paging = TRUE, dom = "tp"
    )
) %>% DT::formatRound(columns = which(names(range_tbl) %in% c("min", "p05", "median", "p95", "max")), digits = 3)
```

The plot below shows the seasonal pattern of the `r ind`, with each panel representing one epidemiological year (`r season_label`). That boundary is set by the `season_start_month` parameter; it defaults to July so that a northern-hemisphere respiratory season falls inside a single panel, following the usual influenza convention. Set it to another month for indicators whose season peaks elsewhere in the year. The line traces the monthly median across all locations, the inner band covers the 25th-75th percentile range (the middle 50% of observations that month), and the outer band spans the full observed range, from the lowest to the highest value.

```{r}
#| label: sanity-seasonality
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 2.5), 15)
#| eval: !expr as.numeric(difftime(max(df_candidate$time_value), min(df_candidate$time_value), units = "days")) >= 365
df_seasonal <- df_candidate %>%
    as_tibble() %>%
    add_season() %>%
    group_by(signal, season, season_month, across(all_of(params$extra_keys))) %>%
    summarize(
        median_val = median(value, na.rm = TRUE),
        min_val = min(value, na.rm = TRUE),
        max_val = max(value, na.rm = TRUE),
        p25 = quantile(value, 0.25, na.rm = TRUE),
        p75 = quantile(value, 0.75, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    # Keep only signal-season (x extra-key) combos with >= 3 months of data
    semi_join(
        count(., across(all_of(c("signal", "season", params$extra_keys)))) %>% filter(n >= 3),
        by = c("signal", "season", params$extra_keys)
    ) %>%
    mutate(signal_label = signal_labels[signal]) %>%
    label_strata_coverage(params$extra_keys)

ggplot(df_seasonal, aes(x = season_month, color = season, fill = season, group = season)) +
    geom_ribbon(aes(ymin = min_val, ymax = max_val), alpha = 0.10, linewidth = 0.2) +
    geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.30, linewidth = 0.3) +
    geom_line(aes(y = median_val), linewidth = 0.8) +
    geom_point(aes(y = median_val), size = 1.5) +
    {
        row_vars <- c(if (multi_signal) "signal_label", params$extra_keys)
        if (length(row_vars) > 0) {
            facet_grid(as.formula(paste(paste(row_vars, collapse = " + "), "~ season")), scales = "free_y")
        } else {
            facet_wrap(~season, ncol = 2, scales = "free_y")
        }
    } +
    scale_x_continuous(labels = season_month_labels) +
    scale_y_continuous(labels = label_number(scale_cut = cut_si(""))) +
    labs(
        title = if (multi_signal) "Seasonal patterns by indicator"
                else sprintf("Seasonal pattern: %s", signal_labels[signals_vec]),
        subtitle = "Median by month across all locations — inner band: 25th-75th pct, outer band: min-max",
        x = "Month", y = "Median value"
    ) +
    theme(legend.position = "none")
```

```{r}
#| label: sanity-geo-aggregation
#| echo: false
#| results: asis
#| eval: !expr params$geo_type == "state"
pacman::p_load(nanoparquet)

crosswalk_path <- here("tests", "test_data", "delphi_geomap", "geomapper_crosswalk_20200101.parquet")
crosswalk <- tryCatch(
    {
        nanoparquet::read_parquet(crosswalk_path) %>%
            distinct(state_geo = state_abbr, hhs_region = as.character(hhs))
    },
    error = function(e) {
        # Fallback to standard HHS region mapping if the file is missing
        tibble(
            state_geo = c(
                "CT", "ME", "MA", "NH", "RI", "VT",
                "NJ", "NY", "PR", "VI",
                "DE", "DC", "MD", "PA", "VA", "WV",
                "AL", "FL", "GA", "KY", "MS", "NC", "SC", "TN",
                "IL", "IN", "MI", "MN", "OH", "WI",
                "AR", "LA", "NM", "OK", "TX",
                "IA", "KS", "MO", "NE",
                "CO", "MT", "ND", "SD", "UT", "WY",
                "AZ", "CA", "HI", "NV", "AS", "GU", "MP",
                "AK", "ID", "OR", "WA"
            ),
            hhs_region = c(
                rep("1", 6),
                rep("2", 4),
                rep("3", 6),
                rep("4", 8),
                rep("5", 6),
                rep("6", 5),
                rep("7", 4),
                rep("8", 6),
                rep("9", 7),
                rep("10", 4)
            )
        )
    }
)

is_pct_indicator <- any(grepl("%|pct", name_vec, ignore.case = TRUE))

fetch_coarser_candidate <- function(geo_type_coarser) {
    tryCatch(
        # pub_covidcast accepts a vector of signals
        pub_covidcast(
            params$source, signals_vec,
            geo_type_coarser, params$time_type,
            time_values = time_range(params$start_day, params$end_day),
            as_of = Sys.Date()
        ) %>% select(time_value, signal, geo_value, coarser_val = value),
        error = function(e) NULL
    )
}

state_vals <- df_candidate %>%
    as_tibble() %>%
    select(signal, time_value, state_geo = geo_value, state_val = value)

check_aggregation <- function(df_coarser, group_label, state_to_coarser) {
    if (is.null(df_coarser)) {
        return(NULL)
    }

    state_bounds <- state_vals %>%
        inner_join(state_to_coarser, by = "state_geo") %>%
        group_by(signal, time_value, coarser_geo) %>%
        summarize(
            state_min = min(state_val, na.rm = TRUE),
            state_max = max(state_val, na.rm = TRUE),
            n_states  = n(),
            .groups   = "drop"
        )

    df_coarser %>%
        rename(coarser_geo = geo_value) %>%
        inner_join(state_bounds, by = c("signal", "time_value", "coarser_geo")) %>%
        filter(!is.na(coarser_val) & (coarser_val < state_min - 1e-9 | coarser_val > state_max + 1e-9)) %>%
        mutate(
            geo_level    = group_label,
            signal_label = signal_labels[signal],
            direction    = if_else(coarser_val < state_min, "below min", "above max")
        ) %>%
        select(geo_level, signal_label, time_value, coarser_geo, coarser_val, state_min, state_max, direction, n_states)
}

df_hhs_raw <- fetch_coarser_candidate("hhs")
df_nation_raw <- fetch_coarser_candidate("nation")

hhs_map <- crosswalk %>% select(state_geo, coarser_geo = hhs_region)
nation_map <- crosswalk %>%
    distinct(state_geo) %>%
    mutate(coarser_geo = "us")

violations <- bind_rows(
    check_aggregation(df_hhs_raw, "HHS", hhs_map),
    check_aggregation(df_nation_raw, "Nation", nation_map)
)

unavailable <- c(
    if (is.null(df_hhs_raw)) "HHS" else NULL,
    if (is.null(df_nation_raw)) "Nation" else NULL
)

if (length(unavailable) > 0) {
    cat(sprintf(
        "**Note**: %s-level data not available for %s / %s.\n\n",
        paste(unavailable, collapse = " and "),
        params$source, paste(signals_vec, collapse = ", ")
    ))
}

if (nrow(violations) == 0) {
    cat(sprintf(
        "No geo-aggregation violations found: all HHS/nation values are within the [min, max] of constituent state values for %s.\n",
        paste(name_vec, collapse = ", ")
    ))
} else {
    cat(sprintf(
        "**%d observations** where the aggregated value falls outside the [min, max] of its constituent states (%s indicator):\n\n",
        nrow(violations),
        if (is_pct_indicator) "percentage" else "non-percentage"
    ))
    print(htmltools::tagList(
        DT::datatable(
            violations,
            colnames = c("Level", "Indicator", "Date", "Geo", "Agg Value", "State Min", "State Max", "Direction", "N States"),
            caption = sprintf(
                "Geo-aggregation bounds check for %s. Aggregated values should fall within [min, max] of constituent state values.",
                paste(name_vec, collapse = ", ")
            ),
            rownames = FALSE,
            options = list(
                scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
                paging = TRUE, pageLength = 20, dom = "tp"
            )
        ) %>% DT::formatRound(columns = 5:7, digits = 3)
    ))
}
```





The table below flags locations whose seasonal peak is at least twice the typical (median) peak across all locations in the same season.

```{r}
#| label: sanity-seasonality-anomalies
#| echo: false
#| eval: !expr as.numeric(difftime(max(df_candidate$time_value), min(df_candidate$time_value), units = "days")) >= 365
df_seasonal_loc <- df_candidate %>%
    as_tibble() %>%
    add_season() %>%
    group_by(across(all_of(c("signal", params$extra_keys, "season", "geo_value")))) %>%
    summarize(loc_max = max(value, na.rm = TRUE), .groups = "drop")

typical_max <- df_seasonal_loc %>%
    group_by(across(all_of(c("signal", params$extra_keys, "season")))) %>%
    summarize(typical_max = median(loc_max, na.rm = TRUE), .groups = "drop")

anomalies <- df_seasonal_loc %>%
    left_join(typical_max, by = c("signal", params$extra_keys, "season")) %>%
    mutate(
        ratio        = loc_max / typical_max,
        signal_label = signal_labels[signal]
    ) %>%
    filter(ratio >= 2) %>%
    arrange(desc(ratio)) %>%
    select(
        any_of(if (multi_signal) "signal_label" else character(0)),
        all_of(params$extra_keys), season, geo_value, loc_max, typical_max, ratio
    )

if (nrow(anomalies) == 0) {
    cat("No locations found with a seasonal max more than 2x the typical (median) seasonal max.")
} else {
    DT::datatable(
        anomalies,
        colnames = c(
            if (multi_signal) "Indicator" else NULL,
            str_to_title(str_replace_all(params$extra_keys, "_", " ")),
            "Season", "Location", "Location Max", "Typical Max", "Ratio"
        ),
        caption = sprintf(
            "Locations where seasonal max exceeds 2x the typical (median) seasonal max%s.",
            if (multi_signal) "" else sprintf(" for %s", signal_labels[signals_vec])
        ),
        rownames = FALSE,
        options = list(
            scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
            paging = TRUE, pageLength = 20, dom = "tp"
        )
    ) %>% DT::formatRound(columns = which(names(anomalies) %in% c("loc_max", "typical_max", "ratio")), digits = 3)
}
```