---
title: "Indicator Exploratory Analysis: `r paste(if (!is.null(params$name)) params$name else params$signal, collapse = ', ')`"
author: "Delphi Group"
date: today
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
params:
  # Candidate Indicator
  source:
  signal:
  name:
  input_csv: !expr "NULL"
  # Shared settings
  geo_type:
  time_type:
  start_day:
  end_day:
  # Extra key columns beyond geo_value/time_value (character vector), e.g. "age_group"
  extra_keys: !expr "character(0)"
  # First month of the epidemiological year used by the seasonality panels.
  # 7 = July-June, the usual influenza convention.
  season_start_month: 7
  # Smooth the quantile trends with a moving average. Meant for raw data with a
  # day-of-week effect; set to false for indicators already smoothed at source.
  smooth_quantiles: true
  # Show the cross-sectional geographic outlier table (2.2.1)
  show_geo_outliers: true
  # Show the versioning summary. Off also skips the version-history query, which
  # is the slowest part of the report; use revision_analysis.qmd for the detail.
  show_versioning: true
  # Maximum locations to show in plots/tables
  max_locations_plot: 60
  max_archive_locs: 60
  min_archive_days: 60  # Minimum days of version history to enable revision analysis
  # A confidence band wider than (this multiplier x the value range) is dropped
  # from the location dropdown plot (2.2.2) as a near-empty-denominator outlier.
  # Raise it (e.g. Inf) to draw every reported band regardless of width.
  ci_band_limit_mult: 1
  # Is the indicator's level comparable across locations (a rate, percentage or
  # per-capita value)?
  comparable_across_locations: !expr "NULL"
  # What kind of value the indicator carries. One of: "cumulative", "count", "rate", "percent", "proportion", "other".
  value_type: !expr "NULL"
knitr:
  opts_chunk:
    comment: "#>"
    message: FALSE
    warning: FALSE
    out.width: "100%"
    dev: "png"
    fig.width: 8
    fig.height: 4.5
    fig.align: center
editor_options: 
  chunk_output_type: console
execute:
  cache: false
---

```{r}
#| label: setup
#| include: false
#| cache: false
# When running interactively, populate params from defaults
if (!exists("params")) {
    params <- list(
        source = "beta_pophive",
        signal = "epic_pct_covid_total",
        name = "PopHive: Epic % COVID Total",
        input_csv = NULL,
        geo_type = "state",
        time_type = NULL,
        end_day = format(Sys.Date()),
        start_day = format(Sys.Date() - 365 * 2),
        extra_keys = character(0),
        season_start_month = 7,
        smooth_quantiles = TRUE,
        show_geo_outliers = TRUE,
        show_versioning = TRUE,
        max_locations_plot = 60,
        max_archive_locs = 60,
        min_archive_days = 60,
        ci_band_limit_mult = 1,
        comparable_across_locations = NULL,
        value_type = NULL
    )
}

# Runtime defaults for dynamic params. Unlock to allow modification.
if (exists("params", envir = .GlobalEnv, inherits = FALSE) &&
    bindingIsLocked("params", .GlobalEnv)) {
    unlockBinding("params", .GlobalEnv)
}
# Helper to flatten parameter lists (e.g. when Quarto CLI parses colons as nested lists)
flatten_param <- function(x) {
    if (is.null(x)) return(NULL)
    if (is.list(x)) {
        if (!is.null(names(x)) && any(nzchar(names(x)))) {
            paste(names(x), unlist(x), sep = ": ")
        } else {
            as.character(unlist(x))
        }
    } else {
        as.character(x)
    }
}

params$source <- flatten_param(params$source)
params$signal <- flatten_param(params$signal)
params$name <- flatten_param(params$name)
params$geo_type <- flatten_param(params$geo_type)
params$time_type <- flatten_param(params$time_type)
params$input_csv <- flatten_param(params$input_csv)

# The CSV path is relative to whichever directory quarto renders from, so try it
# as given, with the current directory's name stripped, and one level up.
if (!is.null(params$input_csv) && nzchar(params$input_csv)) {
    candidates <- c(
        params$input_csv,
        sub(paste0("^", basename(getwd()), "/"), "", params$input_csv),
        file.path("..", params$input_csv)
    )
    found <- candidates[file.exists(candidates)]
    if (length(found) == 0) {
        stop(sprintf("Local CSV file not found at path: %s", params$input_csv))
    }
    params$input_csv <- found[1]
}

# When using CSV route, source/indicator labels default to the CSV filename
if (!is.null(params$input_csv) &&
    (is.null(params$source) || !nzchar(params$source))) {
    params$source <- tools::file_path_sans_ext(basename(params$input_csv))
}

if (is.null(params$end_day)   || !nzchar(params$end_day))   params$end_day   <- format(Sys.Date())
if (is.null(params$start_day) || !nzchar(params$start_day)) params$start_day <- format(as.Date(params$end_day) - 365 * 2)

# Load libraries
if (!require("pak")) install.packages("pak")
if (!requireNamespace("epiprocess", quietly = TRUE)) pak::pkg_install("cmu-delphi/epiprocess@dev")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    epidatr, epiprocess,
    tidyverse, stringr, scales, patchwork,
    here, glue, ggtext
)
# Custom HTML-friendly label wrapper for ggtext::element_markdown
label_wrap_html <- function(width = 25) {
  fun <- label_wrap_gen(width = width)
  function(labels) {
    res <- fun(labels)
    res[] <- lapply(res, function(x) {
      if (is.character(x) || is.factor(x)) {
        gsub("\n", "<br>", as.character(x), fixed = TRUE)
      } else {
        x
      }
    })
    res
  }
}

# Invalidate knitr cache when params change
knitr::opts_chunk$set(cache.extra = list(params))

# Normalize signal/name params to unique character vectors.
signals_vec <- if (!is.null(params$signal) && length(params$signal) > 0 &&
    any(nzchar(as.character(params$signal)))) {
    unique(as.character(params$signal)[nzchar(as.character(params$signal))])
} else if (!is.null(params$input_csv) && file.exists(params$input_csv)) {
    # Signal omitted: peek at CSV to resolve signal names early (needed for config table)
    csv_cols <- readr::read_csv(params$input_csv, n_max = 0, show_col_types = FALSE)
    if ("signal" %in% colnames(csv_cols)) {
        sort(unique(readr::read_csv(params$input_csv, col_select = "signal", show_col_types = FALSE)$signal))
    } else {
        "value"
    }
} else {
    character(0)
}
multi_signal <- length(signals_vec) > 1

# `signal` is just another key column. Where the epi_df/epi_archive are built
# below, signal_format = "add_key" folds it into other_keys when the column is
# present, and "as_is" leaves the frame alone when it isn't.

# Used inline in the markdown so the prose is plural only when more than one
# indicator was loaded.
ind <- if (multi_signal) "indicators" else "indicator"
ind_poss <- if (multi_signal) "indicators'" else "indicator's"
sg_pl <- function(singular, plural) if (multi_signal) plural else singular

# Data with no signal column carries one indicator, so its label is constant
# and subsetting by indicator is a no-op. `has_signal` is set at load time; both
# helpers resolve it when called.
add_indicator <- function(df) {
    mutate(df, indicator = if (has_signal) unname(signal_labels[signal]) else name_vec[1])
}
filter_signal <- function(df, sig) if (has_signal) filter(df, signal == sig) else df

# Build signal labels (name falls back to signal if not provided)
name_vec <- {
    n <- params$name
    if (!is.null(n) && length(n) == length(signals_vec)) as.character(n)
    else signals_vec
}
name_vec <- str_wrap(name_vec, width = 15)
signal_labels <- setNames(name_vec, signals_vec)

# Resolve geo_type parameter (infer if not provided)
if (is.null(params$geo_type) || !nzchar(params$geo_type)) {
    if (!is.null(params$input_csv) && file.exists(params$input_csv)) {
        sample_df <- readr::read_csv(params$input_csv, n_max = 5, show_col_types = FALSE) %>%
                epiprocess::as_epi_df()
        params$geo_type <- attributes(sample_df)$metadata$geo_type
    } else {
        stop("Parameter 'geo_type' is required when querying datasets from the API.")
    }
}

# Set theme and colors
colours <- c(
    "#172431",
    "#ffffff",
    "#C41230",
    "#941120",
    "#007BC0",
    "#002676",
    "#FDB515",
    "#002145",
    "#00A7E1",
    "#8C338F",
    "#368f33"
)

# Palette selection
dis_pal_list <- colours[c(3, 5, 7, 10, 11, 9)]

theme_set(
    theme_minimal(base_size = 13) +
        theme(
            plot.title = ggtext::element_textbox_simple(hjust = 0.5, halign = 0.5, size = 15, face = "bold", margin = margin(b = 6)),
            plot.subtitle = element_text(hjust = 0.5, color = "gray50", margin = margin(b = 10)),
            plot.caption = ggtext::element_textbox_simple(hjust = 0, halign = 0, color = "gray50", margin = margin(t = 10)),
            panel.grid = element_line(linetype = 2, linewidth = 0.3, color = "gray90"),
            panel.grid.minor = element_blank(),
            legend.position = "top",
            legend.text = ggtext::element_markdown(size = 10),
            legend.title = ggtext::element_markdown(size = 10, face = "bold"),
            strip.text = ggtext::element_markdown(hjust = 0.5, size = 11, face = "bold", margin = margin(t = 4, b = 4)),
            strip.background = element_rect(fill = "gray95", linetype = "blank"),
            panel.border = element_rect(color = "gray95", fill = NA)
        )
)

# Palette as global defaults for discrete and continuous scales
options(
    ggplot2.discrete.colour = dis_pal_list,
    ggplot2.discrete.fill = dis_pal_list,
    ggplot2.continuous.colour = function() scale_colour_gradientn(colours = dis_pal_list),
    ggplot2.continuous.fill = function() scale_fill_gradientn(colours = dis_pal_list)
)

# Update defaults for geoms/stats
tibble::tribble(
    ~name,      ~aes,     ~alpha, ~is_stat,
    "line",     "colour", 1.0,    FALSE,
    "point",    "colour", 0.7,    FALSE,
    "function", "colour", 1.0,    FALSE,
    "bar",      "fill",   0.9,    FALSE,
    "col",      "fill",   0.9,    FALSE,
    "boxplot",  "fill",   0.8,    FALSE,
    "bin",      "fill",   0.8,    TRUE,
    "identity", "fill",   0.7,    TRUE
) %>%
    purrr::pwalk(function(name, aes, alpha, is_stat) {
        updater <- if (is_stat) update_stat_defaults else update_geom_defaults
        updater(name, setNames(list(dis_pal_list[1], alpha), c(aes, "alpha")))
    })

# An epirange in whatever format the indicator's time_type needs
time_range <- function(start, end, time_type = params$time_type) {
    if (time_type == "week") {
        to_yw <- function(d) {
            d <- as.Date(d)
            m <- MMWRweek::MMWRweek(d)
            m$MMWRyear * 100L + m$MMWRweek
        }
        epirange(to_yw(start), to_yw(end))
    } else {
        epirange(start, end)
    }
}

# Map geo codes to readable names, for display only: the codes stay in the data
# because the API queries and the HHS crosswalk join on them. Unmapped codes and
# unrecognized geo types are returned unchanged.
geo_to_name <- function(geos, geo_type = params$geo_type) {
    names <- switch(geo_type,
        state  = covidcast::abbr_to_name(toupper(geos)),
        # County names alone repeat across states (e.g. many "Washington
        # County"s), so tag each with its state abbreviation. ifelse (not
        # paste0 directly) keeps unmapped codes NA, so the fallback below
        # still catches them instead of yielding a literal "NA, NA".
        county = {
            cty_name <- covidcast::county_fips_to_name(as.character(geos))
            st_abbr <- covidcast::fips_to_abbr(as.character(geos))
            ifelse(is.na(cty_name) | is.na(st_abbr), NA_character_,
                   paste0(cty_name, ", ", st_abbr))
        },
        msa    = covidcast::cbsa_to_name(as.character(geos)),
        return(geos)
    )
    names[is.na(names)] <- geos[is.na(names)]
    names
}

# Reproducible sample of at most n elements
random_sample <- function(vec, n = 15) {
    if (length(vec) <= n) {
        return(vec)
    }
    withr::with_seed(42, sort(sample(vec, n)))
}

# Comma-separated list of locations, capped at 15
format_excluded <- function(locs) {
    if (length(locs) == 0) {
        return(NULL)
    }
    entries <- geo_to_name(random_sample(locs))
    paste0(paste(entries, collapse = ", "), if (length(locs) > 15) ", etc." else "")
}

# Order stratum levels as a reader expects rather than alphabetically: aggregate
# levels first, then by leading number so "5-17" precedes "18-49", with
# open-ended low bins like "<1" below their own number.
agg_level_names <- c("all", "total", "overall", "all ages", "all_ages", "any")
order_strata <- function(x) {
    lv <- sort(unique(as.character(x[!is.na(x)])))
    is_agg <- tolower(lv) %in% agg_level_names
    rest <- lv[!is_agg]
    num <- suppressWarnings(as.numeric(str_extract(rest, "[0-9]+\\.?[0-9]*")))
    num <- num - 0.5 * str_detect(tolower(rest), "^\\s*(<|under|less|up to)")
    factor(x, levels = c(lv[is_agg], rest[order(num, rest, na.last = TRUE)]))
}

# Order stratum columns and append each level's location count to its label,
# since panels drawn from different location sets are not comparable. The count
# is per indicator too: pooling would report the union on every panel.
label_strata_coverage <- function(df, keys, src = df_candidate) {
    src <- as_tibble(src) %>% filter(!is.na(value))
    sig <- intersect(signal_key, names(df))
    for (k in keys) {
        f <- order_strata(df[[k]])
        cnt <- src %>%
            group_by(across(all_of(c(sig, k)))) %>%
            summarize(n = n_distinct(geo_value), .groups = "drop")
        # Match on the stratum level, and on the indicator too when there is one.
        idx <- if (length(sig) == 0) {
            match(as.character(f), as.character(cnt[[k]]))
        } else {
            match(
                paste(df[[sig]], as.character(f)),
                paste(cnt[[sig]], as.character(cnt[[k]]))
            )
        }
        n <- cnt$n[idx]
        lab <- if_else(is.na(n), as.character(f), sprintf("%s (%s locs)", as.character(f), n))
        # Order by the underlying level, so appending the count does not resort
        # the panels back into alphabetical order.
        ord <- unique(lab[order(f, lab)])
        df[[k]] <- factor(lab, levels = ord[!is.na(ord)])
    }
    df
}

# Interactive time series: one dropdown entry per location, one trace per
# stratum, gaps left unbridged, markers on reported periods. A confidence band
# shares its stratum's legend entry, so hiding a stratum hides its band too.
# Called once, but kept a function on purpose: it is meant to move to
# epiprocess, which autoplot() cannot do because it draws no intervals.
plot_series_dropdown <- function(df, trace_col = NULL, ci_cols = NULL) {
    groups <- as.character(sort(unique(df$geo_value)))
    # An estimate on a near-empty denominator has an interval about as wide as
    # the whole indicator; drawn, it is a full-height column that buries every
    # other series. Skip those and report how many. The cut-off is the span of
    # the values, so it scales with the indicator rather than being a constant.
    band_limit <- if (!is.null(ci_cols)) {
        diff(range(df$value, na.rm = TRUE)) * params$ci_band_limit_mult
    } else NULL
    n_band_wide <- 0L
    n_band_obs <- 0L
    strata <- if (is.null(trace_col)) NA_character_ else levels(order_strata(df[[trace_col]]))
    # Interpolate rather than recycle: two strata sharing a colour is unreadable
    # once bands are drawn as well.
    pal <- setNames(
        if (length(strata) <= length(dis_pal_list)) {
            dis_pal_list[seq_along(strata)]
        } else {
            grDevices::colorRampPalette(dis_pal_list)(length(strata))
        },
        strata
    )

    specs <- list()
    for (g in groups) {
        for (s in strata) {
            d <- df %>% filter(geo_value == g)
            if (!is.na(s)) d <- d %>% filter(.data[[trace_col]] == s)
            d <- d %>% arrange(time_value)
            if (nrow(d) == 0 || all(is.na(d$value))) next
            lab <- if (is.na(s)) "Value" else s
            col <- unname(if (is.na(s)) dis_pal_list[1] else pal[[s]])
            if (!is.null(ci_cols)) {
                # One band per contiguous run of reported bounds. A band is a
                # c(x, rev(x)) polygon, so an NA inside it shifts the return leg
                # and detaches the band from its line. Splitting on gaps also
                # keeps a band from spanning a period with no data.
                ok <- !is.na(d[[ci_cols[1]]]) & !is.na(d[[ci_cols[2]]])
                too_wide <- ok & (d[[ci_cols[2]]] - d[[ci_cols[1]]]) > band_limit
                n_band_obs <- n_band_obs + sum(ok)
                n_band_wide <- n_band_wide + sum(too_wide)
                ok <- ok & !too_wide
                for (seg in split(d[ok, , drop = FALSE], cumsum(!ok)[ok])) {
                    specs <- append(specs, list(list(
                        d = seg, col = col, lab = lab, grp = g,
                        kind = "ci", legend_ok = FALSE
                    )))
                }
            }
            specs <- append(specs, list(list(
                d = d, col = col, lab = lab, grp = g,
                kind = "line", legend_ok = !is.na(s)
            )))
        }
    }
    if (length(specs) == 0) return(NULL)

    trace_grp <- map_chr(specs, "grp")
    legend_ok <- map_lgl(specs, "legend_ok")
    active <- trace_grp == groups[1]

    p <- reduce(seq_along(specs), function(p, i) {
        sp <- specs[[i]]
        if (sp$kind == "ci") {
            plotly::add_ribbons(p,
                x = sp$d$time_value,
                ymin = sp$d[[ci_cols[1]]], ymax = sp$d[[ci_cols[2]]],
                line = list(width = 0), fillcolor = plotly::toRGB(sp$col, 0.2),
                name = sp$lab, legendgroup = sp$lab,
                showlegend = FALSE, visible = active[i],
                hoverinfo = "skip"
            )
        } else {
            plotly::add_trace(p,
                x = sp$d$time_value, y = sp$d$value,
                type = "scatter", mode = "lines+markers", connectgaps = FALSE,
                line = list(color = sp$col, width = 1.5),
                marker = list(color = sp$col, size = 3),
                name = sp$lab, legendgroup = sp$lab,
                showlegend = active[i] && legend_ok[i], visible = active[i],
                hovertemplate = paste0(
                    "%{x|%Y-%m-%d}<br>Value: %{y:.3f}",
                    "<extra>", sp$lab, "</extra>"
                )
            )
        }
    }, .init = plotly::plot_ly())

    # Scale the y axis to the values, not the bands: letting a near-empty
    # denominator's interval set the axis compresses every series into a sliver
    # along the bottom. Overshooting bands are clipped but stay zoomable.
    yrange <- NULL
    if (!is.null(ci_cols)) {
        vals <- df$value[is.finite(df$value)]
        if (length(vals) > 0 && diff(range(vals)) > 0) {
            yrange <- c(min(0, min(vals)), max(vals) + 0.15 * diff(range(vals)))
        }
    }

    buttons <- map(groups, function(g) {
        on <- trace_grp == g
        list(
            method = "update",
            args = list(
                list(visible = as.list(on), showlegend = as.list(on & legend_ok)),
                list(title = list(text = paste0("Location: ", g)))
            ),
            label = g
        )
    })

    p <- p %>%
        plotly::layout(
            title = list(text = paste0("Location: ", groups[1])),
            xaxis = list(title = "Date"),
            yaxis = list(title = "Value", range = yrange),
            legend = list(title = list(text = trace_col %||% "")),
            updatemenus = list(list(
                type = "dropdown", active = 0, buttons = buttons, x = 0.05, y = 1.15
            ))
        ) %>%
        plotly::config(modeBarButtonsToRemove = c("zoomIn2d", "zoomOut2d"))

    # Counts ride along on the plot so the chunk can report them.
    attr(p, "bands_wide") <- n_band_wide
    attr(p, "bands_total") <- n_band_obs
    p
}

# Optional sections, all on when the param is absent.
smooth_quantiles <- !isFALSE(params$smooth_quantiles)
show_geo_outliers <- !isFALSE(params$show_geo_outliers)
show_versioning <- !isFALSE(params$show_versioning)

# A raw count scales with a
# location's size, so its cross-location spread is not what section 2.3 is meant
# to show. Infer from the signal/name unless the caller sets the flag explicitly.
count_like_name <- function(x) {
    grepl("cumulative|(^|[ _])(num|count|counts|n)([ _]|$)|admission|encounter|(^|[ _])(ed|ip|op)$",
          x, ignore.case = TRUE) &&
        !grepl("100k|100000|per.?capita|pct|percent|rate|ratio|prop|/100|%",
               x, ignore.case = TRUE)
}
comparable_across_locations <- if (is.null(params$comparable_across_locations)) {
    !any(vapply(c(signals_vec, name_vec), count_like_name, logical(1)))
} else isTRUE(params$comparable_across_locations)

# The kind of value the indicator carries. A categorical `value_type` that the
# report can branch on; for now only "cumulative" does anything (section 2.4).
# Other categories are inferred and carried through for later use.
infer_value_type <- function(x) {
    x <- paste(x, collapse = " ")
    if (grepl("cumulative", x, ignore.case = TRUE)) {
        "cumulative"
    } else if (grepl("100k|100000|per.?capita|(^|[ _])rate([ _]|$)|incidence", x, ignore.case = TRUE)) {
        "rate"
    } else if (grepl("pct|percent|%", x, ignore.case = TRUE)) {
        "percent"
    } else if (grepl("prop|ratio|frac", x, ignore.case = TRUE)) {
        "proportion"
    } else if (any(vapply(c(signals_vec, name_vec), count_like_name, logical(1)))) {
        "count"
    } else {
        "other"
    }
}
value_type_param <- flatten_param(params$value_type)
value_type <- if (length(value_type_param) == 0 || !nzchar(trimws(value_type_param[1]))) {
    infer_value_type(c(signals_vec, name_vec))
} else {
    tolower(trimws(value_type_param[1]))
}
cumulative_indicator <- identical(value_type, "cumulative")

# Assign each observation to an epidemiological year starting in
# `season_start_month` (7 = July-June, the usual influenza convention).
season_start_month <- as.integer(params$season_start_month %||% 7L)
season_label <- sprintf(
    "%s-%s", month.name[season_start_month],
    month.name[((season_start_month + 10L) %% 12L) + 1L]
)
add_season <- function(df) {
    df %>%
        mutate(
            yr = as.integer(format(time_value, "%Y")),
            month = as.integer(format(time_value, "%m")),
            season_start = if_else(month >= season_start_month, yr, yr - 1L),
            season = paste0(season_start, "-", substr(as.character(season_start + 1L), 3L, 4L)),
            season_month = ((month - season_start_month) %% 12L) + 1L
        )
}

# Set cache
cache_path <- file.path(getwd(), "epidatr_cache")
if (!dir.exists(cache_path)) dir.create(cache_path, recursive = TRUE)

set_cache(cache_dir = cache_path)
```

This notebook evaluates `r sg_pl("a **candidate indicator**", "the **candidate indicators**")` to understand `r sg_pl("its", "their")` characteristics, coverage, missingness, and versioning/revision behavior.

---

## Data Loading {#data}

```{r}
#| label: load-candidate
#| results: asis
# Load candidate indicator. `load_note` records where the data came from and
# `n_dropped` how many rows were discarded, both reported in the prose below.
candidate_archive <- NULL
load_note <- "the source"
n_dropped <- 0L
if (!is.null(params$input_csv)) {
    # geo_value stays character: a numeric guess would strip the leading zero
    # from county FIPS codes in states 01-09 (AL, AK, AZ, AR, CA, CO, CT).
    df_candidate_raw <- read.csv(params$input_csv, colClasses = c(geo_value = "character")) %>%
        mutate(time_value = as.Date(time_value)) %>%
        filter(time_value >= as.Date(params$start_day) & time_value <= as.Date(params$end_day))

    # A file may hold more indicators than were asked for. Keep only the
    # requested ones, or every one of them would be analysed and labelled NA.
    if ("signal" %in% colnames(df_candidate_raw) && length(signals_vec) > 0) {
        df_candidate_raw <- df_candidate_raw %>% filter(signal %in% signals_vec)
    }

    # CSV may include version/issue columns for revision analysis
    has_versions <- any(c("version", "issue") %in% colnames(df_candidate_raw))
    if (has_versions) {
        df_candidate_raw <- df_candidate_raw %>%
            {
                if ("issue" %in% names(.)) rename(., version = issue) else .
            } %>%
            mutate(version = as.Date(version))

        # A row with no version has no place on a version timeline. A file
        # holding both versioned and unversioned copies of the same observations
        # is usually a build accident, and keeping both doubles every series.
        n_dropped <- sum(is.na(df_candidate_raw$version))
        if (n_dropped > 0) df_candidate_raw <- df_candidate_raw %>% filter(!is.na(version))

        # Confidence bounds ride along as extra value columns, so the overview
        # plot can draw bands for versioned CSVs too.
        csv_archive_df  <- df_candidate_raw %>%
            select(geo_value, time_value, version, any_of("signal"), any_of(params$extra_keys),
                   value, any_of(c("ci_lower", "ci_upper"))) %>%
            group_by(geo_value, time_value, version, pick(any_of(c("signal", params$extra_keys)))) %>%
            slice_max(value, n = 1, with_ties = FALSE) %>%
            ungroup()

        candidate_archive <- as_epi_archive(csv_archive_df, other_keys = params$extra_keys,
            signal_format = if ("signal" %in% names(csv_archive_df)) "add_key" else "as_is")
        df_candidate <- epix_as_of(candidate_archive, candidate_archive$versions_end)
        n_versions <- n_distinct(df_candidate_raw$version)
        load_note <- sprintf("a local CSV with %s version%s of history",
                             label_comma()(n_versions), if (n_versions == 1) "" else "s")
    } else {
        df_candidate <- as_epi_df(df_candidate_raw, other_keys = params$extra_keys,
            signal_format = if ("signal" %in% names(df_candidate_raw)) "add_key" else "as_is")
        load_note <- "a local CSV"
    }
} else {
    # Determine if new API can be used for candidate source
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE
    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                all(signals_vec %in% meta_res[[params$source]]$signals)
            },
            error = function(e) FALSE
        )
        if (meta_available) use_new_api <- TRUE
    }
    if (use_new_api) {
        # epidata_snapshot does not accept multiple signals
        df_candidate <- purrr::map(signals_vec, ~ {
            epidata_snapshot(
                source = params$source,
                signals = .x,
                geo_type = params$geo_type
            ) %>%
                rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
                {
                    if ("fill_method" %in% colnames(.)) {
                        filter(., fill_method == "source" | !any(fill_method == "source"))
                    } else .
                } %>%
                filter(
                    time_value >= as.Date(params$start_day),
                    time_value <= as.Date(params$end_day)
                ) %>%
                select(geo_value, any_of("signal"), time_value, any_of(params$extra_keys),
                       value, any_of(c("ci_lower", "ci_upper")))
        }) %>%
            bind_rows() %>%
            as_epi_df(other_keys = params$extra_keys,
                      signal_format = if ("signal" %in% names(.)) "add_key" else "as_is")
    } else {
        # pub_covidcast accepts a signal vector and always returns a signal column
        df_candidate <- pub_covidcast(
            params$source, signals_vec,
            params$geo_type, params$time_type,
            time_values = time_range(params$start_day, params$end_day),
            as_of = Sys.Date()
        ) %>% as_epi_df(other_keys = params$extra_keys, signal_format = "add_key")
    }
    load_note <- sprintf("the %s API", params$source)
}

# Resolved before the archive fetch so the table below can report periods
# and strata.
env_time_type <- attributes(df_candidate)$metadata$time_type
if (is.null(env_time_type) || length(env_time_type) != 1 || is.na(env_time_type) || !is.character(env_time_type)) {
    env_time_type <- params$time_type
}
if (is.null(env_time_type) || !nzchar(env_time_type)) {
    env_time_type <- "day"
}
params$time_type <- env_time_type

env_geo_type <- attributes(df_candidate)$metadata$geo_type
if (is.null(params$geo_type) || length(params$geo_type) != 1 || is.na(params$geo_type) || !is.character(params$geo_type)) {
    params$geo_type <- env_geo_type
}

# The keys beyond geo_value/time_value, `signal` among them whenever the column
# exists. Every grouping, join and denominator below uses this vector, so both
# kinds of key are handled the same way.
has_signal <- "signal" %in% colnames(df_candidate)
signal_key <- if (has_signal) "signal" else character(0)
# Signal first, so tables that show it lead with the Indicator column.
strat_keys <- c(signal_key, params$extra_keys)

n_locations <- n_distinct(df_candidate$geo_value)
n_periods <- n_distinct(df_candidate$time_value)

# Stratum levels per extra key, and the (period x stratum) cells a location
# could fill. Used by the table below and by the missingness section.
strata_levels <- map(set_names(params$extra_keys), ~ sort(unique(as.character(df_candidate[[.x]]))))
n_key_combos <- if (length(params$extra_keys) > 0) {
    df_candidate %>% as_tibble() %>% distinct(across(all_of(params$extra_keys))) %>% nrow()
} else {
    1L
}
n_strata_cells <- max(1L, prod(lengths(strata_levels)))

# Length of the requested window, in the indicator's own periods
req_days <- as.integer(as.Date(params$end_day) - as.Date(params$start_day)) + 1L
req_periods <- round(req_days / switch(env_time_type,
    day = 1, week = 7, month = 30.44, year = 365.25, 1
))
```

Loaded `r label_comma()(nrow(df_candidate))` rows via `r load_note`.`r if (n_dropped > 0) sprintf(" %s row(s) carried no version and were dropped.", label_comma()(n_dropped))` The table below describes the data as loaded.

```{r}
#| label: print-config
#| echo: false
config_tbl <- tibble(
    Field = c(
        "Source", "Indicator(s)", "Label(s)", "Geography", "Locations",
        "Extra key(s)", "Requested period", "Observed period"
    ),
    Value = c(
        params$source,
        if (has_signal) paste(sort(unique(df_candidate$signal)), collapse = ", ") else paste(signals_vec, collapse = ", "),
        paste(str_remove(name_vec, "\\n"), collapse = ", "),
        if (is.null(params$geo_type) || !nzchar(params$geo_type)) NA_character_ else params$geo_type,
        label_comma()(n_locations),
        if (length(params$extra_keys) == 0) {
            NA_character_
        } else {
            paste(sprintf(
                "%s (%d level(s): %s)", names(strata_levels), lengths(strata_levels),
                map_chr(strata_levels, ~ paste(levels(order_strata(.x)), collapse = ", "))
            ), collapse = "; ")
        },
        sprintf("%s to %s (about %s %s(s))", params$start_day, params$end_day,
                label_comma()(req_periods), env_time_type),
        sprintf("%s to %s (%s %s(s) observed)",
                format(min(df_candidate$time_value)), format(max(df_candidate$time_value)),
                label_comma()(n_periods), env_time_type)
    )
) %>%
    filter(!is.na(Value))

DT::datatable(
    config_tbl,
    caption = "Analysis configuration and loaded dataset",
    rownames = FALSE,
    options = list(scrollX = TRUE, dom = "t", paging = FALSE)
)
```

```{r}
#| label: load-candidate-archive
#| results: asis
# Revision history from the API, unless the CSV already supplied it. Skipped
# when the versioning section is off: issues = "*" is the slowest query here.
if (show_versioning && is.null(candidate_archive) && is.null(params$input_csv)) {
    # Only fetch version history if the source supports revisions
    new_api_sources <- c("nssp", "nhsn", "pophive", "nwss", "claims_data_outpatient")
    use_new_api <- FALSE
    if (params$source %in% new_api_sources) {
        meta_available <- tryCatch(
            {
                meta_res <- epidata_meta(params$source)
                all(signals_vec %in% meta_res[[params$source]]$signals)
            },
            error = function(e) FALSE
        )
        if (meta_available) use_new_api <- TRUE
    }

    # Fetch archive for revision behavior
    archive_geos <- unique(df_candidate$geo_value)
    # Limit number of locations for version history queries to avoid API limits
    if (length(archive_geos) > params$max_locations_plot) {
        archive_geos <- random_sample(archive_geos, params$max_locations_plot)
    }

    # Reach back before the analysis window to catch earlier issues
    archive_start <- as.Date(params$start_day) - 60

    archive_info <- list(
        n_sampled = length(archive_geos),
        n_total = length(unique(df_candidate$geo_value)),
        time_value_start = archive_start,
        time_value_end = params$end_day
    )

    if (use_new_api) {
        # epidata_archive does not accept multiple signals
        df_archive_raw <- purrr::map(signals_vec, ~ {
            epidata_archive(
                source = params$source,
                signals = .x,
                geo_type = params$geo_type
            ) %>%
                rename(any_of(c(time_value = "reference_time", version = "report_time"))) %>%
                filter(
                    geo_value %in% archive_geos,
                    time_value >= as.Date(archive_start),
                    time_value <= as.Date(params$end_day)
                )
        }) %>% bind_rows()
    } else {
        # pub_covidcast accepts a signal vector and returns a signal column
        df_archive_raw <- pub_covidcast(
            source = params$source,
            signal = signals_vec,
            geo_type = params$geo_type,
            time_type = params$time_type,
            geo_values = archive_geos,
            time_values = time_range(archive_start, params$end_day),
            issues = "*",
        )
    }

    if (nrow(df_archive_raw) > 0) {
        # ponytail: no dedup here. Sources like NWSS can return several rows per
        # (geo, period, version) from different measurement sources, which
        # as_epi_archive() rejects; add a slice_max if one turns up.
        df_archive_raw <- df_archive_raw %>%
            rename(any_of(c(time_value = "reference_time", version = "issue"))) %>%
            rename(any_of(c(version = "report_time"))) %>%
            select(geo_value, time_value, version, any_of("signal"), any_of(params$extra_keys), value)

        if ("version" %in% colnames(df_archive_raw)) {
            archive_info$version_start <- min(df_archive_raw$version)
            archive_info$version_end <- max(df_archive_raw$version)
            candidate_archive <- as_epi_archive(df_archive_raw, other_keys = params$extra_keys,
                signal_format = if ("signal" %in% names(df_archive_raw)) "add_key" else "as_is")
        }
    }
}
```

```{r}
#| label: revision-analysis-setup
#| echo: false

archive <- NULL
archive_has_history <- FALSE
if (!is.null(candidate_archive) &&
    n_distinct(candidate_archive$DT$version) > 1) {
    archive <- candidate_archive
    archive_span_days <- as.numeric(
        max(archive$DT$version) - min(archive$DT$version),
        units = "days"
    )
    archive_has_history <- archive_span_days >= params$min_archive_days
}
```

```{r}
#| label: quarto-metadata
#| results: asis
#| echo: false
# Quarto metadata for the ::: {.content-visible} blocks below. A section whose
# flag is false is dropped
cat(sprintf(
    "\n---\nmulti_location: %s\nhas_strata: %s\ngeo_outliers: %s\nversioning: %s\nlocation_comparable: %s\ncross_location_ok: %s\n---\n",
    tolower(n_locations > 1),
    tolower(length(params$extra_keys) > 0),
    tolower(show_geo_outliers && n_locations > 1),
    tolower(show_versioning && !is.null(archive)),
    tolower(comparable_across_locations),
    tolower(!cumulative_indicator)
))
```

## Exploratory Data Analysis {#eda}

The data covers `r n_locations` `r params$geo_type`s at `r env_time_type` frequency.

Two words are used throughout. An **observation** is one row: a single value for one location, one `r env_time_type``r if (length(params$extra_keys) > 0) sprintf(" and one level of %s", paste(params$extra_keys, collapse = "/")) else ""`. A **location** is one `r params$geo_type`, so location counts are never observation counts.

::: {.content-visible when-meta="multi_location"}
### Coverage and Missingness

Coverage comes first because everything after it describes only the values that exist.

A value is **missing** when no row exists for a (location, period`r if (length(params$extra_keys) > 0) ", stratum" else ""`) cell. Each location could fill `r label_comma()(n_periods)` `r env_time_type` period(s)`r if (n_strata_cells > 1) sprintf(" x %s stratum cell(s)", label_comma()(n_strata_cells)) else ""` = `r label_comma()(n_periods * n_strata_cells)` **possible periods with values**. Periods before any location's first reported period fall outside that grid, which is why the observed and requested windows differ in the table above.

The chart below tracks how many locations report in each period, against the `r n_locations` in the dataset (dotted line).

```{r}
#| label: eda-coverage
#| echo: false
#| results: asis
coverage <- df_candidate %>%
    group_by(across(all_of(c(strat_keys, "time_value")))) %>%
    summarize(n_available = sum(!is.na(value)), .groups = "drop") %>%
    mutate(geo_value = "all") %>%
    as_epi_df(signal_format = "as_is", other_keys = strat_keys) %>%
    as_tibble() %>%
    add_indicator() %>%
    label_strata_coverage(params$extra_keys)

# The full (period x stratum) x location grid, NA wherever a location has no
# row. Every missingness figure derives from it, so the per-location and
# per-stratum views share one denominator.
summary_base <- df_candidate %>%
    as_tibble() %>%
    select(all_of(c(strat_keys, "geo_value", "time_value")), value) %>%
    complete(nesting(!!!rlang::syms(c(signal_key, "time_value"))),
             !!!rlang::syms(params$extra_keys), geo_value)
```

```{r}
#| label: eda-coverage-plot
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 2), 12)
#| echo: true
ggplot(coverage, aes(x = time_value, y = n_available,
                     color = indicator, linetype = indicator)) +
    geom_hline(yintercept = n_locations, linetype = 3, linewidth = 0.4, color = "gray50") +
    geom_line(linewidth = 0.8) +
    scale_x_date(breaks = breaks_pretty(), label = label_date_short()) +
    scale_y_continuous(labels = label_comma(), limits = c(0, NA)) +
    labs(
        title = "Data coverage over time",
        subtitle = sprintf("Number of locations with non-missing values per %s (dotted line: %d locations in the dataset)",
                           env_time_type, n_locations),
        x = "Date", y = "Number of locations",
        color = "Indicator", linetype = "Indicator"
    ) +
    {if (length(signals_vec) > 6 || length(params$extra_keys) > 0) facet_wrap(reformulate(c(if (length(signals_vec) > 6) "indicator", params$extra_keys)), ncol = 2, labeller = label_wrap_html(width = 25)) else NULL}
```

::: {.content-visible when-meta="has_strata"}

#### Missingness by Stratifier

```{r}
#| label: eda-missingness-by-stratum
#| echo: false
#| eval: !expr length(params$extra_keys) > 0
show_ind_col <- has_signal && multi_signal

strat_miss <- summary_base %>%
    group_by(across(all_of(strat_keys))) %>%
    summarize(
        n_possible = n(),
        n_obs = sum(!is.na(value)),
        n_locs = n_distinct(geo_value[!is.na(value)]),
        n_periods_obs = n_distinct(time_value[!is.na(value)]),
        mean_val = mean(value, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    mutate(
        pct_missing = 100 * (1 - n_obs / n_possible),
        across(all_of(params$extra_keys), order_strata)
    ) %>%
    add_indicator() %>%
    arrange(across(all_of(params$extra_keys)), across(all_of(signal_key))) %>%
    select(
        any_of(if (show_ind_col) "indicator" else character(0)),
        all_of(params$extra_keys), n_locs, n_periods_obs, n_obs, n_possible,
        pct_missing, mean_val
    )

DT::datatable(
    strat_miss,
    colnames = c(
        if (show_ind_col) "Indicator" else NULL,
        str_to_title(str_replace_all(params$extra_keys, "_", " ")),
        "Locations Reporting",
        sprintf("%ss Reporting", str_to_title(env_time_type)),
        "Periods with Values Reported", "Possible Periods with Values", "% Missing", "Avg. Value"
    ),
    caption = sprintf(
        "Missingness by stratifier. Possible periods with values = %s %s period(s) x %s location(s) for each stratum.",
        label_comma()(n_periods), env_time_type, label_comma()(n_locations)
    ),
    rownames = FALSE, options = list(scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
) %>% DT::formatRound(columns = which(names(strat_miss) %in% c("pct_missing", "mean_val")), digits = 2)
```

:::

#### Summary Statistics by Location

This table summarizes the available data across all included locations. Locations with no candidate information are not considered.

```{r}
#| label: eda-summary
#| results: asis
#| echo: false
# Per (indicator, location) against the full grid, so a location missing an
# entire stratum counts as missing rather than being ignored.
summary_tbl_all <- summary_base %>%
    group_by(across(all_of(c(signal_key, "geo_value")))) %>%
    summarize(
        n_possible = n(),
        n_obs = sum(!is.na(value)),
        cand_mean = mean(value, na.rm = TRUE),
        cand_var = var(value, na.rm = TRUE),
        cand_miss_frac = 1 - n_obs / n_possible,
        cand_miss = sprintf("%.1f%%", 100 * (1 - n_obs / n_possible)),
        .groups = "drop"
    ) %>%
    add_indicator()

excluded_locs_df <- summary_tbl_all %>%
    filter(is.na(cand_mean) | cand_miss_frac == 1)
if (nrow(excluded_locs_df) > 0) {
    for (sig in signals_vec) {
        ex_sig <- excluded_locs_df %>% filter_signal(sig)
        if (nrow(ex_sig) > 0) {
            cat(sprintf(
                "**Excluded locations (no data) — %s**: %s\n\n",
                signal_labels[sig],
                format_excluded(ex_sig$geo_value)
            ))
        }
    }
}

zero_var_cand_locs_df <- summary_tbl_all %>%
    filter(!is.na(cand_var) & cand_var == 0 & cand_miss_frac < 1)
if (nrow(zero_var_cand_locs_df) > 0) {
    for (sig in signals_vec) {
        zv_sig <- zero_var_cand_locs_df %>% filter_signal(sig)
        if (nrow(zv_sig) > 0) {
            cat(sprintf(
                "**Zero variance locations (flatlined) — %s**: %s\n\n",
                signal_labels[sig],
                format_excluded(zv_sig$geo_value)
            ))
        }
    }
}

# Summary of missingness levels per indicator
miss_levels <- summary_tbl_all %>%
    mutate(
        cand_miss_group = case_when(
            cand_miss_frac == 1 ~ "100% missing",
            cand_miss_frac < 0.05 ~ "<5% missing",
            cand_miss_frac < 0.25 ~ "5-25% missing",
            cand_miss_frac < 0.50 ~ "25-50% missing",
            TRUE ~ "50-99.9% missing"
        ),
        cand_miss_group = factor(cand_miss_group, levels = c("<5% missing", "5-25% missing", "25-50% missing", "50-99.9% missing", "100% missing"))
    )

for (sig in signals_vec) {
    cat(sprintf("**Missingness Overview%s:**\n\n",
                if (multi_signal) sprintf(" — %s", signal_labels[sig]) else " (All Locations)"))
    cand_counts <- miss_levels %>%
        filter_signal(sig) %>%
        count(cand_miss_group) %>%
        complete(cand_miss_group, fill = list(n = 0))
    for (i in seq_len(nrow(cand_counts))) {
        cat(sprintf("- **%s**: %d locations\n",
                    cand_counts$cand_miss_group[i], cand_counts$n[i]))
    }
    cat("\n")
}

summary_tbl <- summary_tbl_all %>%
    filter(!is.na(cand_mean) & cand_miss_frac < 1) %>%
    arrange(desc(cand_miss_frac)) %>%
    mutate(geo_value = geo_to_name(geo_value)) %>%
    select(
        any_of(if (has_signal && multi_signal) "indicator" else character(0)),
        geo_value, n_obs, n_possible, cand_mean, cand_miss
    )

DT::datatable(
    summary_tbl,
    colnames = c(
        if (has_signal && multi_signal) "Indicator" else NULL,
        sprintf("Location (%s)", params$geo_type),
        "Periods with Values Reported",
        "Possible Periods with Values",
        "Avg. Value",
        "% Missing"
    ),
    caption = sprintf(
        "Summary statistics by location. Possible periods with values per location = %s %s period(s)%s = %s; %% missing is the share of those cells with no reported value.",
        label_comma()(n_periods), env_time_type,
        if (n_strata_cells > 1) sprintf(" x %s stratum cell(s)", label_comma()(n_strata_cells)) else "",
        label_comma()(n_periods * n_strata_cells)
    ),
    rownames = FALSE, options = list(scrollX = TRUE, scrollY = "400px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
) %>% DT::formatRound(columns = which(names(summary_tbl) == "cand_mean"), digits = 2)
```

:::

### Time Series Overview

The candidate `r ind` over time, one location at a time (use the dropdown to switch). Values are shown as published: the line breaks where a period has no value and each marker is one observation, so gaps read as gaps rather than as straight interpolated lines.

```{r}
#| label: eda-overview
#| fig-height: 4
#| results: asis
all_geos <- sort(unique(df_candidate$geo_value))
sample_geos <- random_sample(all_geos, min(length(all_geos), params$max_locations_plot))

# Bands are drawn when the source reports bounds.
ci_cols <- intersect(c("ci_lower", "ci_upper"), colnames(df_candidate))
if (length(ci_cols) < 2) ci_cols <- NULL

df_sample <- df_candidate %>%
    as_tibble() %>%
    filter(geo_value %in% sample_geos) %>%
    # Fill the grid so missing periods become explicit NAs; plotly then leaves a
    # gap instead of joining across them.
    complete(nesting(!!!rlang::syms(c(signal_key, "time_value"))),
             !!!rlang::syms(params$extra_keys), geo_value) %>%
    add_indicator() %>%
    # Name the dropdown entries; nothing downstream joins on the code.
    mutate(geo_value = geo_to_name(geo_value))

# One trace per indicator and, if present, per level of the first stratifier
# (combined into a single key so both distinctions get their own legend entry).
trace_parts <- c(if (multi_signal) "indicator" else NULL,
                  if (length(params$extra_keys) > 0) params$extra_keys[1] else NULL)
trace_col <- if (length(trace_parts) > 1) {
    df_sample[[".trace"]] <- do.call(paste, c(unname(df_sample[trace_parts]), sep = " / "))
    ".trace"
} else if (length(trace_parts) == 1) trace_parts else NULL

if (!is.null(ci_cols)) {
    cat(sprintf(
        "Shaded bands are the confidence interval reported with each value%s.\n\n",
        if (!is.null(trace_col)) {
            sprintf(", drawn in the colour of their %s and sharing its legend entry, so clicking one in the legend hides its line and its band together",
                    if (length(trace_parts) > 1) paste(trace_parts, collapse = "/") else trace_col)
        } else ""
    ))
}

p <- plot_series_dropdown(df_sample, trace_col = trace_col, ci_cols = ci_cols)
if (!is.null(p) && !is.null(attr(p, "bands_wide")) && attr(p, "bands_wide") > 0) {
    cat(sprintf(
        "An estimate on almost no denominator has an interval about as wide as the indicator itself, which drawn is a full-height column hiding every other series. %s of %s reported values (%s) have an interval wider than the full range of the values, so no band is drawn for them. Those observations still show as a line and a marker, and the y axis follows the values rather than the bands, so a band running off the top is clipped.\n\n",
        label_comma()(attr(p, "bands_wide")), label_comma()(attr(p, "bands_total")),
        label_percent(accuracy = 0.1)(attr(p, "bands_wide") / attr(p, "bands_total"))
    ))
}
# SVG, not toWebGL: WebGL cannot fill a ribbon polygon and would break the
# bands. Only one location shows at a time, so SVG is fast enough.
if (!is.null(p)) p %>% plotly::partial_bundle()
```

::: {.content-visible when-meta="multi_location"}
::: {.content-visible when-meta="location_comparable"}

### Location Variability

The distribution of the `r ind` across locations over time. The line is the median, the inner ribbon the 25th-75th percentile range (the middle 50% of locations), the outer ribbon the 5th-95th. Each panel has its own scale.

Ribbon width is geographic variability: narrow means locations agree, wide means they spread.

`r if (smooth_quantiles) "Quantiles are smoothed with a moving average, which is meant for raw data whose day-of-week effect would otherwise dominate the ribbons. An indicator already smoothed at the source does not need it, and smoothing twice only widens the effective window; set <code>smooth_quantiles: false</code> to plot the quantiles as published." else "Quantiles are plotted as published (<code>smooth_quantiles: false</code>). Set the parameter to <code>true</code> for raw data whose day-of-week effect would otherwise dominate the ribbons."`

Each quantile covers only the locations that reported in that window, so panels with different location counts (shown in the panel labels) are not comparable. A spike in one panel alone is often composition: a location reporting only the aggregate stratum, on a small denominator, lifts that panel's upper quantile and no other.

```{r}
#| label: eda-quantiles
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 1.5), 15)
#| eval: !expr comparable_across_locations && n_locations > 1
cat("\n")

# One row per (key, period); geo_value is constant so epi_slide partitions by
# the keys alone.
df_quantiles_wide <- df_candidate %>%
    group_by(across(all_of(c(strat_keys, "time_value")))) %>%
    summarize(
        vals = list(value),
        geo_value = "all",
        .groups = "drop"
    ) %>%
    as_epi_df(signal_format = "as_is", other_keys = strat_keys)

# Smoothing window, matched to the time type.
window_size <- switch(env_time_type %||% "unknown",
    day   = 7L, # 7-day smoothing
    week  = as.difftime(3, units = "weeks"), # 3-week smoothing
    month = as.difftime(3, units = "months"), # 3-month smoothing
    year  = as.difftime(3, units = "years"), # 3-year smoothing
    Inf # Default to cumulative for unknown types
    # (too sparse for sliding averages)
)

q_tbl <- function(v) {
    tibble(
        `0.05` = quantile(v, 0.05, na.rm = TRUE),
        `0.25` = quantile(v, 0.25, na.rm = TRUE),
        `0.5`  = quantile(v, 0.5, na.rm = TRUE),
        `0.75` = quantile(v, 0.75, na.rm = TRUE),
        `0.95` = quantile(v, 0.95, na.rm = TRUE)
    )
}

df_quantiles_wide <- {
    if (smooth_quantiles) {
        df_quantiles_wide %>%
            group_by(across(all_of(c("geo_value", strat_keys)))) %>%
            epi_slide(~ q_tbl(unlist(.x$vals)), .window_size = window_size) %>%
            ungroup()
    } else {
        # Quantiles of the values published in that period alone.
        bind_cols(df_quantiles_wide, map_dfr(df_quantiles_wide$vals, q_tbl))
    }
} %>%
    as_tibble() %>%
    add_indicator() %>%
    label_strata_coverage(params$extra_keys)

ggplot(df_quantiles_wide, aes(x = time_value)) +
    geom_ribbon(aes(ymin = `0.05`, ymax = `0.95`, fill = "5%-95%"), alpha = 0.2) +
    geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`, fill = "25%-75%"), alpha = 0.4) +
    geom_line(aes(y = `0.5`, color = "Median"), linewidth = 1) +
    facet_wrap(reformulate(c("indicator", params$extra_keys)), ncol = 1, scales = "free_y") +
    scale_x_date(breaks = breaks_pretty(), label = label_date_short()) +
    scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_si(""))) +
    scale_fill_manual(values = c("5%-95%" = dis_pal_list[2], "25%-75%" = dis_pal_list[2])) +
    scale_color_manual(values = c("Median" = dis_pal_list[2])) +
    labs(
        title = if (smooth_quantiles) "Smoothed Quantile Trends" else "Quantile Trends",
        subtitle = sprintf("Across the locations reporting each panel (of %d in total)%s",
                           n_locations,
                           if (smooth_quantiles) sprintf(", smoothed window: %s", format(window_size)) else ", as published"),
        x = "Date", y = if (smooth_quantiles) "Value (smoothed)" else "Value",
        fill = "Range", color = "Trend"
    ) +
    theme(legend.position = "bottom")
```

::: {.content-visible when-meta="geo_outliers"}

#### Cross-Sectional Geographic Outliers

The quantile plots summarize the body of the distribution and can hide erratic individual locations. This table places each one within it, counting how often the location sits above the 95th percentile or below the 5th for its own period and stratum. Both shares are out of the periods the location reported, not the analysis window, and the count of reported values is shown so a location with three observations is not read as a stable 100%.

How to read it: 5% is where chance alone puts a location, so only distances from 5% carry information. A location above the 95th percentile most of the time is not having a long outbreak, it is sitting at a different level from everyone else. The same holds at the bottom, where a high zero share is the more common symptom. A location in *both* tails is unstable rather than high or low.

::: {.callout-note}
This compares locations against one another on each period, so it reports relative position only, never whether a value is correct. If the `r ind` `r sg_pl("relies", "rely")` on raw counts rather than rates or percentages, the table will mostly find the largest populations, reflecting size rather than outliers. Set `show_geo_outliers: false` to drop the section.
:::

```{r}
#| label: eda-aberrant-locations
#| echo: false
#| results: asis
#| eval: !expr (show_geo_outliers && n_locations > 1)
# Only reported values take part: a location cannot be in a tail on a period it
# never reported, and it should not be charged for one either.
df_reported <- df_candidate %>%
    as_tibble() %>%
    filter(!is.na(value))

# For a non-negative indicator the zero share is worth reporting alongside the
# lower tail, since the 5th percentile is often zero itself.
show_zero_share <- all(df_reported$value >= 0)

# Cross-sectional tails per (indicator, stratum, period)
daily_tails <- df_reported %>%
    group_by(across(all_of(c(strat_keys, "time_value")))) %>%
    summarize(
        cand_p05 = quantile(value, 0.05),
        cand_p95 = quantile(value, 0.95),
        .groups = "drop"
    )

aberrant_locs <- df_reported %>%
    left_join(daily_tails, by = c(strat_keys, "time_value")) %>%
    group_by(across(all_of(c(strat_keys, "geo_value")))) %>%
    summarize(
        n_obs = n(),
        upper_pct = 100 * mean(value > cand_p95),
        lower_pct = 100 * mean(value < cand_p05),
        zero_pct  = 100 * mean(value == 0),
        .groups = "drop"
    ) %>%
    mutate(geo_value = geo_to_name(geo_value)) %>%
    add_indicator() %>%
    arrange(desc(upper_pct))

if (nrow(aberrant_locs) > 0) {
    aberrant_display <- aberrant_locs %>%
        select(
            any_of(if (has_signal && multi_signal) "indicator" else character(0)),
            all_of(params$extra_keys), geo_value, n_obs,
            upper_pct, lower_pct, any_of(if (show_zero_share) "zero_pct" else character(0))
        )
    DT::datatable(
        aberrant_display,
        colnames = c(
            if (has_signal && multi_signal) "Indicator" else NULL,
            str_to_title(str_replace_all(params$extra_keys, "_", " ")),
            "Location",
            "Periods with Values Reported",
            "% Above 95th",
            "% Below 5th",
            if (show_zero_share) "% Exactly Zero" else NULL
        ),
        caption = "Summary of aberrant locations (time spent in each tail, as a share of the periods the location reported).",
        rownames = FALSE, options = list(scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
    ) %>% DT::formatRound(
        columns = which(names(aberrant_display) %in% c("upper_pct", "lower_pct", "zero_pct")),
        digits = 1
    )
} else {
    cat("Insufficient data to calculate aberrant time for these locations.\n\n")
}
```

:::

:::

:::


::: {.content-visible when-meta="multi_location"}
::: {.content-visible when-meta="cross_location_ok"}

### Temporal Variability

The coefficient of variation (CV) measures how much the `r ind` `r sg_pl("moves", "move")` over time in each location. A high CV is not evidence of noise on its own: epidemic dynamics, seasonality and genuine level shifts all raise it. Read it as the size of the variation, use the seasonality and revision sections to judge how much of it is real, and treat it comparatively.

The CV uses only reported values, so missing periods neither contribute nor count against it, and it is computed per (location, stratum). Pooling strata would measure the spread between them, since a location's strata usually sit at different levels.

The CV is undefined for a location with a mean of zero or fewer than two values. Those series cannot appear in the histogram, so the caption counts them and the bars reconcile against the location count.

$$\text{CV}_i = \frac{\text{SD}(X_i)}{|\bar{X}_i|}$$

```{r}
#| label: eda-noise
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 0.8), 12)
#| eval: !expr n_locations > 1 && isFALSE(cumulative_indicator)
# One CV per (indicator, stratum, location), over reported values only
noise_stats <- df_candidate %>%
    as_tibble() %>%
    filter(!is.na(value)) %>%
    group_by(across(all_of(c(strat_keys, "geo_value")))) %>%
    summarize(
        n_obs = n(),
        cand_cv = sd(value) / abs(mean(value)),
        .groups = "drop"
    )

n_series <- nrow(noise_stats)
noise_undefined <- noise_stats %>% filter(!is.finite(cand_cv))
noise_stats <- noise_stats %>% filter(is.finite(cand_cv))

# Clip extreme outliers so the histogram stays readable
cv_upper <- quantile(noise_stats$cand_cv, 0.99, na.rm = TRUE)
noise_stats_plot <- noise_stats %>%
    filter(cand_cv <= cv_upper) %>%
    add_indicator() %>%
    label_strata_coverage(params$extra_keys)

n_excluded <- nrow(noise_stats) - nrow(noise_stats_plot)

# Summary Table — one row per (indicator, stratum)
noise_summary <- noise_stats %>%
    group_by(across(all_of(strat_keys))) %>%
    summarize(median_cv = median(cand_cv), n_locs = n_distinct(geo_value), .groups = "drop") %>%
    mutate(across(all_of(params$extra_keys), order_strata)) %>%
    add_indicator() %>%
    arrange(across(all_of(params$extra_keys))) %>%
    select(
        any_of(if (has_signal && multi_signal) "indicator" else character(0)),
        all_of(params$extra_keys), n_locs, median_cv
    )

DT::datatable(noise_summary,
    colnames = c(
        if (has_signal && multi_signal) "Indicator" else NULL,
        str_to_title(str_replace_all(params$extra_keys, "_", " ")),
        "Locations", "Median CV"
    ),
    caption = sprintf(
        "CV summary across locations. %d of %d (location%s) series have a defined CV%s%s.",
        nrow(noise_stats), n_series,
        if (length(params$extra_keys) > 0) " x stratum" else "",
        if (nrow(noise_undefined) > 0) {
            sprintf("; %d dropped as undefined (zero mean or a single reported value)", nrow(noise_undefined))
        } else "",
        if (n_excluded > 0) sprintf("; a further %d above the 99th percentile are excluded from the plot", n_excluded) else ""
    ),
    rownames = FALSE, options = list(scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE, paging = TRUE, dom = "tp")
) %>% DT::formatRound(columns = ncol(noise_summary), digits = 3)

# Histogram (faceted by indicator x stratum)
ggplot(noise_stats_plot, aes(x = cand_cv)) +
    geom_histogram(bins = 30, color = "white", fill = dis_pal_list[2], alpha = 0.7) +
    scale_x_continuous(labels = label_percent()) +
    {if (multi_signal || length(params$extra_keys) > 0) facet_wrap(reformulate(c(if (multi_signal) "indicator", params$extra_keys)), scales = "free", labeller = label_wrap_html(width = 25)) else NULL} +
    labs(
        title = "CV Distribution by Location",
        subtitle = "Higher values mean more variation over time, from real dynamics or from noise",
        x = "Coefficient of Variation (CV)", y = "Count"
    ) +
    theme(legend.position = "none")
```

:::

:::


::: {.content-visible when-meta="versioning"}

### Candidate Versioning

Version dynamics is not the focus here. This section answers only whether the `r ind` `r sg_pl("revises", "revise")` at all and roughly how much. Relative revisions are scaled to the `r ind_poss` own magnitude: 0.1 means the observation shifted by 10% of its own size, and values near 1 mean revisions as large as the `r ind` `r sg_pl("itself", "themselves")`.

::: {.callout-tip}
For revision lags by location, convergence over time, and how much of the latest value is knowable at each lag, use `revision_analysis/revision_analysis.qmd`.
:::

```{r}
#| label: revision-analysis-text
#| echo: false
#| results: asis
#| eval: !expr show_versioning && !is.null(archive)
if (!archive_has_history) {
    version_start <- min(archive$DT$version)
    cat(sprintf(
        "Versioning began on %s. At least 60 days of version history are needed for meaningful revision analysis; the earliest this will be possible is %s.\n\n",
        format(version_start), format(version_start + 60)
    ))
    cat(sprintf(
        "Current versions: %s.\n\n",
        paste(sort(unique(archive$DT$version)), collapse = ", ")
    ))
} else {
    cat("> The `revision_summary` is under revision and may produce misleading results in certain edge cases. Use the summary statistics below with caution while development is ongoing.\n\n")

    if (exists("archive_info") && !is.null(archive_info)) {
        cat(sprintf(
            "Revision analysis is performed on a sample of %d out of %d locations, with versions from %s to %s, and time from %s to %s.\n\n",
            archive_info$n_sampled, archive_info$n_total, archive_info$version_start, archive_info$version_end, archive_info$time_value_start, archive_info$time_value_end
        ))
    }

    # A revision is "quick" within 3 days, "few" at 3 or fewer, and its spread is
    # minor below 10% of the observation (or major past 5% of the indicator's max).
    quick_threshold <- 3
    few_threshold <- 3
    rel_spread_threshold <- 0.1
    fmt_summary <- function(x) {
        s <- summary(as.numeric(x, units = "days"))
        paste(names(s), round(as.numeric(s), 1), sep = ": ", collapse = ", ")
    }

    # Per indicator, so the statistics are never pooled across them. The
    # stratifiers stay in the key: dropping them collapses one row per stratum
    # onto the same (geo, period, version), which as_epi_archive() rejects.
    for (sig in signals_vec) {
        if (multi_signal) cat(sprintf("\n#### %s\n\n", signal_labels[sig]))

        sub_archive <- archive$DT %>%
            as_tibble() %>%
            filter_signal(sig) %>%
            select(all_of(c("geo_value", "time_value", "version", params$extra_keys, "value"))) %>%
            as_epi_archive(other_keys = params$extra_keys)

        rd <- revision_summary(sub_archive)
        rb <- rd$revision_behavior
        abs_spread_threshold <- 0.05 * rd$max_val

        cat(sprintf("- **Time range**: %s to %s\n", rd$range_time_values[1], rd$range_time_values[2]))
        cat(sprintf("- **Min lag (days to first version)**: %s\n", fmt_summary(rb$min_lag)))
        cat(sprintf("- **Total number of versions**: %d\n", n_distinct(sub_archive$DT$version)))
        cat(sprintf("- **No revisions**: %s\n", label_percent()(mean(rb$n_revisions == 0))))
        cat(sprintf("- **Quick revisions (<= %d days)**: %s\n", quick_threshold,
                    label_percent()(mean(as.numeric(rb$max_lag, units = "days") <= quick_threshold))))
        cat(sprintf("- **Few revisions (<= %d)**: %s\n", few_threshold,
                    label_percent()(mean(rb$n_revisions <= few_threshold))))

        revised_beh <- rb %>% filter(n_revisions > 0)
        if (nrow(revised_beh) > 0) {
            cat(sprintf("- **Minor relative revisions (< %.1f)**: %s\n", rel_spread_threshold,
                        label_percent()(mean(revised_beh$rel_spread < rel_spread_threshold, na.rm = TRUE))))
            cat(sprintf("- **Major absolute revisions (> %.3f)**: %s\n", abs_spread_threshold,
                        label_percent()(mean(revised_beh$spread > abs_spread_threshold, na.rm = TRUE))))
        }
        cat(sprintf("- **Days until within %d%% of latest**: %s\n\n",
                    rd$within_latest * 100, fmt_summary(rb$lag_near_latest)))
    }

}
```

:::



### Indicator sanity checks

These are hard to automate. Things to look for:

- Do the values match what you expected? This catches the most, and only you can make the call: state the range, level and shape you expect from the `r ind_poss` definition before reading the plots, then compare.
- If it's a percentage or percentile, is it between 0 and 100?
- If it's a count, does it scale with population? One that doesn't points at a coverage or denominator problem. A rate or percentage is already normalized, so its level should instead be comparable across locations of very different sizes, and one far off the rest is worth investigating. The relevant denominator is usually not the resident population but the part the source observes.
- Are the values plausible? A percentage approaching 100 outside a pandemic, for instance.
- For a seasonal disease, do rates vary as expected?
- Do the coarser geographies aggregate plausibly?

The `r sg_pl("histogram", "histograms")` and table below show the distribution of all values and the range per location, to judge whether that range fits the `r ind_poss` expected scale.

```{r}
#| label: sanity-value-range
#| fig-height: !expr min(max(5, length(signals_vec) * 0.5), 12)
df_candidate_plot <- df_candidate %>% add_indicator()

ggplot(df_candidate_plot, aes(x = value)) +
    geom_histogram(bins = 40, alpha = 0.8, color = "white") +
    scale_x_continuous(labels = label_number(scale_cut = cut_si(""))) +
    scale_y_continuous(labels = label_number(scale_cut = cut_si(""))) +
    {if (multi_signal) facet_wrap(~indicator, scales = "free", labeller = label_wrap_html(width = 25)) else NULL} +
    labs(
        title = if (multi_signal) "Value distribution by indicator"
                else sprintf("Value distribution: %s", signal_labels[signals_vec]),
        subtitle = sprintf(
            "%s reported values across %d locations",
            label_comma()(nrow(df_candidate)), n_locations
        ),
        x = "Value", y = "Count"
    )
```

```{r}
#| label: sanity-value-range-table
#| echo: false
range_tbl <- df_candidate %>%
    as_tibble() %>%
    filter(!is.na(value)) %>%
    group_by(across(all_of(c(strat_keys, "geo_value")))) %>%
    summarize(
        n = n(),
        min = min(value),
        p05 = quantile(value, 0.05),
        median = median(value),
        p95 = quantile(value, 0.95),
        max = max(value),
        .groups = "drop"
    ) %>%
    mutate(
        across(all_of(params$extra_keys), order_strata),
        geo_value = geo_to_name(geo_value)
    ) %>%
    add_indicator() %>%
    arrange(desc(max)) %>%
    select(
        any_of(if (has_signal && multi_signal) "indicator" else character(0)),
        all_of(params$extra_keys), geo_value, n:max
    )

DT::datatable(
    range_tbl,
    colnames = c(
        if (has_signal && multi_signal) "Indicator" else NULL,
        str_to_title(str_replace_all(params$extra_keys, "_", " ")),
        "Location", "N", "Min", "5th pct", "Median", "95th pct", "Max"
    ),
    caption = sprintf(
        "Value range per location%s. Sort by Min or Max to surface potential out-of-bounds values.",
        if (multi_signal) "" else sprintf(" for %s", signal_labels[signals_vec])
    ),
    rownames = FALSE,
    options = list(
        scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
        paging = TRUE, dom = "tp"
    )
) %>% DT::formatRound(columns = which(names(range_tbl) %in% c("min", "p05", "median", "p95", "max")), digits = 3)
```

The seasonal pattern of the `r ind`, one panel per epidemiological year (`r season_label`). The line is the monthly median across locations, the inner band the 25th-75th percentile range (the middle 50% of that month's observations), the outer band the full min-max range.

The year boundary comes from `season_start_month`, which defaults to July so a northern-hemisphere respiratory season falls inside one panel, following the usual influenza convention. Set another month for indicators peaking elsewhere in the year.

```{r}
#| label: sanity-seasonality
#| fig-height: !expr min(max(5, length(signals_vec) * n_key_combos * 2.5), 15)
#| eval: !expr as.numeric(difftime(max(df_candidate$time_value), min(df_candidate$time_value), units = "days")) >= 365
df_seasonal <- df_candidate %>%
    as_tibble() %>%
    add_season() %>%
    group_by(across(all_of(c(strat_keys, "season", "season_month")))) %>%
    summarize(
        median_val = median(value, na.rm = TRUE),
        min_val = min(value, na.rm = TRUE),
        max_val = max(value, na.rm = TRUE),
        p25 = quantile(value, 0.25, na.rm = TRUE),
        p75 = quantile(value, 0.75, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    # Keep only (key x season) combos with >= 3 months of data
    semi_join(
        count(., across(all_of(c(strat_keys, "season")))) %>% filter(n >= 3),
        by = c(strat_keys, "season")
    ) %>%
    add_indicator() %>%
    label_strata_coverage(params$extra_keys)

ggplot(df_seasonal, aes(x = season_month, color = season, fill = season, group = season)) +
    geom_ribbon(aes(ymin = min_val, ymax = max_val), alpha = 0.10, linewidth = 0.2) +
    geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.30, linewidth = 0.3) +
    geom_line(aes(y = median_val), linewidth = 0.8) +
    geom_point(aes(y = median_val), size = 1.5) +
    {
        row_vars <- c(if (multi_signal) "indicator", params$extra_keys)
        if (length(row_vars) > 0) {
            facet_grid(as.formula(paste(paste(row_vars, collapse = " + "), "~ season")), scales = "free_y")
        } else {
            facet_wrap(~season, ncol = 2, scales = "free_y")
        }
    } +
    scale_x_continuous(labels = ~ month.abb[((season_start_month - 2L + .x) %% 12L) + 1L]) +
    scale_y_continuous(labels = label_number(scale_cut = cut_si(""))) +
    labs(
        title = if (multi_signal) "Seasonal patterns by indicator"
                else sprintf("Seasonal pattern: %s", signal_labels[signals_vec]),
        subtitle = "Median by month across all locations — inner band: 25th-75th pct, outer band: min-max",
        x = "Month", y = "Median value"
    ) +
    theme(legend.position = "none")
```

```{r}
#| label: sanity-geo-aggregation
#| echo: false
#| results: asis
#| eval: !expr params$geo_type == "state"
pacman::p_load(nanoparquet)

crosswalk_path <- here("tests", "test_data", "delphi_geomap", "geomapper_crosswalk_20200101.parquet")
crosswalk <- tryCatch(
    {
        nanoparquet::read_parquet(crosswalk_path) %>%
            distinct(state_geo = state_abbr, hhs_region = as.character(hhs))
    },
    error = function(e) {
        # Fallback to standard HHS region mapping if the file is missing
        tibble(
            state_geo = c(
                "CT", "ME", "MA", "NH", "RI", "VT",
                "NJ", "NY", "PR", "VI",
                "DE", "DC", "MD", "PA", "VA", "WV",
                "AL", "FL", "GA", "KY", "MS", "NC", "SC", "TN",
                "IL", "IN", "MI", "MN", "OH", "WI",
                "AR", "LA", "NM", "OK", "TX",
                "IA", "KS", "MO", "NE",
                "CO", "MT", "ND", "SD", "UT", "WY",
                "AZ", "CA", "HI", "NV", "AS", "GU", "MP",
                "AK", "ID", "OR", "WA"
            ),
            hhs_region = c(
                rep("1", 6),
                rep("2", 4),
                rep("3", 6),
                rep("4", 8),
                rep("5", 6),
                rep("6", 5),
                rep("7", 4),
                rep("8", 6),
                rep("9", 7),
                rep("10", 4)
            )
        )
    }
)

is_pct_indicator <- any(grepl("%|pct", name_vec, ignore.case = TRUE))

fetch_coarser_candidate <- function(geo_type_coarser) {
    tryCatch(
        # pub_covidcast accepts a vector of signals
        pub_covidcast(
            params$source, signals_vec,
            geo_type_coarser, params$time_type,
            time_values = time_range(params$start_day, params$end_day),
            as_of = Sys.Date()
        ) %>% select(time_value, signal, geo_value, coarser_val = value),
        error = function(e) NULL
    )
}

# The HHS/nation series is unstratified, so only the aggregate stratum is
# comparable. Pooling every stratum widens the bounds until nothing can fall
# outside them and turns the state count into a row count.
agg_stratum <- map(set_names(params$extra_keys), function(k) {
    lv <- unique(as.character(df_candidate[[k]]))
    lv[tolower(lv) %in% agg_level_names][1]
})
strata_comparable <- all(!is.na(unlist(agg_stratum))) || length(params$extra_keys) == 0

state_vals <- df_candidate %>%
    as_tibble() %>%
    {
        if (strata_comparable) reduce(params$extra_keys, function(d, k) filter(d, .data[[k]] == agg_stratum[[k]]), .init = .)
        else .
    } %>%
    select(all_of(signal_key), time_value, state_geo = geo_value, state_val = value)

check_aggregation <- function(df_coarser, group_label, state_to_coarser) {
    if (is.null(df_coarser) || !strata_comparable) {
        return(NULL)
    }
    # No signal column means a single indicator, so match on period and geo.
    if (!has_signal) df_coarser <- df_coarser %>% select(-any_of("signal"))

    state_bounds <- state_vals %>%
        inner_join(state_to_coarser, by = "state_geo") %>%
        group_by(across(all_of(c(signal_key, "time_value", "coarser_geo")))) %>%
        summarize(
            state_min = min(state_val, na.rm = TRUE),
            state_max = max(state_val, na.rm = TRUE),
            n_states  = n_distinct(state_geo),
            .groups   = "drop"
        )

    df_coarser %>%
        rename(coarser_geo = geo_value) %>%
        inner_join(state_bounds, by = c(signal_key, "time_value", "coarser_geo")) %>%
        filter(!is.na(coarser_val) & (coarser_val < state_min - 1e-9 | coarser_val > state_max + 1e-9)) %>%
        add_indicator() %>%
        mutate(
            geo_level = group_label,
            direction = if_else(coarser_val < state_min, "below min", "above max")
        ) %>%
        select(geo_level, indicator, time_value, coarser_geo, coarser_val, state_min, state_max, direction, n_states)
}

df_hhs_raw <- fetch_coarser_candidate("hhs")
df_nation_raw <- fetch_coarser_candidate("nation")

hhs_map <- crosswalk %>% select(state_geo, coarser_geo = hhs_region)
nation_map <- crosswalk %>%
    distinct(state_geo) %>%
    mutate(coarser_geo = "us")

violations <- bind_rows(
    check_aggregation(df_hhs_raw, "HHS", hhs_map),
    check_aggregation(df_nation_raw, "Nation", nation_map)
)

unavailable <- c(
    if (is.null(df_hhs_raw)) "HHS" else NULL,
    if (is.null(df_nation_raw)) "Nation" else NULL
)

if (length(unavailable) > 0) {
    cat(sprintf(
        "**Note**: %s-level data not available for %s / %s.\n\n",
        paste(unavailable, collapse = " and "),
        params$source, paste(signals_vec, collapse = ", ")
    ))
}

# Which slice of the state data the coarser values were compared against.
strata_note <- if (length(params$extra_keys) == 0 || !strata_comparable) {
    ""
} else {
    sprintf(" The comparison uses the aggregate stratum (%s), the only one the unstratified HHS/nation series corresponds to.",
            paste(sprintf("%s = %s", names(agg_stratum), unlist(agg_stratum)), collapse = ", "))
}

if (!strata_comparable) {
    cat(sprintf(
        "**Note**: geo-aggregation check skipped. The HHS/nation series is not stratified, so it can only be compared against an aggregate level of %s, and none of the levels present (%s) is one.\n",
        paste(params$extra_keys, collapse = "/"),
        paste(map_chr(params$extra_keys, ~ paste(unique(as.character(df_candidate[[.x]])), collapse = ", ")), collapse = "; ")
    ))
} else if (length(unavailable) == 2) {
    # Nothing to compare against, so this is not a clean bill of health.
    cat("No coarser series was available, so the geo-aggregation check did not run.\n")
} else if (nrow(violations) == 0) {
    cat(sprintf(
        "No geo-aggregation violations found: all HHS/nation values are within the [min, max] of constituent state values for %s.%s\n",
        paste(name_vec, collapse = ", "), strata_note
    ))
} else {
    cat(sprintf(
        "**%d observations** where the aggregated value falls outside the [min, max] of its constituent states (%s indicator):\n\n",
        nrow(violations),
        if (is_pct_indicator) "percentage" else "non-percentage"
    ))
    print(htmltools::tagList(
        DT::datatable(
            violations,
            colnames = c("Level", "Indicator", "Date", "Geo", "Agg Value", "State Min", "State Max", "Direction", "N States"),
            caption = sprintf(
                "Geo-aggregation bounds check for %s. Aggregated values should fall within [min, max] of constituent state values.",
                paste(name_vec, collapse = ", ")
            ),
            rownames = FALSE,
            options = list(
                scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
                paging = TRUE, pageLength = 20, dom = "tp"
            )
        ) %>% DT::formatRound(columns = 5:7, digits = 3)
    ))
}
```





```{r}
#| label: sanity-seasonality-anomalies
#| echo: false
#| results: asis
has_full_year <- as.numeric(difftime(
    max(df_candidate$time_value), min(df_candidate$time_value), units = "days"
)) >= 365

if (!has_full_year) {
    cat("_Seasonal peak comparison needs at least one full year of history. Not enough data yet._\n")
} else {

cat("The table below flags locations whose seasonal peak is at least twice the typical (median) peak across all locations in the same season.\n\n")

df_seasonal_loc <- df_candidate %>%
    as_tibble() %>%
    add_season() %>%
    group_by(across(all_of(c(strat_keys, "season", "geo_value")))) %>%
    summarize(loc_max = max(value, na.rm = TRUE), .groups = "drop")

typical_max <- df_seasonal_loc %>%
    group_by(across(all_of(c(strat_keys, "season")))) %>%
    summarize(typical_max = median(loc_max, na.rm = TRUE), .groups = "drop")

anomalies <- df_seasonal_loc %>%
    left_join(typical_max, by = c(strat_keys, "season")) %>%
    mutate(
        ratio = loc_max / typical_max,
        across(all_of(params$extra_keys), order_strata),
        geo_value = geo_to_name(geo_value)
    ) %>%
    add_indicator() %>%
    filter(ratio >= 2) %>%
    arrange(desc(ratio)) %>%
    select(
        any_of(if (has_signal && multi_signal) "indicator" else character(0)),
        all_of(params$extra_keys), season, geo_value, loc_max, typical_max, ratio
    )

if (nrow(anomalies) == 0) {
    cat("**(No locations flagged.)** No location's seasonal max exceeds 2x the typical (median) seasonal max.\n")
} else {
    print(htmltools::tagList(
        DT::datatable(
            anomalies,
            colnames = c(
                if (has_signal && multi_signal) "Indicator" else NULL,
                str_to_title(str_replace_all(params$extra_keys, "_", " ")),
                "Season", "Location", "Location Max", "Typical Max", "Ratio"
            ),
            caption = sprintf(
                "Locations where seasonal max exceeds 2x the typical (median) seasonal max%s.",
                if (multi_signal) "" else sprintf(" for %s", signal_labels[signals_vec])
            ),
            rownames = FALSE,
            options = list(
                scrollX = TRUE, scrollY = "300px", scrollCollapse = TRUE,
                paging = TRUE, pageLength = 20, dom = "tp"
            )
        ) %>% DT::formatRound(columns = which(names(anomalies) %in% c("loc_max", "typical_max", "ratio")), digits = 3)
    ))
}

}
```
